### 2015-11-22 description/description.tex
init

\documentclass{scrartcl} 
\usepackage[ngerman,english]{babel}
 \usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[applemac]{inputenc}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsthm}

\usepackage{fancyhdr}
\usepackage{paralist}

\usepackage{mathtools}

\usepackage{hyperref}

\DeclareMathOperator*{\esssup}{ess.sup}
\DeclareMathOperator*{\defgl}{\vcentcolon=}

%\pagestyle{fancy}
%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
%\fancyhead[C]{} %zentrierte Kopfzeile
%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the PRM Method}
\date{}
\author{Manuel Demmeler}

%\thispagestyle{fancy}

\newcommand{\3}{ ^{3\times3} }
\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\M}{\mathbb{M}}
\newcommand{\Sym}{\mathbb{S}^3}
\newcommand{\Mpos}{\M_+^3}
\newcommand{\Orth}{\mathbb{O}^3}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\abs}[1]{\left |#1\right |}
\newcommand{\ub}{\bar{u}}
\newcommand{\yb}{\bar{y}}
\newcommand{\scprod}[2]{\langle#1 , #2 \rangle}

%\vspace{-3.4cm}

\begin{document}
\maketitle

\section{Motivation}
In trajctory optimization with obstacles, it is often necessary to provide feasible inital trajectories. 
For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive. A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. An efficient approximative approach is using probabilistic roadmaps (PRMs).

\section{Problem Statement and Sequential PRM Method}
\newtheorem*{defi}{Definition}
\newtheorem*{alg1}{Algorithm}

The abstract problem is to find a path on a d-dimensional map, which avoids intersecting obstacles. The obstacles are  given by an indicator function
\begin{equation*}
	I: \Omega \rightarrow \{0, 1\}
\end{equation*}
on the domain \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \) and the goal is to find a continuous way
\begin{equation*}
	\Gamma: [0,1] \rightarrow \Omega
\end{equation*}
between a given start and endpoint, \(\Gamma(0)=q_b,\ \Gamma(1)=q_e \), 
such that
\begin{equation*}
	I(\Gamma(s))=0 \text{\ f.a.\ } sÊ\in [0,1].
\end{equation*}
The probabilistic roadmap (PRM) approach grows a graph \(G=(V,E)\) form \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in V\subset \Omega\).  Two nodes \(v,w \in V\) are connected, if they are near enough each other and the linear connection in between is free from obstacles. The algorithm terminates, when there is a path from \(q_s\) to \(q_e\) on G.
\begin{defi}
For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
	\begin{equation*}
		[q_1,q_2]_h \defgl \{ q = \lambda q_1+(1-\lambda) q_2\ |\ \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
	\end{equation*}
	We say, \(q_1\) and \(q_2\) are connected (with stepsize \(h>0\)) on the map \(\Omega\), if
	\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
\end{defi}
\begin{alg1} (Probabilistic Roadmap in general)\\
	Input: \(q_s, q_e\in \Omega, h>0, d_0>0\)	\\
	Output: \(q_s=q_1, ... ,q_n=q_e\in \Omega\), 
		such that all \(q_i\) and \(q_{i+1}\) are connected with stepsize \(h\)
	\begin{itemize}
		\item Initialize graph \(G=(V,E)\) with \(V=\{q_s, q_e\}\), 
			if \(q_s\) and \(q_e\) are connected, \(E \leftarrow\{ q_s, q_e\}\)
		\item While there exists no path in G between \(q_s\) and \(q_e\) do
		\begin{itemize}
			\item Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\)
			\item \(V \leftarrow{q}\)
			\item For all nodes \(\tilde{q} \in V\) which are near enough (\( \norm{q- \tilde q} \le d_0 \)) and
				connected with \(q\) with stepsize h, 
				\(E \leftarrow \{q, \tilde{q}\}\)
		\end{itemize}
		\item Determine the shortest path \(q_1, ... , q_n\in V\) on \(G\) from \(q_s\) to \(q_e\).
	\end{itemize}
\end{alg1}

The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first select randomly a node \(q_0 \in V\) and then sample a new node \(q\) in the neighbourhood of \(q_0\). Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.

\section{Application to Robotics}

In the project, it is planned to use this algorithm to create collisionfree trajectories of a robot arm. 
For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot. Therefore the direct kinematic can be interpreted as a (strongly nonlinear) function
\begin{equation*}
	f: (q_1, ... , q_d) \mapsto (x_1, R_1, ..., x_l, R_l).
\end{equation*}

Now, we have given some obstacles with fixed positions \(x_{l+1}, ..., x_m\) and orientations \(R_{l+1}, ..., R_{m}\), which the robot has to avoid during movement. It is assumed that all objects are given in a form, that for each two objects it can be determined from the positions and orientations, if there is an intersection. For example as polyhedrons.
From this the indicator function can be defined as
\begin{equation*}
	I(q)=
	\begin{cases}
		0,	& \text{if no collision for all object pairs\ } (x_j, R_j), (x_k,R_k),\ j,k\in 1, ... , m \\
		1,	& \text{for at least one collision}
	\end{cases}
\end{equation*}
Now, if we additionally introduce jointlimits and call the set of possible angles \(\Omega\), we see, that the problem of finding feasible trajectories is exactly the pathfinding problem from above.

It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable. Hence, they have to be smoothed before the use on a robot. This is expectedly not part of the project.

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{configspacetrafo.png}
    \caption{Example for the transformation of obstacles into configspace. 
    		The scattered areas on the right are the sets of points with I(q)=1, \cite{prmlec}.}
    \label{fig:awesome_image}
\end{figure}

\section{Ideas for Parallelization}

Like the stated robotics example, in many cases the most costly operation of the PRM algorithm is the evaluation of the indicator function \(I(q)\), which however can be done independently for every new \(q\).
\begin{itemize}
\item 
Therefore in a first step, it is planned to implement \(I(q)\) parallely in a vectorized way, i.e. as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\). For example with help of the GPU. In the robotics example this means, the direct kinematic and collision calculation is made parallely for \(P\) different joint angle values.
\item
Then in a first level of parallelization, this function can be applied to the \(q_i \in [q, \tilde q]_h \) to check, if \(q\) and \(\tilde q\) are connected.
\item
Furthermore, a second level of parallelization could be achieved by sampling and connecting several new points \(\tilde q_i\) parallely. This is expected to be more difficult, because all new nodes have to be connected to the old graph and to each other. Therefore it has to be evaluated, if a parallel implementation stays efficient here.
\item
If there remains time, a next idea could be to execute the algorithm on different discretization levels \(h\). This means for example to compute a way \(q_1^0, ..., q_n^0\) which is connected with a coarse stepsize \(h_0\), then to decrease the stepsize to \(0<h_1<h_0\) and take the path nodes \(q_i^0\) as starting nodes for the next execution of the algorithm, which gives then an \(h_1\)-connected way \(q_1^1, ... , q_n^1\), etc.
\end{itemize}

\bibliographystyle{alpha}
\begin{thebibliography}{999}
	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, source: http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html
	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press, 2005
	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press, 2006
	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl fŸr Angewandte Mechanik, TU MŸnchen, 2014
\end{thebibliography}

\end{document}

### 2015-12-24 description/description.tex
prm projekt angefangen

@@
 
-\section{Problem Statement and Sequential PRM Method}
+\section{Problem Statement and Sequential PRM Method}\label{algintro}
 \newtheorem*{defi}{Definition}
@@
 
-The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first select randomly a node \(q_0 \in V\) and then sample a new node \(q\) in the neighbourhood of \(q_0\).
+The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\).
 Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.
@@
 In the project, it is planned to use this algorithm to create collisionfree trajectories of a robot arm. 
-For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot.
-Therefore the direct kinematic can be interpreted as a (strongly nonlinear) function
+For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot, \cite{robodyn} describes the exact calculations.
+Therefore this direct kinematic can be interpreted as a (strongly nonlinear) function
 \begin{equation*}
@@
 For example as polyhedrons.
-From this the indicator function can be defined as
+From this the indicator function for the obstacles in the configuration space can be defined as
 \begin{equation*}
@@
 
-Like the stated robotics example, in many cases the most costly operation of the PRM algorithm is the evaluation of the indicator function \(I(q)\), which however can be done independently for every new \(q\).
+Like in the stated robotics example, in many cases the most costly operation of the PRM algorithm is the evaluation of the indicator function \(I(q)\), which however can be done independently for every new \(q\).
 \begin{itemize}
@@
 For example with help of the GPU.
-In the robotics example this means, the direct kinematic and collision calculation is made parallely for \(P\) different joint angle values.
+In the robotics example this means, the direct kinematic and collision calculation is made parallely for \(P\) different joint angle vectors.
 \item

### 2015-12-24 prm/lib/configspace.cpp

@@ -1,0 +1,3 @@
+#include "configspace.hpp"
+
+

### 2015-12-24 prm/lib/configspace.hpp

@@ -1,0 +1,16 @@
+#ifndef CONFIGSPACE_HPP
+#define CONFIGSPACE_HPP
+
+class Configspace
+{
+public:
+  virtual Configspace(int d_){d=d_;}
+  virtual ~Configspace();
+
+  virtual int indicator(double*q, int N);
+
+private:
+  int d;
+};
+
+#endif // CONFIGSPACE_HPP

### 2015-12-24 prm/lib/robokin.cpp

@@ -1,0 +1,1 @@
+#include "robokin.hpp"

### 2015-12-24 prm/lib/robokin.hpp

@@ -1,0 +1,130 @@
+#ifndef ROBOKIN_HPP
+#define ROBOKIN_HPP
+
+
+#include <math.h>
+#include <vector>
+#include "ssv.hpp"
+#include "util.hpp"
+
+using namespace std;
+using namespace ssv;
+
+typedef vector<double> vecd;
+typedef vector<int> veci;
+
+class trafomat;
+class robogeo;
+class robokin;
+
+class trafomat{
+public:
+  //!Trafo nach Denavit Hartenberg
+  trafomat(double q, double alpha, double a, double d){
+    double cq=cos(q);
+    double sq=sin(q);
+    double ca=cos(alpha);
+    double sa=sin(alpha);
+
+    val[0]=cq;    val[1]=-sq;   val[2]=0.0;  val[3]=a;
+    val[4]=sq*ca; val[5]=cq*ca; val[6]=-sa;  val[7]=-sa*d;
+    val[8]=sq*sa; val[9]=cq*sa; val[10]=ca;  val[11]=ca*d;
+  }
+  trafomat(){}
+
+  inline void apply(double* vec, double* res){
+    double vec0=vec[0];
+    double vec1=vec[1];
+    double vec2=vec[2];
+
+    res[0]=val[0]*vec0+val[1]*vec1+val[2]*vec2+val[3];
+    res[1]=val[4]*vec0+val[5]*vec1+val[6]*vec2+val[7];
+    res[2]=val[8]*vec0+val[9]*vec1+val[10]*vec2+val[11];
+  }
+
+  inline void apply(const trafomat &trafo, trafomat &restrafo){
+    double* tval=&(trafo.val);
+    double* res=&(restrafo->val);
+
+    for(int i=0;i<3;++i){
+      int i0,i4,i8;
+      double vec0=tval[i0=0+i];
+      double vec1=tval[i4=4+i];
+      double vec2=tval[i8=8+i];
+      res[i0]=val[0]*vec0+val[1]*vec1+val[2]*vec2;
+      res[i4]=val[4]*vec0+val[5]*vec1+val[6]*vec2;
+      res[i8]=val[8]*vec0+val[9]*vec1+val[10]*vec2;
+    }
+    double vec0=tval[3];
+    double vec1=tval[7];
+    double vec2=tval[11];
+    res[3]=val[0]*vec0+val[1]*vec1+val[2]*vec2+val[3];
+    res[7]=val[4]*vec0+val[5]*vec1+val[6]*vec2+val[7];
[…]

### 2015-12-24 prm/lib/ssv.cpp

@@ -1,0 +1,2 @@
+#include "ssv.hpp"
+

### 2015-12-24 prm/lib/ssv.hpp

@@ -1,0 +1,116 @@
+#ifndef SSV_ELEMENT_HPP
+#define SSV_ELEMENT_HPP
+
+#include <math.h>
+
+namespace ssv{
+
+  class ssvele;
+  class ssv1;
+  class ssv2;
+  class ssv3;
+
+  class ssvele
+  {
+  public:
+    ssvele(){}
+  };
+
+  class ssv1: ssvele{
+  public:
+    ssv1();
+
+    inline double dist(ssv1& ele);
+    inline double dist(ssv2& ele);
+    inline double dist(ssv3& ele);
+  private:
+    double p[3];
+    double r2;
+  };
+
+  class ssv2: ssvele{
+  public:
+    ssv2();
+
+    inline double dist(ssv1& ele);
+    inline double dist(ssv2& ele);
+    inline double dist(ssv3& ele);
+
+  private:
+    double p[3];
+    double d[3];
+    double l;
+    double r2;
+  };
+
+  class ssv3: ssvele{
+    ssv3();
+
+    inline double dist(ssv1& ele);
+    inline double dist(ssv2& ele);
+    inline double dist(ssv3& ele);
+
+  private:
+    double p1[3];
+    double d1[3];
+    double d2[3];
+    double l1, l2;
+    double r2;
+  };
+
+
+
+  inline double ssv1::dist(ssv1& ele){
+    return p[0]*ele.p[0]+p[1]*ele.p[1]+p[2]*ele.p[2];
+  }
+
+  inline double ssv1::dist(ssv2& ele){
+
+  }
+
+  inline double ssv1::dist(ssv3& ele){
+
+  }
+
+  inline double ssv2::dist(ssv1& ele){
+    return ele.dist(*this);
+  }
+
+  inline double ssv2::dist(ssv2& ele){
+
+  }
+
+  inline double ssv2::dist(ssv3& ele){
+
+  }
+
+  inline double ssv3::dist(ssv1& ele){
+    return ele.dist(*this);
+  }
+
+  inline double ssv3::dist(ssv2& ele){
+    return ele.dist(*this);
+  }
+
+  inline double ssv3::dist(ssv3& ele){
+
+  }
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+
+}
+
+
+#endif // SSV_ELEMENT_H

### 2015-12-24 prm/lib/util.hpp

@@ -1,0 +1,9 @@
+#ifndef UTIL_HPP
+#define UTIL_HPP
+
+#include <iostream>
+
+#define check(x) if(x){ std::cout<<"check "<<#x<<" failed"<<std::endl;}
+
+
+#endif // UTIL_HPP

### 2016-03-27 prm/src/colltest.cpp
deleted all old tests

@@ -1,46 +1,0 @@
-#include <iostream>
-
-#include "lib/geo4.h"
-
-#include "lib/kinematics.h"
-#include "lib/collision4.h"
-
-using namespace std;
-using namespace geo4;
-using namespace collision4;
-
-const float pi=3.14159265358;
-
-int main()
-{
-
-  trafo4 tp(0.0, 0.0, 0.0, 0.0);
-  trafo4 tq(1.0, 0.0, 0.0, -0.01);
-  trafo4 tr(-1.0, 0.0, 0.0005, 0.0);
-
-
-  polytope4 P;
-  generate_simplex(P, 1.0, 1.0, 1.0);
-
-  polytope4 Q;
-  generate_simplex(Q, 1.0, 1.0, 1.0);
-
-  polytope4 R;
-  generate_quader(R, 1.0, 1.0, 1.0);
-
-
-#if 0
-  p4print(P,tp);
-  p4print(Q,tq);
-  p4print(R,tr);
-#endif
-
-  printvar(seperating_vector_algorithm(P,Q,tp,tq));
-  printvar(seperating_vector_algorithm(P,R,tp,tr));
-  printvar(seperating_vector_algorithm(Q,R,tq,tr));
-
-
-  msg("finished");
-  return 0;
-}
-

### 2016-03-27 prm/src/colltest2.cpp

@@ -1,201 +1,0 @@
-#include <iostream>
-
-#define SILENT
-#define CUDA_IMPLEMENTATION
-
-#include "lib/cuda_head.h"
-
-#include "lib/geo4.h"
-#include "lib/collision4.h"
-#include "lib/polytope4data.h"
-#include "lib/kinematics.h"
-
-using namespace collision4;
-
-const float pi=3.14159265358;
-
-#ifdef CUDA_IMPLEMENTATION
-template<int ndof>
-__global__ void kernel(Robot<ndof>* robot, polytope4data* polydata, float* q, int* coll, int n){
-  int i = blockDim.x * blockIdx.x + threadIdx.x;
-  if(i<n){
-    dmsg("Hello");
-    coll[i]=0;
-    Kinematics<ndof> kin(robot);
-    kin.calculate(&q[i],n);
-    dmsg("Hello");
-    for(int dof0=0;dof0<=ndof;++dof0)
-    for(int dof1=dof0+1;dof1<=ndof;++dof1){
-      dmsg("Hello");
-      dprintvard(dof0);
-      dprintvard(dof1);
-      int numsys0=polydata->get_numsys(dof0), numsys1=polydata->get_numsys(dof1);
-      dprintvard(numsys0);
-      dprintvard(numsys1);
-      dmsg("Hello4");
-      dmsg("Hello4");
-      dmsg("Hello4");
-      dmsg("Hello4");
-      for(int k0=0;k0<numsys0;++k0)
-      for(int k1=0;k1<numsys1;++k1){
-        dmsg("Hello5");
-        polytope4 poly0, poly1;
-        polydata->get_polytope(poly0, dof0,k0);
-        polydata->get_polytope(poly1, dof1,k1);
-#if 1
-        dp4print(poly0);
-        dp4print(poly1);
-#endif
-        int res=seperating_vector_algorithm(poly0,poly1,kin.trafos[dof0],kin.trafos[dof1]);
-        if(res!=0){
-          coll[i]=res;
-        }
-      }
-    }
-  }//if
-}
-#endif
-
-
-template<int ndof>
-void kernel_(Robot<ndof>* robot, polytope4data* polydata, float* q, int* coll, int n){
-  for(int i=0;i<n;++i){
-    if(i<n){
-      coll[i]=0;
[…]

### 2016-03-27 prm/src/test.cpp

@@ -1,83 +1,0 @@
-#include <iostream>
-
-#include "lib/geo4.h"
-
-#include "lib/kinematics.h"
-#include "lib/collision4.h"
-
-using namespace std;
-using namespace geo4;
-using namespace collision4;
-
-const float pi=3.14159265358;
-
-int main()
-{ 
-
-#if 0
-  float4 u=make_float4(sqrt(2),0.0,0.0);
-  trafo4 T(0.0,0.0,pi/4,1.0);
-
-  f4print(u);
-  t4print(T);
-
-  T.apply(u);
-
-  f4print(u);
-
-  float4 v;
-
-  f4print(v);
-
-  T.apply(u,v);
-
-  f4print(u);
-  f4print(v);
-#endif
-
-#if 0
-  Robot<3> robot;
-  robot.a[0]=0.0;
-  robot.alpha[0]=0.0;
-  robot.q[0]=0.0;
-  robot.d[0]=0.0;
-  robot.types[0]=rotational;
-
-  robot.a[1]=1.0;
-  robot.alpha[1]=0.0;
-  robot.q[1]=0.0;
-  robot.d[1]=0.0;
-  robot.types[1]=rotational;
-
-  robot.a[2]=0;
-  robot.alpha[2]=0;
-  robot.q[2]=0;
-  robot.d[2]=0;
-  robot.types[2]=prismatic;
-
-  float q[3]={0.0,pi/4.0,0.9};
-
-  Kinematics<3> kin;
-  kin.robot=&robot;
-
-  kin.calculate(&q[0]);
-
-  t4print(kin.trafos[0]);
-  t4print(kin.trafos[1]);
-  t4print(kin.trafos[2]);
-
-  float4 p=make_float4(1.0,0.0,0.0);
-
-  f4print(p);
-
-  kin.trafos[2].apply(p);
-
-  f4print(p);
-#endif
-
-
-
-  cout<<"test.cpp"<<endl;
-  return 0;
-}
-

### 2016-03-27 prm/src/test2.cpp

@@ -1,94 +1,0 @@
-#include <iostream>
-
-#define CUDA_IMPLEMENTATION
-#include <cuda.h>
-
-
-#include "lib/geo4.h"
-#include "lib/kinematics.h"
-#include "lib/collision4.h"
-
-#include "lib/util.hpp"
-
-using namespace std;
-using namespace geo4;
-
-const float pi=3.14159265358;
-
-
-__global__ void kernel(float* q, float* res, int n){
-  int i = blockDim.x * blockIdx.x + threadIdx.x;
-  if(i<n){
-      float4 u=make_float4(1.0,0.0,0.0);
-      trafo4 T(0.0,0.0,q[i],0.0);
-      T.apply(u);
-      res[i]=u.y;
-  }
-}
-
-void handle(cudaError_t res){
-  if(res!=0){
-    cout<<"cudaError_t:  "<<cudaGetErrorString(res)<<endl;
-  }
-}
-
-int main()
-{
-
-  int device = 0;
-  cudaError_t res=cudaSetDevice(device);
-  if(res!=0){
-    cout<<"Error setting device: "<<cudaGetErrorString(res)<<endl;
-    return -1;
-  }
-
-  cudaDeviceProp p;
-  cudaGetDeviceProperties(&p, device);
-  cout << "Device: " << p.name << endl;
-  cout << "MP: " << p.multiProcessorCount << endl;
-  cout << "Compute: " << p.major << "." << p.minor << endl;
-  cout << "maxThreadsPerBlock: " << p.maxThreadsPerBlock << endl;
-  cout << "maxThreadsDim: " << p.maxThreadsDim <<endl;
-  cout << "maxGridSize: " << p.maxGridSize << endl;
-  cout << "warpSize: " << p.warpSize << endl;
-  cout << "totalGlobalMem: " << p.totalGlobalMem/1024/1024 << "MB" << endl;
-
-  int n = 32, BLOCK = 32, GRID = (n + BLOCK - 1)/BLOCK;
-
-  printvar(n);
-  printvar(GRID);
-  printvar(BLOCK);
-
-
-  float *hst_q=new float[n]();
-  float *dev_q;
-  float *hst_res=new float[n]();
-  float *dev_res;
-
-  for(int i=0;i<n;++i){
-    hst_q[i]=pi/4.0+i*pi/4.0/n;
-  }
-
[…]

### 2016-03-27 prm/src/test3.cpp

@@ -1,38 +1,0 @@
-#include <iostream>
-
-#include "lib/util.hpp"
-
-#include "lib/configspace.hpp"
-#include "lib/arrayconfigspace.hpp"
-
-using namespace std;
-
-
-int main()
-{
-  int h=480, b=640, n=h*b;
-  int* array=new int[n];
-  read_file("array.bin",array, n);
-
-  ArrayConfigspace space(array, b, h, 0.0, 1.0, 0.0, 1.0);
-
-  printvar(space.dim());
-  printvar(space.min(0));
-  printvar(space.max(0));
-  printvar(space.min(1));
-  printvar(space.max(1));
-  printvar(space.deltaq());
-
-  float qs[4]={0.8,0.1,
-               0.4,0.4};
-  float qe[4]={0.99,0.9,
-               0.4,0.4};
-
-  float res[2]={7,7};
-  space.indicator2(&qs[0],&qe[0],&res[0],2);
-  printarr(res,2);
-
-  //printarr(array,3000);
-
-  //printarr(array,n);
-}

### 2016-03-27 prm/src/test4.cpp

@@ -1,109 +1,0 @@
-
-#include <iostream>
-
-#include <cuda.h>
-
-#define CUDA_IMPLEMENTATION
-#include "lib/util.hpp"
-
-
-using namespace std;
-
-#define loop 3
-
-
-struct teststruct{
-    int *x;
-    int y;
-};
-
- __device__ void test(int i, int* x_){
-
-    dprintvard(data.x[0]);
-    dprintvard(data.x[1]);
-    dprintvard(data.y);
-    //for(int k=0;k<loop;++k){
-    //    dprintvard(k);
-    //}
-}
-
-__global__ void kernel(float* q, float* res, int n){
-  int i = blockDim.x * blockIdx.x + threadIdx.x;
-  if(i<n){
-    int x[10];
-    x[0]=8;
-    x[1]=4;
-    teststruct data;
-    data.x=x_;
-    data.y=5;
-    dmsg("Hello");
-    test(i,&x[0]);
-  }
-}
-
-void handle(cudaError_t res){
-  if(res!=0){
-    cout<<"cudaError_t:  "<<cudaGetErrorString(res)<<endl;
-  }
-}
-
-int main()
-{
-
-  int device = 0;
-  cudaError_t res=cudaSetDevice(device);
-  if(res!=0){
-    cout<<"Error setting device: "<<cudaGetErrorString(res)<<endl;
-    return -1;
-  }
-
-  cudaDeviceProp p;
-  cudaGetDeviceProperties(&p, device);
-  cout << "Device: " << p.name << endl;
-  cout << "MP: " << p.multiProcessorCount << endl;
-  cout << "Compute: " << p.major << "." << p.minor << endl;
-  cout << "maxThreadsPerBlock: " << p.maxThreadsPerBlock << endl;
-  cout << "maxThreadsDim: " << p.maxThreadsDim <<endl;
-  cout << "maxGridSize: " << p.maxGridSize << endl;
-  cout << "warpSize: " << p.warpSize << endl;
-  cout << "totalGlobalMem: " << p.totalGlobalMem/1024/1024 << "MB" << endl;
-
-  int n = 1, BLOCK = 1, GRID = (n + BLOCK - 1)/BLOCK;
-
-  printvar(n);
-  printvar(GRID);
-  printvar(BLOCK);
-
-
-  float *hst_q=new float[n]();
-  float *dev_q;
[…]

### 2016-03-27 prm/src/test5.cpp

@@ -1,155 +1,0 @@
-
-
-#include <iostream>
-
-#include "lib/util.hpp"
-#include <map>
-#include <vector>
-
-#include <math.h>
-
-#include <chrono>
-#include <cstdlib>
-#include <ctime>
-
-
-using namespace std;
-
-int pow(int x, int n){
-    if(n>1)return x*pow(x,n-1);
-    else if(n==1) return x;
-    else if(n==0) return 1;
-    else return 0;
-}
-
-class point{
-public:
-    point(const point& p){
-        val[0]=p.val[0];
-        val[1]=p.val[1];
-    }
-    point(){}
-    point(int x, int y){val[0]=x;val[1]=y;}
-    int val[2];
-};
-
-#define printpoint(P) printarr(P.val,2);
-#define many(x) x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
-    x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x x \
[…]

### 2016-03-27 prm/src/test7.cpp

@@ -1,255 +1,0 @@
-
-#include <iostream>
-
-#define SILENT
-#define NO_IO
-
-#include "lib/util.hpp"
-
-#include "lib/configspace.hpp"
-#include "lib/robotconfigspace.h"
-#include "lib/geo4.h"
-using namespace geo4;
-
-
-#include "lib/vertexlist3.h"
-
-using namespace std;
-using namespace collision4;
-
-
-//void build_example1(Robot<1>* &robot, polytope4data* &polydata)
-void build_example1(Robot<2>* &robot, polytope4* &polys, int* &sys, int &N){
-  robot=new Robot<2>;
-
-  robot->a[0]=1.1;
-  robot->alpha[0]=0.0;
-  robot->q[0]=0.0;
-  robot->d[0]=0.0;
-  robot->types[0]=prismatic;
-
-  robot->a[1]=0.0;
-  robot->alpha[1]=0.0;
-  robot->q[1]=0.0;
-  robot->d[1]=0.0;
-  robot->types[1]=rotational;
-
-
-  polys=new polytope4[2];
-  sys=new int[2];
-  N=2;
-
-  trafo4 t0(0.0, 0.0, 0.0, 0.0);
-  generate_quader(polys[0], 1.0, 1.0, 1.0);
-  transform(polys[0],t0);
-  sys[0]=0;
-
-  trafo4 t1(0.0, 0.0, 0.0, 0.0);
-  generate_quader(polys[1], 1.0, 1.0, 1.0);
-  transform(polys[1],t1);
-  sys[1]=2;
-
-  //trafo4 t2(0.0, -pi/2.0, 0.0, 0.0);
-  //generate_quader(P[2], 1.0, 1.0, 1.0);
-  //transform(P[2],t2);
-
-}
-
-template<int ndof>
-int load_config(std::string path, Robot<ndof>* &robot, polytope4* &polys, int* &sys, int &N, bool printmsg=false){
-
-    //! DH params
-
-
-  int ndof_loaded=0;
-  read_file(path+"/ndof.bin",&ndof_loaded,1);
-  if(ndof_loaded!=ndof){
-    msg("error: wrong ndof");
-    return -1;
-  }
-
-  robot=new Robot<ndof>;
-
-  read_file(path+"/dh/a.bin",&(robot->a[0]),ndof);
-  read_file(path+"/dh/alpha.bin",&(robot->alpha[0]),ndof);
[…]

### 2016-03-27 prm/src/test8.cpp

@@ -1,268 +1,0 @@
-#include <iostream>
-
-#include "lib/config.h"
-
-#include "lib/util.hpp"
-
-#include "lib/configspace.hpp"
-#include "lib/robotconfigspace.h"
-#include "lib/geo4.h"
-using namespace geo4;
-
-
-#include "lib/vertexlist3.h"
-
-using namespace std;
-using namespace collision4;
-
-
-//void build_example1(Robot<1>* &robot, polytope4data* &polydata)
-void build_example1(Robot<2>* &robot, polytope4* &polys, int* &sys, int &N){
-  robot=new Robot<2>;
-
-  robot->a[0]=1.1;
-  robot->alpha[0]=0.0;
-  robot->q[0]=0.0;
-  robot->d[0]=0.0;
-  robot->types[0]=prismatic;
-
-  robot->a[1]=0.0;
-  robot->alpha[1]=0.0;
-  robot->q[1]=0.0;
-  robot->d[1]=0.0;
-  robot->types[1]=rotational;
-
-
-  polys=new polytope4[2];
-  sys=new int[2];
-  N=2;
-
-  trafo4 t0(0.0, 0.0, 0.0, 0.0);
-  generate_quader(polys[0], 1.0, 1.0, 1.0);
-  transform(polys[0],t0);
-  sys[0]=0;
-
-  trafo4 t1(0.0, 0.0, 0.0, 0.0);
-  generate_quader(polys[1], 1.0, 1.0, 1.0);
-  transform(polys[1],t1);
-  sys[1]=2;
-
-  //trafo4 t2(0.0, -pi/2.0, 0.0, 0.0);
-  //generate_quader(P[2], 1.0, 1.0, 1.0);
-  //transform(P[2],t2);
-
-}
-
-template<int ndof>
-int load_config(std::string path, Robot<ndof>* &robot, polytope4* &polys, int* &sys, int &N, int* &from, int* &to, int& M, bool printmsg=false){
-
-    //! DH params
-
-
-  int ndof_loaded=0;
-  read_file(path+"/ndof.bin",&ndof_loaded,1);
-  if(ndof_loaded!=ndof){
-    msg("error: wrong ndof");
-    return -1;
-  }
-
-  robot=new Robot<ndof>;
-
-  read_file(path+"/dh/a.bin",&(robot->a[0]),ndof);
-  read_file(path+"/dh/alpha.bin",&(robot->alpha[0]),ndof);
[…]

### 2016-03-27 prm/src/test8cuda.cpp

@@ -1,271 +1,0 @@
-#include <iostream>
-
-
-#define CUDA_IMPLEMENTATION
-
-
-#define SILENT
-#define NO_IO
-
-#include "lib/util.hpp"
-
-#include "lib/configspace.hpp"
-#include "lib/robotconfigspace.h"
-#include "lib/geo4.h"
-using namespace geo4;
-
-
-#include "lib/vertexlist3.h"
-
-using namespace std;
-using namespace collision4;
-
-
-//void build_example1(Robot<1>* &robot, polytope4data* &polydata)
-void build_example1(Robot<2>* &robot, polytope4* &polys, int* &sys, int &N){
-  robot=new Robot<2>;
-
-  robot->a[0]=1.1;
-  robot->alpha[0]=0.0;
-  robot->q[0]=0.0;
-  robot->d[0]=0.0;
-  robot->types[0]=prismatic;
-
-  robot->a[1]=0.0;
-  robot->alpha[1]=0.0;
-  robot->q[1]=0.0;
-  robot->d[1]=0.0;
-  robot->types[1]=rotational;
-
-
-  polys=new polytope4[2];
-  sys=new int[2];
-  N=2;
-
-  trafo4 t0(0.0, 0.0, 0.0, 0.0);
-  generate_quader(polys[0], 1.0, 1.0, 1.0);
-  transform(polys[0],t0);
-  sys[0]=0;
-
-  trafo4 t1(0.0, 0.0, 0.0, 0.0);
-  generate_quader(polys[1], 1.0, 1.0, 1.0);
-  transform(polys[1],t1);
-  sys[1]=2;
-
-  //trafo4 t2(0.0, -pi/2.0, 0.0, 0.0);
-  //generate_quader(P[2], 1.0, 1.0, 1.0);
-  //transform(P[2],t2);
-
-}
-
-template<int ndof>
-int load_config(std::string path, Robot<ndof>* &robot, polytope4* &polys, int* &sys, int &N, int* &from, int* &to, int& M, bool printmsg=false){
-
-    //! DH params
-
-
-  int ndof_loaded=0;
-  read_file(path+"/ndof.bin",&ndof_loaded,1);
-  if(ndof_loaded!=ndof){
-    msg("error: wrong ndof");
-    return -1;
-  }
-
-  robot=new Robot<ndof>;
-
-  read_file(path+"/dh/a.bin",&(robot->a[0]),ndof);
[…]

### 2016-03-27 prm/src/lib/arrayconfigspace.cpp
new lib folder structure

@@ -1,146 +1,0 @@
-#include "arrayconfigspace.hpp"
-#include <math.h>
-
-ArrayConfigspace::ArrayConfigspace(const int* array_, int b_, int h_, float minx, float maxx, float miny, float maxy){
-  b=b_; h=h_; n=b*h;
-  mins[0]=minx;
-  mins[1]=miny;
-  maxs[0]=maxx;
-  maxs[1]=maxy;
-  factor[0]=b/(maxs[0]-mins[0]);
-  factor[1]=h/(maxs[1]-mins[1]);
-  dq=1.0/factor[0];
-  if(1.0/factor[1]<dq)dq=1.0/factor[1];
-  array=new int[n];
-  for(int i=0;i<n;++i){
-      array[i]=array_[i];
-  }
-}
-ArrayConfigspace::~ArrayConfigspace(){
-  delete array;
-}
-
-//!initialization function
-int ArrayConfigspace::init(const int ressource_rank, const int ressource_size){
-  return 0;
-}
-//! indicator function of obstacles
-//! q: length d*N, array of structures: q[N*k+i]= k-th component of i-th q-vector
-//! res: length N
-int ArrayConfigspace::indicator(const float* q, int* res, const int N, const int offset){
-  int iy=offset;
-  for(int ix=0;ix<N;++ix,++iy){
-    int x=(int)((q[ix]-mins[0])*factor[0]);
-    int y=(int)((q[iy]-mins[1])*factor[1]);
-    int resix=-1;
-    if(x>=0 && x<b && y>=0 && y<h){
-      int index=b*y+x;
-      resix=array[index];
-    }
-    res[ix]=resix;
-  }
-  return 0;
-}
-//!case N=1
-int ArrayConfigspace::indicator(const float* q){
-  int res=0;
-  indicator(q,&res,1,1);
-  return res;
-}
-
-//! checks if indicator function = 1 somewhere on the line between qs and qe
-//! res is return value
-int ArrayConfigspace::indicator2(const float* qs, const float* qe, int *res, const int N, const int offset){
-  for(int k=0;k<N;++k){
-    float qs_[]={qs[k],qs[k+offset]};
-    float qe_[]={qe[k],qe[k+offset]};
-    if(0 != check_boundaries(&qs_[0]) || 0 != check_boundaries(&qe_[0])){
-      res[k]=2;
-    }else{
[…]

### 2016-03-27 prm/src/lib/arrayconfigspace.hpp

@@ -1,51 +1,0 @@
-#ifndef ARRAYCONFIGSPACE_HPP
-#define ARRAYCONFIGSPACE_HPP
-
-#include "configspace.hpp"
-
-class ArrayConfigspace : public Configspace<2>
-{
-public:
-  ArrayConfigspace(const int* array_, int b_, int h_, float minx, float maxx, float miny, float maxy);
-  ~ArrayConfigspace();
-
-  //!initialization function
-  int init(const int ressource_rank=0, const int ressource_size=1);
-  //! indicator function of obstacles
-  //! q: length d*N, array of structures: q[N*k+i]= k-th component of i-th q-vector
-  //! res: length N
-  int indicator(const float* q, int* res, const int N, const int offset);
-  //!case N=1
-  int indicator(const float* q);
-  //! checks if indicator function = 1 somewhere on the line between (qs[i],qs[i+N]) and (qe[i],qe[i+N])
-  //! res is return value
-  int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
-  //! same paircheck as above, but with compressed storage:
-  //! checks pairs: (qs[i],...) ->  (qe(posqe[i]),...) , ...., (qe[posqe[i]+numqe[i]-1],...) for i=0,...,M-1
-  int indicator2(const float* qs, const int M, const float* qe, int *res, const int *posqe, const int *numqe, const int offset);
-  //! structure like indicator function
-  //! returns if lies in boundaries
-  int check_boundaries(const float* q, int* res, int N, const int offset);
-  //! for N=1
-  int check_boundaries(const float* q);
-
-  //! minimal value of qi, i=0...d-1
-  float min(int i);
-  //! maximal value of qi, i=0...d-1
-  float max(int i);
-  //! dq used for indcator2
-  float deltaq(){return dq;}
-  //! dimension d
-  int dim(){return 2;}
-private:
-  float mins[2];
-  float maxs[2];
-  float factor[2];
-  int* array;
-  int b;
-  int h;
-  int n;
-  float dq;
-};
-
-#endif // ARRAYCONFIGSPACE_HPP

### 2016-03-27 prm/src/lib/configspace.cpp

@@ -1,3 +1,0 @@
-#include "configspace.hpp"
-
-

### 2016-03-27 prm/src/lib/prm/configspace.hpp

@@ -1,6 +1,12 @@
 #ifndef CONFIGSPACE_HPP
 #define CONFIGSPACE_HPP
 
+
+//!
+//!  interface to pass different configuration spaces to the PRMSolver
+//!
+
+
 template<int ndof>
 class Configspace
 {

### 2016-03-27 prm/src/lib/prm/prmsolver.cpp

@@ -6,8 +6,8 @@
 #include <stdlib.h>
 #include <queue>
 
-#include "util.hpp"
-#include "tictoc.h"
+#include <util.hpp>
+#include <tictoc.h>
 
 
 

### 2016-03-27 prm/src/lib/prm/prmsolver.h

@@ -8,7 +8,7 @@
 
 #define MPI_CODE
 
-#include "config.h"
+#include <config.h>
 
 #include "configspace.hpp"
 

### 2016-03-27 prm/src/lib/robotspace/kinematics.h

@@ -2,10 +2,8 @@
 #define KINEMATICS_H
 
 
-#include "cuda_head.h"
-
-#include "geo4.h"
-//using namespace geo4;
+#include <collision4/cuda_head.h>
+#include <collision4/geo4.h>
 
 #include "robot.h"
 

### 2016-03-27 prm/src/lib/robotspace/robotconfigspace.cpp

@@ -1,14 +1,14 @@
 
-#include "config.h"
+#include <config.h>
 
 #include "robotconfigspace.h"
 
 
-#include "cuda_head.h"
-#include "util.hpp"
+#include <collision4/cuda_head.h>
+#include <util.hpp>
 
-#include "collision4.h"
-#include "polytope4data.h"
+#include <collision4/collision4.h>
+#include <collision4/polytope4data.h>
 #include "kinematics.h"
 
 #include <vector>

### 2016-03-27 prm/src/lib/robotspace/robotconfigspace.h

@@ -6,9 +6,10 @@
 #include <vector>
 #include <set>
 
-#include "configspace.hpp"
+#include <prm/configspace.hpp>
+#include <collision4/polytope.h>
+
 #include "robot.h"
-#include "polytope.h"
 
 template<int ndof> class Kinematics;
 

### 2016-03-27 prm/src/main.cpp

@@ -8,13 +8,11 @@
 #include "lib/util.hpp"
 #include "lib/tictoc.h"
 
-#include "lib/robotconfigspace.h"
-#include "lib/polytope.h"
-#include "lib/prmsolver.h"
+#include "lib/robotspace/robotconfigspace.h"
+#include "lib/prm/prmsolver.h"
 
 
 using namespace std;
-
 
 
 int main(int argc, char** argv)

### 2016-03-27 prm/src/test10.cpp

@@ -8,9 +8,8 @@
 #include "lib/util.hpp"
 #include "lib/tictoc.h"
 
-#include "lib/robotconfigspace.h"
-#include "lib/polytope.h"
-#include "lib/prmsolver.h"
+#include "lib/robotspace/robotconfigspace.h"
+#include "lib/prm/prmsolver.h"
 
 
 using namespace std;

### 2016-03-27 prm/src/lib/collision4/geo4.h
makefile

@@ -7,9 +7,7 @@
 #include <iomanip>
 
 
-//#ifndef CUDA_IMPLEMENTATION
-//#ifndef __CUDA_ARCH__
-#if 0
+#ifdef NO_CUDA
   struct float4{
     float x,y,z,w;
   };

### 2016-03-27 prm/src/lib/collision4/cuda_head.h
some warnings fixed

@@ -2,7 +2,7 @@
 #define CUDA_HEAD_H
 
 #undef qualifier
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
   #include <cuda.h>
   #define qualifier __host__ __device__ inline
   #define qualifierd __host__ __device__ inline
@@ -15,7 +15,7 @@
   #define hostonly(x) x
 #endif
 
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
   #define sin_ sinf
   #define cos_ cosf
   #define sqrt_ sqrtf
@@ -26,8 +26,8 @@
   #define sqrt_ sqrt
 #endif
 
-#ifndef CUDA_IMPLEMENTATION
-  #warning ("NO CUDA_IMPLEMENTATION: version without gpu");
+#ifndef GPU_VERSION
+  #pragma message("GPU_VERSION not defined: version without gpu");
 #endif
 
 

### 2016-03-27 prm/src/lib/collision4/geo4.h

@@ -7,7 +7,8 @@
 #include <iomanip>
 
 
-#ifdef NO_CUDA
+#ifndef CUDA_CODE
+  #pragma message("NO CUDA_CODE: defining float2 and float4");
   struct float4{
     float x,y,z,w;
   };
@@ -312,7 +313,7 @@ namespace geo4{
   }
 
 #ifndef SILENT
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
     #define df2print(v) printf("float2: %s= %f %f\n",#v, v.x,v.y);
     #define df4print(v) printf("float4: %s= %f %f %f\n",#v, v.x,v.y,v.z);
     #define dt4print(T) printf("trafo4: %s=\n%f\t%f\t%f\t%f\n%f\t%f\t%f\t%f\n%f\t%f\t%f\t%f\n\n",\

### 2016-03-27 prm/src/lib/collision4/polytope4data.h

@@ -69,7 +69,7 @@ namespace collision4{
         inline int build(const polytope4* polys, const int* sys, int N, int ndof_, const int* from_=0x0, const int* to_=0x0, int M_=0); //! arrays must be sortet w.r.t. sys
         inline int build(const polytope* polys, const int* sys, int N, int ndof_, const int* from_=0x0, const int* to_=0x0, int M_=0); //! arrays must be sortet w.r.t. sys
 
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
         friend int copy_host_to_device(polytope4data& devdata, const polytope4data& hostdata, bool withpairs_=false);
         friend int copy_device_to_host(polytope4data& hostdata, const polytope4data& devdata);
         //TODO: free memory
@@ -249,7 +249,7 @@ namespace collision4{
         return 0; //Todo error handling?
     }
 
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
     inline int copy_host_to_device(polytope4data& devdata, const polytope4data& hostdata, bool withpairs){
         int N=devdata.N=hostdata.N;
         int sumn=devdata.sumn=hostdata.sumn;

### 2016-03-27 prm/src/lib/config.h

@@ -4,7 +4,6 @@
 
 #define SILENT
 #define NO_IO
-//#define CUDA_IMPLEMENTATION
 
 
 #endif // CONFIG_H

### 2016-03-27 prm/src/lib/prm/prmsolver.cpp

@@ -74,6 +74,7 @@
   int PRMSolver<ndof>::init(const float* qstart, const float* qend){
     i0l=insert(qstart,graphl);
     i0r=insert(qend,graphr);
+    return 0;
   }
 
 
@@ -159,9 +160,18 @@
       }
     }
 
-    delete[] qnew, qnewall, qstartlist,qendlist, resbuf,
-            leftconn, rightconn, leftconnall, rightconnall,
-            surrnumnew,surrnumnewall,positionsall;
+    delete qnew;
+    delete qnewall;
+    delete qstartlist;
+    delete qendlist;
+    delete resbuf;
+    delete leftconn;
+    delete rightconn;
+    delete leftconnall;
+    delete rightconnall;
+    delete surrnumnew;
+    delete surrnumnewall;
+    delete positionsall;
 
     return 0;
   }
@@ -573,7 +583,8 @@
       return 1;
     }
 
-    delete[] poslist, distlist;
+    delete poslist;
+    delete distlist;
     return 0;
 
   }
@@ -691,7 +702,13 @@
       }
     }
 
-    delete[] qnew, qstartlist,qendlist, resbuf, resbufloc, leftconn, rightconn;
+    delete qnew;
+    delete qstartlist;
+    delete qendlist;
+    delete resbuf;
+    delete resbufloc;
+    delete leftconn;
+    delete rightconn;
 
     return 0;
   }
@@ -885,7 +902,8 @@
 
       MPI_Allgatherv(resbufloc,count,MPI_INT,resbuf,counts,disps,MPI_INT,MPI_COMM_WORLD);
 
-      delete[] counts, disps;
+      delete counts;
+      delete disps;
 
   #ifndef NO_IO
       printarr(qnew,ndof*num);
@@ -1007,7 +1025,8 @@
            if(res1==0){msg("ERROR: no path found by dijkstra in graphr");}
            //tock(tdijkstra);
 
-           delete[] poslist, distlist;
+           delete poslist;
+           delete distlist;
            return 1;
          }
 
@@ -1015,7 +1034,8 @@
 
 
 
-      delete[] poslist, distlist;
+      delete poslist;
+      delete distlist;
       return 0;
 
     }
@@ -1108,7 +1128,17 @@
       }
     }
 
-    delete[] qnew, qstartlist,qendlist, resbuf, resbufloc, leftconn, rightconn, poslist, distlist, dsp, cnt;
+    delete qnew;
[…]

### 2016-03-27 prm/src/lib/robotspace/robotconfigspace.cpp

@@ -77,7 +77,8 @@ RobotConfigspace<ndof>::RobotConfigspace(std::string path, const float dq_, cons
 
     construct(robot_, polys_, sys_, N_, from_, to_, M_, mins_, maxs_, dq_, nbuf_);
 
-    delete mins_, maxs_;
+    delete mins_;
+    delete maxs_;
 }
 
 template<int ndof>
@@ -205,7 +206,7 @@ int RobotConfigspace<ndof>::init(const int ressource_rank, const int ressource_s
   polydata=new collision4::polytope4data;
   polydata->build(polylist.polys,polylist.sys,polylist.N,ndof,polylist.from, polylist.to, polylist.M);
 
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
   int devcount;
   cudaGetDeviceCount(&devcount);
 
@@ -329,7 +330,7 @@ int RobotConfigspace<ndof>::indicator(const float* q)
 
 //!help kernels
 
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
 template<class T>
 __global__ void set_kernel(T *array, T val, int n){
     int i = blockDim.x * blockIdx.x + threadIdx.x;
@@ -346,7 +347,7 @@ __global__ void set_kernel(T *array, T val, int n){
 //! ************************
 
 template<int ndof>
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
 __global__ void kernel_indicator2(const Robot<ndof>* robot,
                                   const collision4::polytope4data* polydata,
                                   const float* qs, int offsets,
@@ -433,7 +434,7 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
     data.res=res;
     data.N=N;
 
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
     int id;
     if(free_resdevbuffer_ids.empty()){
         msg("RobotConfigspace: adding new resdevbuffer");
@@ -481,7 +482,7 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
       numthreads+=testnum[l];
     }
 
-  #ifdef CUDA_IMPLEMENTATION
+  #ifdef GPU_VERSION
     int GRID = (numthreads + BLOCK - 1)/BLOCK;
 
     for(int i=0;i<ndof;++i){
[…]

### 2016-03-27 prm/src/lib/util.hpp

@@ -5,6 +5,7 @@
 #include <fstream>
 #include <string>
 #include <sstream>
+#include <assert.h>
 
 #include "config.h"
 
@@ -40,8 +41,7 @@ const float pi=3.1415926535897;
                     std::cout<<stream__.str();}
 #endif
 
-#if 1//def CUDA_IMPLEMENTATION
-  #include <assert.h>
+#ifdef CUDA_CODE
   #define cudaassert(x) {if(x!=cudaSuccess){printvar(x); printvar(cudaGetErrorString(x));}  assert(x==cudaSuccess);}
 #endif
 
@@ -52,7 +52,7 @@ const float pi=3.1415926535897;
 ///   **************************
 
 #ifndef SILENT
-#ifdef CUDA_IMPLEMENTATION
+#ifdef GPU_VERSION
 #if 0
     #define dprintsetup std::stringstream str; str<<blockIdx.x<<" "<<threadIdx.x;
     #define dcheck(x) if(!(x)) printf( "check %s failed.\n", #x );

### 2016-03-27 prm/src/lib/arrayspace/arrayconfigspace.cpp
added forgotten files

@@ -1,0 +1,152 @@
+#include "arrayconfigspace.hpp"
+#include <math.h>
+
+ArrayConfigspace::ArrayConfigspace(const int* array_, int b_, int h_, float minx, float maxx, float miny, float maxy){
+  b=b_; h=h_; n=b*h;
+  mins[0]=minx;
+  mins[1]=miny;
+  maxs[0]=maxx;
+  maxs[1]=maxy;
+  factor[0]=b/(maxs[0]-mins[0]);
+  factor[1]=h/(maxs[1]-mins[1]);
+  dq=1.0/factor[0];
+  if(1.0/factor[1]<dq)dq=1.0/factor[1];
+  array=new int[n];
+  for(int i=0;i<n;++i){
+      array[i]=array_[i];
+  }
+}
+ArrayConfigspace::~ArrayConfigspace(){
+  delete array;
+}
+
+//!initialization function
+int ArrayConfigspace::init(const int ressource_rank, const int ressource_size){
+  return 0;
+}
+//! indicator function of obstacles
+//! q: length d*N, array of structures: q[N*k+i]= k-th component of i-th q-vector
+//! res: length N
+int ArrayConfigspace::indicator(const float* q, int* res, const int N, const int offset){
+  int iy=offset;
+  for(int ix=0;ix<N;++ix,++iy){
+    int x=(int)((q[ix]-mins[0])*factor[0]);
+    int y=(int)((q[iy]-mins[1])*factor[1]);
+    int resix=-1;
+    if(x>=0 && x<b && y>=0 && y<h){
+      int index=b*y+x;
+      resix=array[index];
+    }
+    res[ix]=resix;
+  }
+  return 0;
+}
+//!case N=1
+int ArrayConfigspace::indicator(const float* q){
+  int res=0;
+  indicator(q,&res,1,1);
+  return res;
+}
+
+//! checks if indicator function = 1 somewhere on the line between qs and qe
+//! res is return value
+int ArrayConfigspace::indicator2(const float* qs, const float* qe, int *res, const int N, const int offset){
+  for(int k=0;k<N;++k){
+    float qs_[]={qs[k],qs[k+offset]};
+    float qe_[]={qe[k],qe[k+offset]};
+    if(0 != check_boundaries(&qs_[0]) || 0 != check_boundaries(&qe_[0])){
+      res[k]=2;
+    }else{
[…]

### 2016-03-27 prm/src/lib/arrayspace/arrayconfigspace.hpp

@@ -1,0 +1,54 @@
+#ifndef ARRAYCONFIGSPACE_HPP
+#define ARRAYCONFIGSPACE_HPP
+
+#include <prm/configspace.hpp>
+
+class ArrayConfigspace : public Configspace<2>
+{
+public:
+  ArrayConfigspace(const int* array_, int b_, int h_, float minx, float maxx, float miny, float maxy);
+  ~ArrayConfigspace();
+
+  //!initialization function
+  int init(const int ressource_rank=0, const int ressource_size=1);
+  //! indicator function of obstacles
+  //! q: length d*N, array of structures: q[N*k+i]= k-th component of i-th q-vector
+  //! res: length N
+  int indicator(const float* q, int* res, const int N, const int offset);
+  //!case N=1
+  int indicator(const float* q);
+  //! checks if indicator function = 1 somewhere on the line between (qs[i],qs[i+N]) and (qe[i],qe[i+N])
+  //! res is return value
+  int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
+  int indicator2_async(const float* qs, const float* qe, int *res, const int N, const int offset, int &request);
+  int indicator2_async_wait(int request);
+
+  //! same paircheck as above, but with compressed storage:
+  //! checks pairs: (qs[i],...) ->  (qe(posqe[i]),...) , ...., (qe[posqe[i]+numqe[i]-1],...) for i=0,...,M-1
+  int indicator2(const float* qs, const int M, const float* qe, int *res, const int *posqe, const int *numqe, const int offset);
+  //! structure like indicator function
+  //! returns if lies in boundaries
+  int check_boundaries(const float* q, int* res, int N, const int offset);
+  //! for N=1
+  int check_boundaries(const float* q);
+
+  //! minimal value of qi, i=0...d-1
+  float min(int i);
+  //! maximal value of qi, i=0...d-1
+  float max(int i);
+  //! dq used for indcator2
+  float deltaq(){return dq;}
+  //! dimension d
+  int dim(){return 2;}
+private:
+  float mins[2];
+  float maxs[2];
+  float factor[2];
+  int* array;
+  int b;
+  int h;
+  int n;
+  float dq;
+};
+
+#endif // ARRAYCONFIGSPACE_HPP

### 2016-03-27 prm/src/lib/collision4/cuda_head.h
...

@@ -27,7 +27,7 @@
 #endif
 
 #ifndef GPU_VERSION
-  #pragma message("GPU_VERSION not defined: version without gpu");
+  #warning ("GPU_VERSION not defined: version without gpu");
 #endif
 
 

### 2016-03-27 prm/src/lib/collision4/geo4.h

@@ -8,7 +8,7 @@
 
 
 #ifndef CUDA_CODE
-  #pragma message("NO CUDA_CODE: defining float2 and float4");
+  #warning ("NO CUDA_CODE: defining float2 and float4");
   struct float4{
     float x,y,z,w;
   };

### 2016-03-27 paper/paper.tex
paper angefangen

--- 2015-12-24 description/description.tex
@@
 %\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
-\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the PRM Method}
+\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
 \date{}
@@
 
-\section{Problem Statement and Sequential PRM Method}\label{algintro}
+\section{Introduction}
+\subsection{Problem Statement and the Sequential PRM Method}\label{algintro}
 \newtheorem*{defi}{Definition}
@@
 \begin{equation*}
-	I(\Gamma(s))=0 \text{\ f.a.\ } sÊ\in [0,1].
+	I(\Gamma(s))=0 \text{\ f.a.\ } s \in [0,1].
 \end{equation*}
@@
 
-\section{Application to Robotics}
-
-In the project, it is planned to use this algorithm to create collisionfree trajectories of a robot arm. 
-For a robot with d joint angles \(q_i\) one can calculate straightforwardly all positions \(x_j \in \R^3\) and orientations \(R_j\in \R^{3\times3}\) of the single parts of the robot, \cite{robodyn} describes the exact calculations.
-Therefore this direct kinematic can be interpreted as a (strongly nonlinear) function
-\begin{equation*}
-	f: (q_1, ... , q_d) \mapsto (x_1, R_1, ..., x_l, R_l).
-\end{equation*}
-
-Now, we have given some obstacles with fixed positions \(x_{l+1}, ..., x_m\) and orientations \(R_{l+1}, ..., R_{m}\), which the robot has to avoid during movement.
-It is assumed that all objects are given in a form, that for each two objects it can be determined from the positions and orientations, if there is an intersection.
-For example as polyhedrons.
-From this the indicator function for the obstacles in the configuration space can be defined as
-\begin{equation*}
-	I(q)=
-	\begin{cases}
-		0,	& \text{if no collision for all object pairs\ } (x_j, R_j), (x_k,R_k),\ j,k\in 1, ... , m \\
-		1,	& \text{for at least one collision}
-	\end{cases}
-\end{equation*}
-Now, if we additionally introduce jointlimits and call the set of possible angles \(\Omega\), we see, that the problem of finding feasible trajectories is exactly the pathfinding problem from above.
-
-It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable.
-Hence, they have to be smoothed before the use on a robot.
-This is expectedly not part of the project.
+\subsection{Application to Robotics}
 
@@
 
-\section{Ideas for Parallelization}
+\subsection{Project Overview}
 
@@
 
+\section{Structure of the PRM Solver}
+
+here overview over the components of the programm components:
+
+PRMSolver 1 <-> Configspace object 1 <-> GPU 1
+
+.
+
+.
+
+.
+
+PRMSolver n <-> Configspace object 2 <-> GPU 1
+
+- description of the different versions:
+
+- definitions as pseudocode
+
+version 1:
+
+	- each process generates own nodes -> exchange them, synchronous gpu kernel
+
+version 2:
+
+	- every process generates all nodes (same by exchange of random seed)
+
+	- indicator function splitted over processes, synchronous
+
+version 3:
+
+	- same as version 2 with some pipeline optimizations
+
+version 4:
+
+	- same as version 3 with asynchronous gpu
+
+version 5:
+
+	- each process has 2 workers which are generating their nodes an are using 1 gpu
+
+	- one worker uses the kernel time  of the other to do its work
+
+	- should give use the gpu computation time efficiently
+
+\section{Implementation of an Indicator Function for Robot Arms}
+
+In the project, the probabilistic roadmap algorithm is applied  to create collisionfree trajectories of a robot arm.
+For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
+For this, coordinate frames according to the Denavit Hartenberg convention are used.
+The coordinate frames of the single parts are described by transformation matrices 
+$T_i \in R^{4x4}$, which transform points of the body frame into the world frame.
+They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
+For the exact calculations see \cite{robodyn}.
+
+The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame.
+Applying the corresponding $T_i$, its vertices can be transformed into the world frame.
+Environment obstacles are also defined by convex polytopes, in the world frame.
+
+Now the movement of the robot can be described by a curve $q(t)$ in the state space.
+It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after transformed into the world frame.
+To apply the probabilistic roadmap algorithm, we express this by the indicator function
+
+\begin{equation*}
+	I(q)=
+	\begin{cases}
+		0,	& \text{ for no collision for all given pairs of polytopes in wold frame } \\
+		1,	& \text{ if at least one collision occurs }
+	\end{cases}
+\end{equation*}
+
+It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable.
+Hence, they have to be smoothed before the use on a robot.
+This is not part of the project.
+The task is only to generate feasible trajectories to be passed to an optimizer.
+
+here some overview
+
+- state the principle task: movement of an robot arm through environment with obstacles
+
+- define the resulting state space: $q=(q1,...,qn) \in R^ndof$
+
+- state how the indicator function has to look like in general
+
+- describe the robot:
+
+	- consisting of convex polytopes
+
+ 	- which are located in coordinate frames
+
+	- coordinate frames described by transformation matrices
+
+	- transformation matrices given by state q
+
+- collision tested by the ... algorithm
+
+\subsection{Direct Kinematics}
+
+- coordinate frames
+
+- DH-Parameters
+
+\subsection{2D and 3D Geometry Library}
+
+- vector types: float4, float2
+
+- matrix type: trafo4=float4[4]
+
+- float precision is enough for this application
+
+- everything inline
+
+- hybrid device host functions
+
+- += etc. overloaded, + etc. to make no unnecessary temps
+
+\subsection{Polytopes}
+
+- describe polytope4 structure:
+
+	- vertices stored as float4 array
+
+	- edges stored in crs format
+
+- whole list of polytopes stored compactly in polytope4data on gpu
+
+- maybe describe interface polytope <-> polytope4data <-> polytope4
+
+\subsection{Chung-Wang Collision Algorithm}
+
+- input: polytopes P, Q and trafonsformations Tp, Tq
+
+- output: intersection or not
+
+- describe very shortly how the algorithm works
+
+- implemented as simple device function
+
+\subsection{Robot Kernel}
+
+- the indicator function gets:
+
+	- polytopes in coordinate frames
+
+	- data, which pairs must be tested
+
+	- q
+
+  and computes if for any pair $(i,j)$ the polytopes $P_i, P_j$ collide in the world frame
+
+- implements the indicator function for multiple requests parallely:
+
+- input: lists qstart, qend
+
+- output: res
+
+$res[i] = 0$ if indicator(q) == 0 for all q in $[ qstart[i], qend[i] ]_deltaq$
+
+	!= 0 else
+
+program:
+
+host:
+
+- computes distances between qstart, qend and resulting number of intermediate points/ threads
+
+- qstart, qend, numbers of threads sendt to gpu
+
+gpu kernel:
+
+- each gpu thread computes one intermediate point and its indicator function
+
+- indicator function computation:
+
+	- iterate over all polytope pairs and run the collision algorithm for each
+
+- reduce the indicator functions to res
+
+- send back res to host
+
+--> maybe do this in pseudo code
+
+asynchronous version:
+
+- indicator call: everything untill kernel call
+
+- wait function: res memcopy incl. waiting for kernel
+
 \bibliographystyle{alpha}
@@
 	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press, 2006
-	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl fŸr Angewandte Mechanik, TU MŸnchen, 2014
+	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München, 2014
 \end{thebibliography}

### 2016-03-27 paper/paper.tex
paper

@@
 \usepackage{hyperref}
+
+\usepackage{listings}
+\usepackage{xcolor}
+\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
+
+\lstdefinestyle{customc}{
+  belowcaptionskip=1\baselineskip,
+  breaklines=true,
+  %frame=LR,
+  %xleftmargin=\parindent,
+  language=C,
+  showstringspaces=false,
+  basicstyle=\footnotesize\ttfamily,
+  keywordstyle=\bfseries\color{green!40!black},
+  commentstyle=\itshape\color{purple!40!black},
+  identifierstyle=\color{blue},
+  stringstyle=\color{orange},
+}
+
+\lstdefinestyle{customcpp}{
+	backgroundcolor=\color{lbcolor},
+    	tabsize=4,
+    	language=[GNU]C++,
+        basicstyle=\tiny,%{\scriptsize},
+        upquote=true,
+        %aboveskip={1.5\baselineskip},
+        columns=fixed,
+        showstringspaces=false,
+        extendedchars=false,
+        breaklines=true,
+        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
+        frame=single,
+        numbers=left,
+        showtabs=false,
+        showspaces=false,
+        showstringspaces=false,
+        identifierstyle=\ttfamily,
+        keywordstyle=\color[rgb]{0,0,1},
+        commentstyle=\color[rgb]{0.026,0.112,0.095},
+        stringstyle=\color[rgb]{0.627,0.126,0.941},
+        numberstyle=\color[rgb]{0.205, 0.142, 0.73},
+}
+
+\lstset{escapechar=@,style=customcpp}
 
@@
 For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
-For this, coordinate frames according to the Denavit Hartenberg convention are used.
+For this, coordinate frames according to the Denavit-Hartenberg convention are used.
 The coordinate frames of the single parts are described by transformation matrices 
@@
 They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
-For the exact calculations see \cite{robodyn}.
+%For the exact calculations see \cite{robodyn}.
 
@@
 
-here some overview
-
-- state the principle task: movement of an robot arm through environment with obstacles
-
-- define the resulting state space: $q=(q1,...,qn) \in R^ndof$
-
-- state how the indicator function has to look like in general
-
-- describe the robot:
-
-	- consisting of convex polytopes
-
- 	- which are located in coordinate frames
-
-	- coordinate frames described by transformation matrices
-
-	- transformation matrices given by state q
-
-- collision tested by the ... algorithm
+The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^N$, the indicator function is evaluated for each state by one CUDA thread.
+Therefore, the kinematics and collision algorithm have been implemented as device code for the use in the kernel.
 
@@
 
-- coordinate frames
-
-- DH-Parameters
+The Denavit-Hartenberg (DH) convention is a method to describe the geometry of an robots joints, 
+that is, how the transformations between the robots coordinate frames look like in dependancy of its joint values $q_1,...,q_d$.
+For this it is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
+or prismatic, a translation along an axis.
+The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by the transformation matrix $T_i$ into the world frame.
+For an robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i \in \R$, $i=1,...,d$, 
+the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies now is defined as
+\begin{equation}
+	T_{i-1,i} = R_z(\theta_i) T_z(d_i) R_x(\alpha_{i-1}) R_x(a_{i-1})
+\end{equation}
+where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
+By an proper choice of the body frames, every robot with rotational or prismatic joint can be described by DH parameters.
 
@@
 
-- vector types: float4, float2
+For the CUDA implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes float2 for 2D vectors and float4 for 3D vectors.
+To guarantee aligned access on the GPU, float4 is used instead of float3.
+Furthermore, as the geometric calculations made here are numerically not very critical, the usage of double precision has turned out to be not necessary.
+The library contains hybrid host and device inline functions for common vector operations such as
+\begin{lstlisting}
+__host__ __device__ inline float4& operator += (float4& u, const float4& v);
+__host__ __device__ inline float4& operator *=  (float4& u, const float f);
+__host__ __device__ inline float4& add(const float4& u, const float4& v, float4& w);
+\end{lstlisting}
+Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
 
-- matrix type: trafo4=float4[4]
-
-- float precision is enough for this application
-
-- everything inline
-
-- hybrid device host functions
-
-- += etc. overloaded, + etc. to make no unnecessary temps
+Transformation matrices are implemented float4 arrays with inline functions for common matrix operations. Again operators with temporaries avoided.
+\begin{lstlisting}
+class trafo4{
+  public:
+    float4 col[4];
+    __host__ __device__ trafo4(){}
+    __host__ __device__ trafo4(float a, float alpha, float q, float d);
+    __host__ __device__ inline float4& apply(float4& u) const;
+    __host__ __device__ inline float4& apply(const float4& u, float4& Tu) const;
+    .
+    .
+    .
+}
+\end{lstlisting}
 
@@
 
-- describe polytope4 structure:
-
-	- vertices stored as float4 array
-
-	- edges stored in crs format
+To calculate the intersection between two polytopes, the Chung-Wang algorithm only needs their vertices 
+and the data which vertices are connected by an edge. 
+For this the adjacency matrix of the edge graph is stored compactly in the crs format.
+\begin{lstlisting}
+struct polytope4{
+    float4* vertices;	//length n
+    int n;
+    //!edges in crs format
+    int* dsp;	//length n
+    int* cnt;	//length n
+    int* dest;	//length m
+    int m;
+};
+\end{lstlisting}
+It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
 
@@
 \subsection{Chung-Wang Collision Algorithm}
+
+\begin{lstlisting}
+__host__ __device__ int seperating_vector_algorithm(const polytope4& P, const polytope4& Q, 
+													const trafo4& tp, const trafo4& tq);
+\end{lstlisting}
 

### 2016-03-27 prm/src/lib/collision4/polytope4.h

@@ -13,7 +13,7 @@ namespace collision4{
 
   struct polytope4{
     float4* vertices;
-    int n; //Reihenfolge?
+    int n;
     //!edges saved in crs format
     int* dsp;
     int* cnt;

### 2016-03-28 paper/paper.tex
robot kernel chapter...

@@
 \usepackage{hyperref}
+
+\usepackage[ruled]{algorithm2e}
 
@@
 
-The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^N$, the indicator function is evaluated for each state by one CUDA thread.
+The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
 Therefore, the kinematics and collision algorithm have been implemented as device code for the use in the kernel.
@@
 \subsection{Chung-Wang Collision Algorithm}
-
+The indicator function is implemented in a compressed way
 \begin{lstlisting}
@@
 
-- the indicator function gets:
+\begin{lstlisting}
+int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
+\end{lstlisting}
+It gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$.
+This are the nodes passed by the PRMSolver.
+The indicator function checks now for each pair, if on the line in between the indicator function is always 0.
+\begin{equation}
+	res[j]=\begin{cases}
+		0	&\text{if $I(q)=0$ f.a. $q \in \left[qs[j], qe[j]\right]_{\Delta q}$} \\
+		1	&else
+	\end{cases}
+\end{equation}
+The border points are stored in structures of arrays.
+\begin{equation}
+\begin{aligned}
+	q_{start}^j &= \left(qs[j],qs[j+\text{offset}],...,qs[j+(d-1)*\text{offset}]\right)^T
+	q_{end}^j &= \left(qe[j],qe[j+\text{offset}],...,qe[j+(d-1)*\text{offset}]\right)^T
+\end{aligned}
+\end{equation}
 
-	- polytopes in coordinate frames
+\begin{algorithm}
+	\KwData{qs, qe, N, offset}
+	\KwResult{res}
+	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
+	testnum[j] = $d_j/{\Delta q}$: number of threads for each line \;
+	testpos[j] = thread displacements \;
+	reset res \;
+	call kernel \;
+	\caption{Host indicator function}
+\end{algorithm}
 
-	- data, which pairs must be tested
-
-	- q
-
-  and computes if for any pair $(i,j)$ the polytopes $P_i, P_j$ collide in the world frame
-
-- implements the indicator function for multiple requests parallely:
-
-- input: lists qstart, qend
-
-- output: res
-
-$res[i] = 0$ if indicator(q) == 0 for all q in $[ qstart[i], qend[i] ]_deltaq$
-
-	!= 0 else
+\begin{algorithm}
+	\KwData{arrays qs, qe, testpos, testnum and threadindex}
+	\KwResult{res}
+	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
+	$c = (threadindex - testpos[j]) / (testnum[k]-1) $ \;
+	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
+	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
+	$res_{tmp} = 0; $ \;
+	\For{all registered pairs pairs of polytopes $P, Q$}{
+		$res_{tmp}=seperating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
+		\If{$res_{tmp} \neq 0$}{
+			$res[j]=res_{tmp}$ \;
+			break \;
+		}
+	}
+	\caption{Robot kernel: code for one thread}
+\end{algorithm}
 
@@
 
+In order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
+Here the function is split in everything until the kernel launch and a wait function, which waits until the kernel finishes and recieves the result data.
+The implementation allows also overlapping of multiple requests, for example two calls of the launch function and after this the two waiting calls.
+With this it is possible the full computation time of the GPU.
+
+\begin{lstlisting}
+int indicator2_async(const float* qs, const float* qe, int *res, 
+					 const int N, const int offset, int &request);
+int indicator2_async_wait(int request);
+\end{lstlisting}
+
 \bibliographystyle{alpha}

### 2016-03-28 prm/src/lib/robotspace/robotconfigspace.cpp

@@ -376,7 +376,7 @@ void kernel_indicator2(const Robot<ndof>* robot,
     //! calculate q (convex combination)
     float q[ndof];
     float c1,c2;
-    if(testnum[k]>0){
+    if(testnum[k]>0){ //TODO: 1 statt 0 ?
       c1=(float)(i-testpos[k])/(float)(testnum[k]-1);
       c2=1.0-c1;
     }else{

### 2016-03-28 paper/paper.tex
paper...

@@
 
-\subsection{Direct Kinematics}
+\subsection{Direct Kinematics}\label{dk}
 
@@
 \subsection{Chung-Wang Collision Algorithm}
-The indicator function is implemented in a compressed way
+
+To check if two polytopes collide an algorithm by Chung and Wang is used here.
+It uses the separating vector theorem and tries to find a seperating vector by iterative search.
+Furthermore, a subalgorithm determines in each iteration from the candidate history, if it is still theoretically possible to find a seperating vector.
+If not, it is returned a collision.
+If on the other side a seperating vector is found, the algorithm can return no collision.
+For detailed informations see \cite{chung} and \cite{ericson}, pp. 410-412.
+The method is implemented as device code for the use on GPU kernels.
 \begin{lstlisting}
@@
 \end{lstlisting}
-
-- input: polytopes P, Q and trafonsformations Tp, Tq
-
-- output: intersection or not
-
-- describe very shortly how the algorithm works
-
-- implemented as simple device function
+In our application, the polytopes are mostly parts of the robot, which are given in a body frame and have to be transformed into world frame according to section \ref{dk} first.
+However, from the structure of the Chung-Wang algorithm drops out that mostly only a small part of the polytopes vertices are needed for the computations.
+Therefore it would be rather time wasting to transform all vertices before passing to the algorithm.
+To avoid this, all polytopes are passed in their body frames together with their transformations and the algorithm transforms vertices at the time they occur in calculations.
 
@@
 
+The robot configuration space is implemented in the form
 \begin{lstlisting}
+template<int ndof>
+class RobotConfigspace : public Configspace<ndof>
+{
+public:
+  RobotConfigspace(const Robot<ndof>* robot_,
+                   const polytope *polys_,
+                   const int* sys_,
+                   const int N_,
+                   const int *from_, const int *to_,
+                   const int M_,
+                   const float* mins_, const float* maxs_, const float dq_,
+                   const int nbuf_);
+  int init(const int ressource_rank=0, const int ressource_size=1);
   int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
+  .
+  .
+  .
+private:
+  const Robot<ndof>* robot; //host
+  Robot<ndof>* robotdev;    //GPU
+  collision4::polytope4data* polydata;            //data on host
+  collision4::polytope4data* polydatadev;         //pointer to GPU obj
+  .
+  .
+  .
+}
 \end{lstlisting}
-It gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$.
-This are the nodes passed by the PRMSolver.
-The indicator function checks now for each pair, if on the line in between the indicator function is always 0.
+\begin{lstlisting}
+__global__ void kernel_indicator2(const Robot<ndof>* robot,
+                                  const collision4::polytope4data* polydata,
+                                  const float* qs, int offsets,
+                                  const float* qe, int offsete,
+                                  int* res,
+                                  const int* testpos, const int* testnum,
+                                  int N, int numthreads);
+\end{lstlisting}
+where the constant geometry and robot data is passed once at the beginning and copied to the GPU by an init function.
+The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if on the line in between the indicator function is always 0.
 \begin{equation}
@@
 \begin{aligned}
-	q_{start}^j &= \left(qs[j],qs[j+\text{offset}],...,qs[j+(d-1)*\text{offset}]\right)^T
+	q_{start}^j &= \left(qs[j],qs[j+\text{offset}],...,qs[j+(d-1)*\text{offset}]\right)^T\\
 	q_{end}^j &= \left(qe[j],qe[j+\text{offset}],...,qe[j+(d-1)*\text{offset}]\right)^T
@@
 \end{equation}
+
+As the PRMSolver only needs to know, which node points can be connected by a line, and not, which points on the line exactly are free, this version delivers exactly the minimal necessary information.
+Furthermore it saves CPU and communication time, because the intermediate points on the lines can be directly computed by the GPU threads and do not need to be exchanged.
+
+Algorithms 1 and 2 two show in pseudo code, how the work is split between host and device.
 
@@
 
-program:
-
-host:
-
-- computes distances between qstart, qend and resulting number of intermediate points/ threads
-
-- qstart, qend, numbers of threads sendt to gpu
-
-gpu kernel:
-
-- each gpu thread computes one intermediate point and its indicator function
-
-- indicator function computation:
-
-	- iterate over all polytope pairs and run the collision algorithm for each
-
-- reduce the indicator functions to res
-
-- send back res to host
-
---> maybe do this in pseudo code
-
-asynchronous version:
-
-- indicator call: everything untill kernel call
-
-- wait function: res memcopy incl. waiting for kernel
-
 In order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
 Here the function is split in everything until the kernel launch and a wait function, which waits until the kernel finishes and recieves the result data.
-The implementation allows also overlapping of multiple requests, for example two calls of the launch function and after this the two waiting calls.
-With this it is possible the full computation time of the GPU.
 
@@
 
+The implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corrsponding waiting calls. 
+With this it is possible to use the full computation time of the GPU.
+
 \bibliographystyle{alpha}
@@
 	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München, 2014
+	\bibitem{chung} K.
+Chung, Wenping Wang, Quick Collision Detection of Polytopes in Virtual Environments, Proceedings of the ACM Symposium on Virtual Reality Software and Technology (VRST 96), Mark Green (ed.), 1996
+	\bibitem{ericson} C. Ericson, Real-Time Collision Detection, Elsevier, 2005
 \end{thebibliography}

### 2016-03-28 prm/src/lib/robotspace/robotconfigspace.cpp

@@ -405,7 +405,7 @@ void kernel_indicator2(const Robot<ndof>* robot,
           int k1=dest[l];
           collision4::polytope4 poly1;
           polydata->get_polytope(poly1, k1);
-          int result=collision4::seperating_vector_algorithm(poly0,poly1,kin.trafos[polydata->sys[k0]],kin.trafos[polydata->sys[k1]]);
+          int result=collision4::seperating_vector_algorithm(poly0,poly1,kin.trafos[polydata->sys[k0]],kin.trafos[polydata->sys[k1]]); //-> TODO: seperate environment? code sparen?
           if(result!=0){
             resext=result;
           }

### 2016-03-28 paper/paper.tex
paper...

@@
 
-The abstract problem is to find a path on a d-dimensional map, which avoids intersecting obstacles. The obstacles are  given by an indicator function
+The abstract problem is to find a path on a d-dimensional space, which avoids intersecting obstacles. The obstacles are  given by an indicator function
 \begin{equation*}
@@
     \includegraphics[width=0.8\textwidth]{configspacetrafo.png}
-    \caption{Example for the transformation of obstacles into configspace. 
-    		The scattered areas on the right are the sets of points with I(q)=1, \cite{prmlec}.}
+    \caption{Example for an indicator function in robotics from \cite{prmlec}. 
+    		The real obstacles on the left are transformed 
+		into occupied areas of the configuration space on the right.}
     \label{fig:awesome_image}
 \end{figure}
+
+In the project, the PRM algorithm is used to find trajectories for a robot arm, which avoid self collisions and collisions with obstacles in the environment.
+Such trajectories can be expressed as functions describing the joint variables changing over time, that means curves in the configuration space with dimension equal to the number of degrees of freedom of the robot.
+With association the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form stated in the previous chapter.
+The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given set of joints, if a collision occurs.
 

### 2016-03-29 paper/paper.tex
paper...

@@
 
+% PSEUDOCODE
+
 \usepackage[ruled]{algorithm2e}
+
+% C++ CODE
 
@@
 \lstset{escapechar=@,style=customcpp}
+
+% TIKZ DIAGRAMS
+
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+\usetikzlibrary{positioning}
+
+\tikzset{
+    block/.style={
+           rectangle,
+           rounded corners,
+           draw=black, very thick,
+           minimum height=2em,
+           inner sep=2pt,
+           text centered,
+           },
+}
+
+\usepackage{tikz-cd}
+\tikzset{
+  shift left/.style ={commutative diagrams/shift left={#1}},
+  shift right/.style={commutative diagrams/shift right={#1}}
+}
 
@@
 Such trajectories can be expressed as functions describing the joint variables changing over time, that means curves in the configuration space with dimension equal to the number of degrees of freedom of the robot.
-With association the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form stated in the previous chapter.
-The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given set of joints, if a collision occurs.
+With this association, the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form stated in the previous chapter.
+The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given set of joints, if the configuration is feasible or a collision occurs.
 
@@
 
-Like in the stated robotics example, in many cases the most costly operation of the PRM algorithm is the evaluation of the indicator function \(I(q)\), which however can be done independently for every new \(q\).
-\begin{itemize}
-\item 
-Therefore in a first step, it is planned to implement \(I(q)\) parallely in a vectorized way, i.e.
-as a function \((q_1, ...
-, q_P) \mapsto (I(q_1),...,I(q_P))\).
-For example with help of the GPU.
-In the robotics example this means, the direct kinematic and collision calculation is made parallely for \(P\) different joint angle vectors.
-\item
+Like in the stated robotics example, in many cases the most costly operation of the PRM algorithm are the many evaluations of the indicator function \(I(q)\), 
+which however can be done independently for every new \(q\).
+Therefore $I(q)$ is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every thread computes one evaluation.
+In the robotics case, this means, the direct kinematic and collision calculation are made on the GPU parallely for different joint angle configurations.
+Although this will the only major application in the project, the PRM solver and the configuration space are implemented seperated from each other and connected by an interface to be free to simply add other applications.
+As an example and for testing, additionally a simple 2D configuration space is provided,
+whose indicator function is read from a black-and-white image.
+
+\ \\ 
+\begin{tikzpicture}[auto]%node distance=0.5cm and 1.8cm]
+	\node[block] (prm) {
+		\begin{tabular}{c}
+		\textbf{PRMSolver}\\
+		generate nodes\\
+		store graph
+		\end{tabular}
+	};
+	\node [block, right=of prm] (config) {
+		\begin{tabular}{c}
+		\textbf{Configpace}\\
+		indicator fct. \\
+		interface
+		\end{tabular}
+	};
+	\node [block, right=of config] (robo) {
+		\begin{tabular}{c}
+		\textbf{RobotConfigspace}\\
+		robot kinematics\\
+		collision detection
+		\end{tabular}
+	};
+	\node [block, right=of robo] (gpu) {\textbf{GPU}};
+
+	\path[-stealth,thick, shift left=1.0ex]	(prm) edge (config) ;
+	\path[-stealth,thick, shift left=1.0ex]	(config) edge  (prm) ;
+	\path[-stealth,thick, shift left=1.0ex]	(config) edge (robo);
+	\path[-stealth,thick, shift left=1.0ex]	(robo) edge  (config);
+	\path[-,thick]					(robo) edge (gpu);
+\end{tikzpicture}
+\ \\
+
 Then in a first level of parallelization, this function can be applied to the \(q_i \in [q, \tilde q]_h \) to check, if \(q\) and \(\tilde q\) are connected.
-\item
+
 Furthermore, a second level of parallelization could be achieved by sampling and connecting several new points \(\tilde q_i\) parallely.
@@
 Therefore it has to be evaluated, if a parallel implementation stays efficient here.
-\item
+
 If there remains time, a next idea could be to execute the algorithm on different discretization levels \(h\).
@@
 , q_n^1\), etc.
-\end{itemize}
 
@@
 They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
-%For the exact calculations see \cite{robodyn}.
 
@@
 By an proper choice of the body frames, every robot with rotational or prismatic joint can be described by DH parameters.
+For the exact calculations, see \cite{robodyn}.
 

### 2016-03-29 paper/paper.tex
paper...

@@
 %\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
-\title{\vspace{-1.3cm} \textmd{\normalsize{Project description}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
+\title{\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
 \date{}
@@
 
-Then in a first level of parallelization, this function can be applied to the \(q_i \in [q, \tilde q]_h \) to check, if \(q\) and \(\tilde q\) are connected.
+While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
+a second level can be achieved by sampling and connecting several new nodes parallely. 
+This is more difficult, because all new nodes have to be connected 
+and the graph data has to be updated on all processes. 
+For this, different approaches have been implemented, which are described in chapter \ref{solver}.
 
-Furthermore, a second level of parallelization could be achieved by sampling and connecting several new points \(\tilde q_i\) parallely.
-This is expected to be more difficult, because all new nodes have to be connected to the old graph and to each other.
-Therefore it has to be evaluated, if a parallel implementation stays efficient here.
+Before that, chapter \ref{indicator} gives more details about the robotics indicator function.
 
-If there remains time, a next idea could be to execute the algorithm on different discretization levels \(h\).
-This means for example to compute a way \(q_1^0, ..., q_n^0\) which is connected with a coarse stepsize \(h_0\), then to decrease the stepsize to \(0<h_1<h_0\) and take the path nodes \(q_i^0\) as starting nodes for the next execution of the algorithm, which gives then an \(h_1\)-connected way \(q_1^1, ...
-, q_n^1\), etc.
-
-\section{Structure of the PRM Solver}
-
-here overview over the components of the programm components:
-
-PRMSolver 1 <-> Configspace object 1 <-> GPU 1
-
-.
-
-.
-
-.
-
-PRMSolver n <-> Configspace object 2 <-> GPU 1
-
-- description of the different versions:
-
-- definitions as pseudocode
-
-version 1:
-
-	- each process generates own nodes -> exchange them, synchronous gpu kernel
-
-version 2:
-
-	- every process generates all nodes (same by exchange of random seed)
-
-	- indicator function splitted over processes, synchronous
-
-version 3:
-
-	- same as version 2 with some pipeline optimizations
-
-version 4:
-
-	- same as version 3 with asynchronous gpu
-
-version 5:
-
-	- each process has 2 workers which are generating their nodes an are using 1 gpu
-
-	- one worker uses the kernel time  of the other to do its work
-
-	- should give use the gpu computation time efficiently
-
-\section{Implementation of an Indicator Function for Robot Arms}
+\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}.
 
@@
 The coordinate frames of the single parts are described by transformation matrices 
-$T_i \in R^{4x4}$, which transform points of the body frame into the world frame.
+$T_i \in R^{4x4}$, which transform points from the body frame frame into the world frame.
 They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
@@
 The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame.
-Applying the corresponding $T_i$, its vertices can be transformed into the world frame.
+Applying the corresponding transformation $T_i$, its vertices can be transformed into the world frame.
 Environment obstacles are also defined by convex polytopes, in the world frame.
 
-Now the movement of the robot can be described by a curve $q(t)$ in the state space.
-It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after transformed into the world frame.
+Now the movement of the robot can be described by a curve $q(t) \in \R^d$ in the configuration space.
+It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
 To apply the probabilistic roadmap algorithm, we express this by the indicator function
@@
 	\begin{cases}
-		0,	& \text{ for no collision for all given pairs of polytopes in wold frame } \\
+		0,	& \text{ for no collision for all given pairs of polytopes } \\
 		1,	& \text{ if at least one collision occurs }
@@
 The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
-Therefore, the kinematics and collision algorithm have been implemented as device code for the use in the kernel.
+Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.
 
@@
 or prismatic, a translation along an axis.
-The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by the transformation matrix $T_i$ into the world frame.
-For an robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i \in \R$, $i=1,...,d$, 
-the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies now is defined as
+The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by a transformation matrix $T_i$ into the world frame.
+For a robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
+it is now defined, that the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies are
 \begin{equation}
@@
 where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
-By an proper choice of the body frames, every robot with rotational or prismatic joint can be described by DH parameters.
+From this follows, that from given DH parameters all transformations are determined.
+The other way round, one can show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by a set of DH parameters.
 For the exact calculations, see \cite{robodyn}.
@@
 
-Transformation matrices are implemented float4 arrays with inline functions for common matrix operations. Again operators with temporaries avoided.
+Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
 \begin{lstlisting}
@@
 
+\section{Structure of the PRM Solver}\label{solver}
+
+here overview over the components of the programm components:
+
+PRMSolver 1 <-> Configspace object 1 <-> GPU 1
+
+.
+
+.
+
+.
+
+PRMSolver n <-> Configspace object 2 <-> GPU 1
+
+- description of the different versions:
+
+- definitions as pseudocode
+
+version 1:
+
+	- each process generates own nodes -> exchange them, synchronous gpu kernel
+
+version 2:
+
+	- every process generates all nodes (same by exchange of random seed)
+
+	- indicator function splitted over processes, synchronous
+
+version 3:
+
+	- same as version 2 with some pipeline optimizations
+
+version 4:
+
+	- same as version 3 with asynchronous gpu
+
+version 5:
+
+	- each process has 2 workers which are generating their nodes an are using 1 gpu
+
+	- one worker uses the kernel time  of the other to do its work
+
+	- should give use the gpu computation time efficiently
+
+\section{Results}
+
+- timings of different versions at same conditions
+
+\section{Conclusion}
+
 \bibliographystyle{alpha}

### 2016-03-29 paper/paper.tex
diagramme...

@@
 \usepackage{hyperref}
+
+\usepackage{graphicx}
 
@@
            draw=black, very thick,
+           minimum height=2em,
+           inner sep=2pt,
+           text centered,
+           },
+     block2/.style={
+           %rectangle,
+           %rounded corners,
+           fill=black!20,
+           %draw=black, very thick,
+           minimum height=2em,
+           inner sep=2pt,
+           text centered,
+           },
+     frameless/.style={
+           %rectangle,
+           %rounded corners,
+           %draw=black, very thick,
            minimum height=2em,
@@
 
-\ \\ 
+%\ \\
+\begin{center} 
 \begin{tikzpicture}[auto]%node distance=0.5cm and 1.8cm]
@@
 \end{tikzpicture}
-\ \\
+\end{center}
+%\ \\
 
@@
 
-\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}.
+\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}
 
@@
 
-To calculate the intersection between two polytopes, the Chung-Wang algorithm only needs their vertices 
+To calculate the intersection between two convex polytopes, the Chung-Wang algorithm only needs their vertices 
 and the data which vertices are connected by an edge. 
@@
 It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
-
-- whole list of polytopes stored compactly in polytope4data on gpu
-
-- maybe describe interface polytope <-> polytope4data <-> polytope4
+The polytopes used in this projects are given by the user as convex hull of a list of vertices.
+However, to determine the edge graph only from the vertices is a nontrivial task, which is done by Matlab here using provided high level functions.
 
@@
 
+\subsection{Storage of the Graph}
+
+\subsection{Parallelization Models}
+
 here overview over the components of the programm components:
-
-PRMSolver 1 <-> Configspace object 1 <-> GPU 1
-
-.
-
-.
-
-.
-
-PRMSolver n <-> Configspace object 2 <-> GPU 1
 
@@
 
-- definitions as pseudocode
+- mpi
 
-version 1:
+- every process has its own GPU
 
-	- each process generates own nodes -> exchange them, synchronous gpu kernel
+\subsubsection*{Version 1}
 
-version 2:
+In the first approach, that was evaluated, every process generates independantly from each other new nodes.
+Then the possible neighbors are determined and the connections, which have to be tested.
+The indicator function is called and returnes the edges.
+After this the graph must be synchronized on every processes, which leads to an alltoall communication.
+As in this graph building does not need the edge data, it is sufficient to exchange only the nodes in this phase.
+
+	\begin{center}
+\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
+	\node[block, minimum width=3cm] (prm) {
+		\begin{tabular}{c}
+		\textbf{PRMSolver 1}
+		\end{tabular}
+	};
+
+	\node[frameless, below=of prm, minimum width=3cm] (space) {$\vdots$};
+
+	\node[block, below=of space, minimum width=3cm] (prm2) {
+		\begin{tabular}{c}
+		\textbf{PRMSolver n}
+		\end{tabular}
+	};
+
+	\node [block, right=of prm] (config) {
+		\begin{tabular}{c}
+		\textbf{Configpace 1}
+		\end{tabular}
+	};
+	\node[frameless, below=of config] (space2) {$\vdots$};
+	\node [block, below=of space2] (config2) {
+		\begin{tabular}{c}
+		\textbf{Configpace n}
+		\end{tabular}
+	};
+
+	\node [block, right=of config] (gpu) {\textbf{GPU 1}};
+	\node [block, right=of config2] (gpu2) {\textbf{GPU n}};
+
+	\path[-latex',thick, shift left=1.0ex]	(prm) edge (config) 
+								(config) edge (prm);
+	\path[-,thick]					(config) edge (gpu);
+
+	\path[-latex',thick, shift left=1.0ex]	(prm2) edge (config2) 
+								(config2) edge (prm2) ;
+	\path[-,thick]					(config2) edge (gpu2);
+
+	\path[-latex',thick, shift left=1.0ex]	(prm) edge (space)
+								(space) edge (prm)
+								(prm2) edge (space)
+								(space) edge (prm2) ;
+
+\end{tikzpicture}
+\end{center}
+
+\subsubsection*{Version 2}
+The second approach
 
@@
 
-version 3:
+	\begin{center}
+\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
+	\node[block, minimum width=3cm] (prm) {
+		\begin{tabular}{c}
+		\textbf{PRMSolver 1}
+		\end{tabular}
+	};
+
+	\node[frameless, below=of prm, minimum width=3cm] (space) {$\vdots$};
+
+	\node[block, below=of space, minimum width=3cm] (prm2) {
+		\begin{tabular}{c}
+		\textbf{PRMSolver n}
+		\end{tabular}
+	};
+
+	\node [block, right=of prm] (config) {
+		\begin{tabular}{c}
+		\textbf{Configpace 1}
+		\end{tabular}
+	};
+	\node[frameless, below=of config] (space2) {$\vdots$};
+	\node [block, below=of space2] (config2) {
+		\begin{tabular}{c}
+		\textbf{Configpace n}
+		\end{tabular}
+	};
+
+	\node [block, right=of config] (gpu) {\textbf{GPU 1}};
+	\node [block, right=of config2] (gpu2) {\textbf{GPU n}};
+
+	\path[-latex',thick, shift left=1.5ex]	(prm) edge (config) ;
+	\path[-,thick]					(config) edge (gpu);
+
+	\path[-latex',thick, shift right=1.5ex]	(prm2) edge (config2) ;
+	\path[-,thick]					(config2) edge (gpu2);
+
+	\path[-latex',thick]				(config.west) edge  (prm.east) 
+								(config.west) edge (space.east)
+								(config.west) edge (prm2.east)
+								(config2.west) edge  (prm.east) 
+								(config2.west) edge (space.east)
+								(config2.west) edge (prm2.east);
+
+\end{tikzpicture}
+\end{center}
+
+\subsubsection*{Version 3}
 
@@
 
+Stateflow for each PRMSolver instance:
+
+\begin{center}
+\begin{tikzpicture}
+	\tikzset{mywidth/.style={minimum width=2.5cm},};
+	\node[block, mywidth] (l1) {\textbf{Worker 1}};
+	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
+	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
+										generate\\ nodes, \\
+										start kernel
+									\end{tabular}};
+	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
+	\node[frameless, mywidth, below=of l6] (l7) {};
+	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
+										recieve\\ result, \\
+										store nodes \\ 
+										and edges
+									\end{tabular}};
+	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};
+
+	\node[block, mywidth, right=of l1] (r1) {\textbf{Worker 2}};
+	\node[frameless, mywidth, right=of l2, minimum height=1.5cm] (r2) {$\vdots$};
+	\node[frameless, mywidth, right=of l3] (r3) {};
+	\node[block, mywidth, right=of l6] (r6)	{\begin{tabular}{c}
+										recieve\\ result, \\
+										store nodes \\
+										and edges
+									\end{tabular}};
+	\node[block, mywidth, right=of l7] (r7) 	{\begin{tabular}{c}
+										generate\\ nodes, \\
+										start kernel
+									\end{tabular}};
+	\node[frameless, mywidth, right=of l12] (r12) {$\vdots$};
+
+	\node[frameless, left=of l1] (g1) {};
+	\node[block, below=of g1, minimum height=14cm] (graph) {\rotatebox{90}{\textbf{Graph}}};
+
+	%gpu column
+	\tikzset{gpuwidth/.style={minimum width=0.8cm},};
+
+	\node[block, right=2cm of r1, minimum width=2.5cm] {\textbf{GPU}};
+	\node[frameless, right=2cm of r2, gpuwidth] (topa) {$\vdots$};
+	\node[frameless, right=of topa, gpuwidth] (topb) {$\vdots$};
+	\node[block2, below=0.7cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
+	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=4.2cm] (stream1b) {\rotatebox{90}{kernel}};
+	\node[frameless, below=0.3cm of stream1b, gpuwidth] (bota) {$\vdots$};
+
+	\node[block2, below=2.2cm of topb, gpuwidth, minimum height=7.6cm] (stream2) {\rotatebox{90}{kernel}};
+	\node[frameless, right=of bota, gpuwidth] (botb) {$\vdots$};
+
+	%EDGES
+
+	\path[-latex',thick]	(graph.east |- l3.west) edge (l3.west)
+					(l11.west) edge (graph.east |- l11.west);
+	\path[-latex',thick]	(graph.east |- r7.west) edge (r7.west)
+					(r6.west) edge (graph.east |- r6.west);
+	\path[dotted,thick]	(l2) edge (l3)
+					(l3) edge (l11)
+					(l11) edge (l12)
+					(r2) edge (r6)
+					(r6) edge (r7)
+					(r7) edge (r12);
+
+	\path[-latex',thick]	(r6.east -| stream1b.west) edge (r6.east);
+	\path[-latex',thick]	(r7.east) edge (r7.east -| stream1b.west);
+
+	\path[-,thick]		(l3.east) edge (l3.east -| stream1a.west);
+	\path[-latex',thick]	(stream1a.east |- l3.east) edge (l3.east -| stream2.west);
+	\path[latex'-,thick]		(l11.east) edge (l11.east -| stream1b.west);
+	\path[-,thick]		(stream1b.east |- l11.east) edge (l11.east -| stream2.west);
+
+\end{tikzpicture}
+\end{center}
+
 \section{Results}
@@
 - timings of different versions at same conditions
+
+- difficult to test because
+	- not deterministic
+	- dependant on density of nodes: how many neighbours have to be tested per connection
 

### 2016-03-29 paper/paper.tex
paper...

@@
 \end{defi}
-\begin{alg1} (Probabilistic Roadmap in general)\\
-	Input: \(q_s, q_e\in \Omega, h>0, d_0>0\)	\\
-	Output: \(q_s=q_1, ... ,q_n=q_e\in \Omega\), 
-		such that all \(q_i\) and \(q_{i+1}\) are connected with stepsize \(h\)
-	\begin{itemize}
-		\item Initialize graph \(G=(V,E)\) with \(V=\{q_s, q_e\}\), 
-			if \(q_s\) and \(q_e\) are connected, \(E \leftarrow\{ q_s, q_e\}\)
-		\item While there exists no path in G between \(q_s\) and \(q_e\) do
-		\begin{itemize}
-			\item Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\)
-			\item \(V \leftarrow{q}\)
-			\item For all nodes \(\tilde{q} \in V\) which are near enough (\( \norm{q- \tilde q} \le d_0 \)) and
-				connected with \(q\) with stepsize h, 
+
+\begin{algorithm} 
+	\caption{Probabilistic Roadmap in general}
+	\KwData{ \(q_s, q_e\in \Omega, h>0, d_0>0\)}
+	\KwResult{ \(q_s=q_1, ... ,q_n=q_e\in \Omega\), 
+		such that all \(q_i\) and \(q_{i+1}\) are connected with stepsize \(h\)}
+	Initialize graph \(G=(V,E)\) with \(V=\{q_s, q_e\}\) \\
+	If \(q_s\) and \(q_e\) are connected, \(E \leftarrow\{ q_s, q_e\}\) \\
+	\While{ there exists no path in G between \(q_s\) and \(q_e\)}{
+		Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\) \\
+		\For{ all nodes \(\tilde{q} \in V\), with \( \norm{q- \tilde q} \le d_0 \) }{
+			\If { $q$ and $\tilde q$ connected with stepsize h} {
 				\(E \leftarrow \{q, \tilde{q}\}\)
-		\end{itemize}
-		\item Determine the shortest path \(q_1, ... , q_n\in V\) on \(G\) from \(q_s\) to \(q_e\).
-	\end{itemize}
-\end{alg1}
+			}
+		}
+		\(V \leftarrow{q}\), if at least one edge was inserted \\
+	}
+	Determine the shortest path \(q_1, ... , q_n\in V\) on \(G\) from \(q_s\) to \(q_e\).
+\end{algorithm}
 
@@
 
-%\ \\
 \begin{center} 
@@
 \end{center}
-%\ \\
 
@@
     int n;
-    //!edges in crs format
+    //edges
     int* dsp;	//length n
@@
 
+- introduction
+
+- dijkstra at the end makes almost no contribution to the whole time
+
+	- can be parallelized, but this is again complicated
+
+- focus on building the graph
+
 \subsection{Storage of the Graph}
+
+- describe
+
+- compare different models...?
 
@@
 
-here overview over the components of the programm components:
-
-- description of the different versions:
-
-- mpi
-
-- every process has its own GPU
+The parallelization is done with MPI. All processes store their own versions of the graph, which are hold up to date between each other.
+Furthermore, it is assumed that every process has its own GPU.
+The work of the single MPI processes could be parallelized further with OpenMP, but this has not been done here.
+Instead only different approaches to distribute the work on several GPUs have been investigated.
 
@@
 
-In the first approach, that was evaluated, every process generates independantly from each other new nodes.
-Then the possible neighbors are determined and the connections, which have to be tested.
+In the first approach, that was evaluated, every process generates new nodes independantly from each other.
+Then the possible neighbours are determined and the connections, which have to be tested.
 The indicator function is called and returnes the edges.
@@
 \subsubsection*{Version 2}
-The second approach
-
-	- every process generates all nodes (same by exchange of random seed)
-
-	- indicator function splitted over processes, synchronous
+The second approach uses the fact, that by synchronizing the random seed, 
+different processes can produce the same sequence of random numbers independently from each other.
+By using this, every process can generate all new nodes and determine the connection data. 
+Then the indicator function computation is splitted over their configuration space instances and the returned data shared afterwards with each other.
+With this data, all graphs can be updated and it is guaranteed, that they are the same on each process.
+The purpose of this version compared to the first one is, that less data has to be sent. 
+Furthermore, in the first version it can occur, that the different generated nodes can result in very different counts of neighbours and therefore different loads on the GPUs.
+Here, the distribution of the computation amount over the different GPUs can be chosen almost freely.
 
@@
 
-	- same as version 2 with some pipeline optimizations
-
-version 4:
-
-	- same as version 3 with asynchronous gpu
-
-version 5:
-
-	- each process has 2 workers which are generating their nodes an are using 1 gpu
-
-	- one worker uses the kernel time  of the other to do its work
-
-	- should give use the gpu computation time efficiently
-
-Stateflow for each PRMSolver instance:
+The third version does a further optimization by using the asynchronous version of the indicator function.
+Here, every process runs two workers overlapped.
+Here, in the time, when the first worker has generated his indicator function request and launched the kernel and is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round.
+With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
 
@@
 - difficult to test because
+
 	- not deterministic
+
 	- dependant on density of nodes: how many neighbours have to be tested per connection

### 2016-03-30 paper/paper.tex
...

@@
 \section{Motivation}
-In trajctory optimization with obstacles, it is often necessary to provide feasible inital trajectories. 
+In trajectory optimization with obstacles, it is often necessary to provide feasible initial trajectories. 
 For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive.
@@
 An efficient approximative approach is using probabilistic roadmaps (PRMs).
+
+In this project, a parallelized version of the PRM method has been developed in general.
+The robotics application then was implemented as an special case of this with GPU support.
 
@@
 
-\subsection{Storage of the Graph}
+\subsection{Storing the Graph}
 
-- describe
+Apart from the indicator function evaluation, the main problem in building the probabilistic roadmap is, that for each new node $q$ it have to be determined all possible neighbors.
+That is the set of all nodes $\tilde q$ of the graph, with $\norm{\tilde q - q} \le D$.
+If no special order of the nodes is given, one has to calculate the norms for all nodes of the graph, which is rather time consuming.
 
-- compare different models...?
+Therefore a special indexing is used here. 
+The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map.
+
+\begin{lstlisting}
+struct block{
+	int pos;	 //! position of first vertex
+	int num;	 //! current number of vertices stored
+	block* next; //! if num==blocksize -> pointer to next block
+};
+\end{lstlisting}
+
+\begin{lstlisting}
+struct graph{
+    std::map<int,block*> map; //map for accessing blocks
+    std::vector<block> blocks;
+
+    std::vector<float> qstorage; //length ndof*N
+
+    int newblockpos;  //position of next new block
+    int blocknum;     //number of used blocks
+    .
+    .
+    .
+};
+\end{lstlisting}
+
+The key of an block is computed by the mapping
+\begin{equation*}
+	key(q) = \lfloor q(1)/H \rfloor
+\end{equation*}
+such that $H>0$ becomes a parameter to adjust the grid density.
+This has to be chosen so large that enough nodes are stored in one block in order to have efficient memory access.
+On the other side, the smaller it is, the less computations have to be done for the neighborhood requests.
+For illustration the following pseudocode shows the insertion of a node.
+
+\begin{algorithm}[H]
+	\KwData{$q\in\R^d$}
+	Compute key=key(q)\\ 
+	\eIf{key exists in map}{
+		block=map[key]\\ 
+		\While{ block full }{
+			block=block->next
+		}
+	}{
+		insert new block at map[key]
+	}
+	Insert node in block
+	\caption{Inserting a node}
+\end{algorithm}
 
@@
 
+Tested have been made on a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
+
 - timings of different versions at same conditions

### 2016-03-30 prm/matlab/lib/pathplot.m

@@ -1,16 +1,18 @@
-function [] = pathplot( dhparams, ndof, polys, N, Q, dt, infinite )
+function [] = pathplot( dhparams, ndof, polys, N, Q, dt, infinite)
     
     for i=1:N
         polys{i}.K=convhulln(polys{i}.vertices);
     end
     
-    clf;
+    colors=repmat([1,1,0.5; 0.5,0.5,1],ndof+1,1);
+    
     l=size(Q,2);
     while 1
     for t=[1:l,l:-1:1]
         lim=[xlim;ylim;zlim];
+        
         clf;
-        stateplot(dhparams,ndof,polys,N,Q(:,t));
+        stateplot(dhparams,ndof,polys,N,Q(:,t),colors);
         
         limneu=[xlim;ylim;zlim];
         limneu=[min(limneu(:,1),lim(:,1)), max(limneu(:,2),lim(:,2))];
@@ -24,5 +26,7 @@ function [] = pathplot( dhparams, ndof, polys, N, Q, dt, infinite )
             break;
         end
     end
+    
+    hold off
     
 end

### 2016-03-30 prm/matlab/lib/stateplot.m

@@ -1,15 +1,18 @@
-function [] = stateplot( dhparams, ndof, polys, N, q )
+function [] = stateplot( dhparams, ndof, polys, N, q , colors)
 % Q= [q1, ...., ql] describing a path
 
     hold on
     
     T=calctrafos(dhparams,ndof,q);
     
+    
     for i=1:N
         V=[polys{i}.vertices,ones(size(polys{i}.vertices,1),1)];
         X=V*T{polys{i}.sys+1}';
         X=X(:,1:3);
-        patch('Faces',polys{i}.K,'Vertices',X,'FaceColor',[1,1,0.5],'FaceAlpha',1.0);
+        patch('Faces',polys{i}.K,'Vertices',X,...
+            'FaceColor',colors(polys{i}.sys+1,:),...
+            'FaceAlpha',1.0);
     end
     
     x=zeros(1,ndof);

### 2016-03-30 paper/paper.tex
robotsolution picture

@@
 \documentclass{scrartcl} 
+%\documentclass[a4paper]{article}
+%\usepackage{a4wide}
+
 \usepackage[ngerman,english]{babel}
@@
 \usepackage{graphicx}
+
+\usepackage{wrapfig}
 
@@
 
+\newpage
 \section{Results}
 
-Tested have been made on a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
+\begin{wrapfigure}{r}{0.5\textwidth}\label{robotsolution}
+	\centering
+	\includegraphics[width=0.5\textwidth]{robotsolution.pdf}
+	\caption{PRM generated collisionfree movement for a 4-axis robot}
+\end{wrapfigure}
+
+Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
 
@@
 
-	- dependant on density of nodes: how many neighbours have to be tested per connection
+	- dependant on density of nodes: how many neighbors have to be tested per connection
 

### 2016-03-30 prm/matlab/lib/pathplot.m

@@ -4,11 +4,16 @@ function [] = pathplot( dhparams, ndof, polys, N, Q, dt, infinite)
         polys{i}.K=convhulln(polys{i}.vertices);
     end
     
-    colors=repmat([1,1,0.5; 0.5,0.5,1],ndof+1,1);
-    
+    colors=[0.8,0.8,0.8;
+            repmat([1,1,0.5; 0.5,0.5,1],ndof+1,1)];
+     
+    clf;
     l=size(Q,2);
     while 1
-    for t=[1:l,l:-1:1]
+    for t=[1:l]%,l:-1:1]
+        %colors=repmat([t/l,1,0.5; 0.5,0.5,t/l],ndof+1,1);
+        %colors=repmat([t/l,1,0.5],ndof+1,1);
+    
         lim=[xlim;ylim;zlim];
         
         clf;

### 2016-03-30 prm/matlab/lib/stateplot.m

@@ -44,7 +44,7 @@ function [] = stateplot( dhparams, ndof, polys, N, q , colors)
     text(x,y,z,labels, 'HorizontalAlignment', 'right');
     for i=1:(ndof+1)
         l=1000;
-        plot3(x(:,i)+[1,-1]*l*xr(:,i),y(:,i)+[1,-1]*l*yr(:,i),z(:,i)+[1,-1]*l*zr(:,i),'-.');
+        %plot3(x(:,i)+[1,-1]*l*xr(:,i),y(:,i)+[1,-1]*l*yr(:,i),z(:,i)+[1,-1]*l*zr(:,i),'-.');
     end
     xlim(lim(1,:));
     ylim(lim(2,:));

### 2016-03-30 prm/matlab/robotplot.m

@@ -16,7 +16,7 @@ t=repmat(seg, 1, l)+repelem(1:l,length(seg));
 
 Qpathcont=interp1q((1:l)',prmoutput.Qpath',(1:0.2:l)')';
 %Qpathcont=interp1q((1:l)',prmoutput.Qpath',t')';
-
+%Qpathcont=interp1q((1:l)',prmoutput.Qpath',[1;18;20;21;25;l])';
 
 
 pathplot(dhparams,ndof,polys,N,Qpathcont, 0.01, false);

### 2016-03-30 prm/src/lib/collision4/cuda_head.h

@@ -27,7 +27,7 @@
 #endif
 
 #ifndef GPU_VERSION
-  #warning ("GPU_VERSION not defined: version without gpu");
+  //#warning ("GPU_VERSION not defined: version without gpu");
 #endif
 
 

### 2016-03-30 prm/src/lib/collision4/geo4.h

@@ -8,7 +8,7 @@
 
 
 #ifndef CUDA_CODE
-  #warning ("NO CUDA_CODE: defining float2 and float4");
+  //#warning ("NO CUDA_CODE: defining float2 and float4");
   struct float4{
     float x,y,z,w;
   };

### 2016-03-30 prm/src/lib/prm/prmsolver.cpp

@@ -1795,14 +1795,14 @@
     float *distlist2=new float[nbuf];
 
 
-    processor5 processor1(rank,size,
+    worker processor1(rank,size,
                          qnew, num, dsp, cnt,
                          leftconn, rightconn,
                          poslist, distlist,
                          qstartlist,qendlist,resbuf, resbufloc,nbuf,offset,
                          this);
 
-    processor5 processor2(rank,size,
+    worker processor2(rank,size,
                          qnew2, num, dsp, cnt,
                          leftconn2, rightconn2,
                          poslist2, distlist2,
@@ -1878,7 +1878,7 @@
 
 
   template<int ndof>
-  PRMSolver<ndof>::processor5::processor5( const int mpirank_, const int mpisize_,
+  PRMSolver<ndof>::worker::worker( const int mpirank_, const int mpisize_,
                                    float* qnew_, const int num_, const int *dsp_, const int *cnt_,
                                    int *leftconn_, int *rightconn_,
                                    int *poslist_, float *distlist_,
@@ -1905,7 +1905,7 @@
     }
 
     template<int ndof>
-    PRMSolver<ndof>::processor5::~processor5(){
+    PRMSolver<ndof>::worker::~worker(){
         delete posqlist;
         delete numqlist;
         delete numqlistleft;
@@ -1917,7 +1917,7 @@
 
 
     template<int ndof>
-    int PRMSolver<ndof>::processor5::processing_step_part1()
+    int PRMSolver<ndof>::worker::processing_step_part1()
     {
 
 
@@ -2001,7 +2001,7 @@
 
 
     template<int ndof>
-    int PRMSolver<ndof>::processor5::processing_step_part2(){
+    int PRMSolver<ndof>::worker::processing_step_part2(){
 
 #ifndef NO_IO
     printarr(dsp,mpisize);
@@ -2042,7 +2042,7 @@
 
 
     template<int ndof>
-    int PRMSolver<ndof>::processor5::processing_step_part3(){
+    int PRMSolver<ndof>::worker::processing_step_part3(){
 
       for(int j=0;j<num;++j){
 

### 2016-03-30 prm/src/lib/prm/prmsolver.h

@@ -164,9 +164,9 @@ public:
   int process_mpi5(const int numall, const int nbuf, const int maxsteps, int seed);
 
 
-  friend class processor5;
+  friend class worker;
 
-  class processor5{
+  class worker{
   private:
       const int mpirank;
       const int mpisize;
@@ -210,13 +210,13 @@ public:
       graph &graphr;
 
   public:
-    processor5( const int mpirank_, const int mpisize_,
+    worker( const int mpirank_, const int mpisize_,
                                    float* qnew_, const int num_, const int *dsp_, const int *cnt_,
                                    int *leftconn_, int *rightconn_,
                                    int *poslist_, float *distlist_,
                                    float* qstart_, float* qend_, int* resbuf_, int *resbufloc_, const int nbuf_, const int offset_,
                                    PRMSolver *instance_);
-    ~processor5();
+    ~worker();
         inline int processing_step_part1();
         inline int processing_step_part2();
         inline int processing_step_part3();

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.cpp
numthreadsall

@@ -48,6 +48,8 @@ void RobotConfigspace<ndof>::construct(const Robot<ndof>* robot_,
   devloaded=false;
 
   requeststack_id=0;
+
+  numthreads_all=0;
 }
 
 //! robot_: Robot Object with Denavit-Hartenberg data
@@ -512,6 +514,8 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
 
     //printarr(res,N);
     //printvar(numthreads);
+
+    numthreads_all+=numthreads;
 
     return 0; //TODO error handling?
 }

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.h

@@ -105,6 +105,8 @@ public:
   //! dimension d
   int dim(){return ndof;}
 
+  int get_numthreads_all() const {return numthreads_all;}
+
 private:
 
   Kinematics<ndof>* kin;
@@ -156,6 +158,9 @@ private:
       int N;
   };
   std::map<int,request_data> requeststack;
+
+  //!number of threads launched since creation
+  int numthreads_all;
 
 };
 

### 2016-03-30 prm/src/main.cpp

@@ -103,6 +103,7 @@ int main(int argc, char** argv)
   //prm.print();
 
   MPI_Barrier(MPI_COMM_WORLD);
+  printvar(space.get_numthreads_all());
   if(rank==0){
       tick(twrite);
 

### 2016-03-30 prm/src/lib/config.h
...

@@ -3,7 +3,6 @@
 
 
 #define SILENT
-#define NO_IO
 
 
 #endif // CONFIG_H

### 2016-03-30 prm/src/lib/prm/prmsolver.cpp

@@ -10,6 +10,7 @@
 #include <tictoc.h>
 
 
+#define NO_IO
 
 
     //! *******************

### 2016-03-30 prm/src/main.cpp

@@ -103,7 +103,9 @@ int main(int argc, char** argv)
   //prm.print();
 
   MPI_Barrier(MPI_COMM_WORLD);
-  printvar(space.get_numthreads_all());
+  int numthreads=space.get_numthreads_all();
+  int numthreadsall;
+  MPI_Reduce(&numthreads, &numthreadsall,1,MPI_INT,MPI_SUM,0,MPI_COMM_WORLD);
   if(rank==0){
       tick(twrite);
 

### 2016-03-30 prm/src/main.cpp
...

@@ -107,6 +107,8 @@ int main(int argc, char** argv)
   int numthreadsall;
   MPI_Reduce(&numthreads, &numthreadsall,1,MPI_INT,MPI_SUM,0,MPI_COMM_WORLD);
   if(rank==0){
+      printvar(numthreadsall);
+
       tick(twrite);
 
       prm.store_results("prmoutput");

### 2016-03-30 prm/src/main.cpp
...

@@ -107,13 +107,14 @@ int main(int argc, char** argv)
   int numthreadsall;
   MPI_Reduce(&numthreads, &numthreadsall,1,MPI_INT,MPI_SUM,0,MPI_COMM_WORLD);
   if(rank==0){
-      printvar(numthreadsall);
 
       tick(twrite);
 
       prm.store_results("prmoutput");
 
       tock(twrite);
+
+      printvar(numthreadsall);
 
       printvar(rand());
       printvar(firstrand);

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.cpp
...

@@ -439,10 +439,10 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
 #ifdef GPU_VERSION
     int id;
     if(free_resdevbuffer_ids.empty()){
-        msg("RobotConfigspace: adding new resdevbuffer");
-        printvar(resdevbuffers.size());
+        //msg("RobotConfigspace: adding new resdevbuffer");
+        //printvar(resdevbuffers.size());
         id=resdevbuffers.size();
-        streams.resize(0x0);
+        streams.push_back(0x0);
         cudaassert(cudaStreamCreate(&streams[id]));
         resdevbuffers.push_back(0x0);
         cudaassert(cudaMalloc((void**)&resdevbuffers[id], nbufres*sizeof(int)));

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.cpp
...

@@ -439,8 +439,8 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
 #ifdef GPU_VERSION
     int id;
     if(free_resdevbuffer_ids.empty()){
-        //msg("RobotConfigspace: adding new resdevbuffer");
-        //printvar(resdevbuffers.size());
+        msg("RobotConfigspace: adding new resdevbuffer");
+        printvar(resdevbuffers.size());
         id=resdevbuffers.size();
         streams.push_back(0x0);
         cudaassert(cudaStreamCreate(&streams[id]));

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.cpp
...

@@ -454,6 +454,7 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
         cudaassert(cudaMalloc((void**)&testnumdev[id], nbuftest*sizeof(int)));
         testposdev.push_back(0x0);
         cudaassert(cudaMalloc((void**)&testposdev[id], nbuftest*sizeof(int)));
+        printvar(resdevbuffers.size());
     }else{
         std::set<int>::iterator it=free_resdevbuffer_ids.begin();
         id=*it;

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.cpp
...

@@ -439,8 +439,8 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
 #ifdef GPU_VERSION
     int id;
     if(free_resdevbuffer_ids.empty()){
-        msg("RobotConfigspace: adding new resdevbuffer");
-        printvar(resdevbuffers.size());
+        //msg("RobotConfigspace: adding new resdevbuffer");
+        //printvar(resdevbuffers.size());
         id=resdevbuffers.size();
         streams.push_back(0x0);
         cudaassert(cudaStreamCreate(&streams[id]));
@@ -454,7 +454,7 @@ int RobotConfigspace<ndof>::indicator2_async(const float* qs, const float* qe, i
         cudaassert(cudaMalloc((void**)&testnumdev[id], nbuftest*sizeof(int)));
         testposdev.push_back(0x0);
         cudaassert(cudaMalloc((void**)&testposdev[id], nbuftest*sizeof(int)));
-        printvar(resdevbuffers.size());
+        //printvar(resdevbuffers.size());
     }else{
         std::set<int>::iterator it=free_resdevbuffer_ids.begin();
         id=*it;

### 2016-03-30 prm/src/lib/prm/prmsolver.cpp
...

@@ -1815,7 +1815,7 @@
     processor1.processing_step_part1();
     processor2.processing_step_part1();
 
-    tick(tloop);
+    //tick(tloop);
     for(int i=0;i<maxsteps;++i){
       //tick(evaluating1);
       processor1.processing_step_part2();
@@ -1847,7 +1847,7 @@
         printvar(i);
       }
     }
-    tock(tloop);
+    //tock(tloop);
 
     delete qnew;
     delete qstartlist;

### 2016-03-30 prm/src/lib/robotspace/robotconfigspace.cpp

@@ -123,10 +123,7 @@ int RobotConfigspace<ndof>::load_config(std::string path, Robot<ndof>* &robot, p
   read_file(path+"/mins.bin",mins,ndof);
   read_file(path+"/maxs.bin",maxs,ndof);
 
-  printvar(mins[0]+pi);
-  printvar(maxs[0]-1.5*pi);
-
-    //! Polytopes
+  //! Polytopes
 
   read_file(path+"/polys/N.bin",&N,1);
   if(N<=0){

### 2016-03-30 prm/src/lib/prm/prmsolver.cpp
...

@@ -1872,7 +1872,7 @@
     delete distlist2;
 
 
-    msg("process_mpi5 finished");
+    //msg("process_mpi5 finished");
     return 0;
   }
 

### 2016-03-30 prm/src/lib/prm/prmsolver.h
...

@@ -81,7 +81,7 @@ public:
 
   //! *******************
   //! *                 *
-  //! *    process 1    *
+  //! *    process 1    * (=version 1 in paper)
   //! *                 *
   //! *******************
 
@@ -139,7 +139,7 @@ public:
 
   //! *******************
   //! *                 *
-  //! *    process 4    *
+  //! *    process 4    * (=version 2 in paper)
   //! *                 *
   //! *******************
 
@@ -156,7 +156,7 @@ public:
 
   //! *******************
   //! *                 *
-  //! *    process 5    *
+  //! *    process 5    * (=version 3 in paper)
   //! *                 *
   //! *******************
 

### 2016-03-30 prm/src/bench.cpp
...

@@ -1,3 +1,4 @@
+
 #include <iostream>
 #include <time.h>
 #include <mpi.h>
@@ -27,8 +28,6 @@ int main(int argc, char** argv)
   const int ndof=4;
 
 
-  tick(tinit);
-
   //! Parameters
 
   int num=(argc>=2 ? atoi(argv[1]) : 32 );
@@ -39,12 +38,11 @@ int main(int argc, char** argv)
   int seed=(argc>=6 ? atoi(argv[5]) : 0 );
   int blocksize=(argc>=7 ? atoi(argv[6]) : 256);
   int prmversion=(argc>=8 ? atoi(argv[7]) : 5 );
-  int maxstorage=1024*1024;//(argc>=8 ? atoi(argv[7]) : 1024*1024);
+  int maxstorage=1024*1024;
   int maxsteps=100000;
 
-  //srand(time(NULL));
-  //srand(clock());
-  //srand(rank+time(NULL));
+
+
   srand(seed+rank*10);
   int firstrand=rand();
 
@@ -64,16 +62,8 @@ int main(int argc, char** argv)
     qend[dof]=3;
   }
 
-  if(rank==0){
-    printvar(space.indicator(&qstart[0]));
-    printvar(space.indicator(&qend[0]));
-  }
-
   prm.init(&qstart[0],&qend[0]);
 
-  //prm.print();
-
-  tock(tinit);
 
 
   //! run
@@ -81,10 +71,10 @@ int main(int argc, char** argv)
   tick(trun);
 
 
+  int version;
+
   //prm.process_mpi(num,nbuf,maxsteps);
-  if(prmversion==1){
-    prm.process_mpi(num,nbuf,maxsteps, seed);
-  }if(prmversion==2){
+  if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
   }else if(prmversion==3){
     prm.process_mpi3(num,nbuf,maxsteps, seed);
@@ -99,35 +89,17 @@ int main(int argc, char** argv)
 
   tock(trun);
 
-
   //! write output
-
-  //prm.print();
 
   MPI_Barrier(MPI_COMM_WORLD);
   int numthreads=space.get_numthreads_all();
   int numthreadsall;
   MPI_Reduce(&numthreads, &numthreadsall,1,MPI_INT,MPI_SUM,0,MPI_COMM_WORLD);
   if(rank==0){
-
-      tick(twrite);
-
-      prm.store_results("prmoutput");
-
-      tock(twrite);
-
[…]

### 2016-03-30 prm/src/lib/tictoc.h

@@ -12,6 +12,10 @@
 #define tock(t1) {std::chrono::high_resolution_clock::time_point endtime___=std::chrono::high_resolution_clock::now(); \
                  int time=std::chrono::duration_cast<std::chrono::milliseconds>(endtime___-t1).count();                \
                  std::cout << "Since " << #t1 << ": \t" << time << "ms" << std::endl; }
+#define tockval(t1,time__) {std::chrono::high_resolution_clock::time_point endtime___=std::chrono::high_resolution_clock::now(); \
+                 time__=std::chrono::duration_cast<std::chrono::milliseconds>(endtime___-t1).count();                \
+                 std::cout << "Since " << #t1 << ": \t" << time__ << "ms" << std::endl; }
+
 #define timedelta(t1,t2) {int time=std::chrono::duration_cast<std::chrono::millizseconds>(endtime___-t1).count(); \
                           std::cout << "Since " << #t1 << ": \t" << time << "ms" << std::endl;}
 

### 2016-03-30 prm/src/bench.cpp
...

@@ -74,14 +74,22 @@ int main(int argc, char** argv)
   int version;
 
   //prm.process_mpi(num,nbuf,maxsteps);
-  if(prmversion==2){
+
+  if(prmversion==1){
+    prm.process_mpi(num,nbuf,maxsteps, seed);
+    version=1;
+  }if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
+    version=-1;
   }else if(prmversion==3){
     prm.process_mpi3(num,nbuf,maxsteps, seed);
+    version=-1;
   }else if(prmversion==4){
     prm.process_mpi4(num,nbuf,maxsteps, seed);
+    version=2;
   }else if(prmversion==5){
-      prm.process_mpi5(num,nbuf,maxsteps, seed);
+    prm.process_mpi5(num,nbuf,maxsteps, seed);
+    version=3;
   }else {
     msg("Error: prmprocess not valid");
     return 0;

### 2016-03-30 prm/src/bench.cpp
...

@@ -65,18 +65,15 @@ int main(int argc, char** argv)
   prm.init(&qstart[0],&qend[0]);
 
 
-
   //! run
 
   tick(trun);
 
 
-  int version;
-
-  //prm.process_mpi(num,nbuf,maxsteps);
+  int version=-1;
 
   if(prmversion==1){
-    prm.process_mpi(num,nbuf,maxsteps, seed);
+    prm.process_mpi(num,nbuf,maxsteps);
     version=1;
   }if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);

### 2016-03-30 prm/src/main.cpp

@@ -47,6 +47,7 @@ int main(int argc, char** argv)
   //srand(rank+time(NULL));
   srand(seed+rank*10);
   int firstrand=rand();
+  printvar(firstrand);
 
 
   //! Initialization
@@ -71,8 +72,6 @@ int main(int argc, char** argv)
 
   prm.init(&qstart[0],&qend[0]);
 
-  //prm.print();
-
   tock(tinit);
 
 
@@ -80,10 +79,8 @@ int main(int argc, char** argv)
 
   tick(trun);
 
-
-  //prm.process_mpi(num,nbuf,maxsteps);
   if(prmversion==1){
-    prm.process_mpi(num,nbuf,maxsteps, seed);
+    prm.process_mpi(num,nbuf,maxsteps);
   }if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
   }else if(prmversion==3){

### 2016-03-30 prm/src/bench.cpp
...

@@ -103,7 +103,7 @@ int main(int argc, char** argv)
   if(rank==0){
     ofstream file;
     file.open("results.txt",ios::app);
-    file<<size<<"\t"<<prmversion<<"\"
+    file<<size<<"\t"<<prmversion<<"\n";
     file.close();
   }
 

### 2016-03-30 prm/src/main.cpp
...

@@ -47,7 +47,7 @@ int main(int argc, char** argv)
   //srand(rank+time(NULL));
   srand(seed+rank*10);
   int firstrand=rand();
-  printvar(firstrand);
+  //printvar(firstrand);
 
 
   //! Initialization
@@ -66,8 +66,8 @@ int main(int argc, char** argv)
   }
 
   if(rank==0){
-    printvar(space.indicator(&qstart[0]));
-    printvar(space.indicator(&qend[0]));
+    assert(space.indicator(&qstart[0])==0);
+    assert(space.indicator(&qend[0])==0);
   }
 
   prm.init(&qstart[0],&qend[0]);
@@ -94,7 +94,8 @@ int main(int argc, char** argv)
     return 0;
   }
 
-  tock(trun);
+  int time;
+  tockval(trun,time);
 
 
   //! write output
@@ -125,6 +126,11 @@ int main(int argc, char** argv)
       printvar(seed);
       printvar(blocksize);
       printvar(prmversion);
+
+      ofstream file;
+      file.open("stats.txt",ios::app);
+      file<<size<<"\t"<<prmversion<<"\t"<<time<<"\t"<<numthreads<<"\t"<<(time*1000000)/numthreads<<"\n";
+      file.close();
   }
 
   MPI_Finalize();

### 2016-03-30 prm/src/main.cpp
...

@@ -129,7 +129,7 @@ int main(int argc, char** argv)
 
       ofstream file;
       file.open("stats.txt",ios::app);
-      file<<size<<"\t"<<prmversion<<"\t"<<time<<"\t"<<numthreads<<"\t"<<(time*1000000)/numthreads<<"\n";
+      file<<size<<"\t"<<prmversion<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreads<<"\n";
       file.close();
   }
 

### 2016-03-30 prm/src/main.cpp
...

@@ -79,16 +79,22 @@ int main(int argc, char** argv)
 
   tick(trun);
 
+  int version=-1;
   if(prmversion==1){
     prm.process_mpi(num,nbuf,maxsteps);
+    version=1;
   }if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
+    version=-1;
   }else if(prmversion==3){
     prm.process_mpi3(num,nbuf,maxsteps, seed);
+    version=-1;
   }else if(prmversion==4){
     prm.process_mpi4(num,nbuf,maxsteps, seed);
+    version=2;
   }else if(prmversion==5){
-      prm.process_mpi5(num,nbuf,maxsteps, seed);
+    prm.process_mpi5(num,nbuf,maxsteps, seed);
+    version=3;
   }else {
     msg("Error: prmprocess not valid");
     return 0;
@@ -129,7 +135,7 @@ int main(int argc, char** argv)
 
       ofstream file;
       file.open("stats.txt",ios::app);
-      file<<size<<"\t"<<prmversion<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreads<<"\n";
+      file<<size<<"\t"<<version<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreads<<"\n";
       file.close();
   }
 

### 2016-03-30 prm/src/bench.cpp
...

@@ -75,7 +75,7 @@ int main(int argc, char** argv)
   if(prmversion==1){
     prm.process_mpi(num,nbuf,maxsteps);
     version=1;
-  }if(prmversion==2){
+  }else if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
     version=-1;
   }else if(prmversion==3){

### 2016-03-30 prm/src/main.cpp

@@ -83,7 +83,7 @@ int main(int argc, char** argv)
   if(prmversion==1){
     prm.process_mpi(num,nbuf,maxsteps);
     version=1;
-  }if(prmversion==2){
+  }else if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
     version=-1;
   }else if(prmversion==3){

### 2016-03-30 prm/src/main.cpp
...

@@ -135,7 +135,7 @@ int main(int argc, char** argv)
 
       ofstream file;
       file.open("stats.txt",ios::app);
-      file<<size<<"\t"<<version<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreads<<"\n";
+      file<<size<<"\t"<<version<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
       file.close();
   }
 

### 2016-03-30 prm/src/main.cpp
...

@@ -135,7 +135,7 @@ int main(int argc, char** argv)
 
       ofstream file;
       file.open("stats.txt",ios::app);
-      file<<size<<"\t"<<version<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
+      file<<size<<"\t"<<version<<"\t"<<seed<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
       file.close();
   }
 

### 2016-03-30 prm/run.sh
...

@@ -1,0 +1,10 @@
+mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 1
+mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 1
+mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 1
+mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 4
+mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 4
+mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 4
+mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 5
+mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 5
+mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 5
+

### 2016-03-30 prm/src/main.cpp

@@ -135,7 +135,7 @@ int main(int argc, char** argv)
 
       ofstream file;
       file.open("stats.txt",ios::app);
-      file<<size<<"\t"<<version<<"\t"<<seed<<"\t"<<time<<"\t"<<numthreads<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
+      file<<size<<"\t"<<version<<"\t"<<seed<<"\t"<<time<<"\t"<<numthreadsall<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
       file.close();
   }
 

### 2016-03-30 prm/run.sh
...

@@ -1,3 +1,4 @@
+make -j4
 mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 1
 mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 1
 mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 1

### 2016-03-30 prm/run.sh
...

@@ -1,4 +1,4 @@
-make -j4
+#make -j4
 mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 1
 mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 1
 mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 1

### 2016-03-30 prm/src/main.cpp
...

@@ -135,7 +135,7 @@ int main(int argc, char** argv)
 
       ofstream file;
       file.open("stats.txt",ios::app);
-      file<<size<<"\t"<<version<<"\t"<<seed<<"\t"<<time<<"\t"<<numthreadsall<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
+      file<<version<<"\t"<<size<<"\t"<<seed<<"\t"<<time<<"\t"<<numthreadsall<<"\t"<<((long)time*1000000)/numthreadsall<<"\n";
       file.close();
   }
 

### 2016-03-30 paper/paper.tex
paper fast fertig

@@
 
+\newpage
 \subsection{Project Overview}
@@
 
-Algorithms 1 and 2 two show in pseudo code, how the work is split between host and device.
+Algorithms \ref{host} and \ref{kernel} two show in pseudo code, how the work is split between host and device.
 
-\begin{algorithm}
+\begin{algorithm}\label{host}
 	\KwData{qs, qe, N, offset}
@@
 
-\begin{algorithm}
+\begin{algorithm}\label{kernel}
 	\KwData{arrays qs, qe, testpos, testnum and threadindex}
@@
 
-\section{Structure of the PRM Solver}\label{solver}
+\newpage
+\section{The PRM Solver}\label{solver}
 
@@
     std::vector<block> blocks;
-
     std::vector<float> qstorage; //length ndof*N
-
     int newblockpos;  //position of next new block
     int blocknum;     //number of used blocks
-    .
     .
@@
 Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
+Because of the dependancy on random numbers, the algorithm takes different amounts of time for different runs.
+Therefore, 10 runs were made for each version and number of processes.
+However, the different versions lead to a different average amount of GPU threads needed.
+So, to be able to better compare how they scale, the time per number of GPU threads used as a quantity.
+Table \ref{resulttable} shows the results.
+One can see, that while the versions with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
+The versions with less communication however have better absolute times.
+Furthermore, one sees the improvement from version 2 to version 3 from the asynchronous indicator function.
 
-- timings of different versions at same conditions
+\ \\ 
+\begin{table}[h]\label{resulttable}
+\centering
+\begin{tabular}{| c | c | r | r | r |}
 
-- difficult to test because
+\hline
+version	&  processes/	& time (ms)	& GPU  		& time per 1e6  \\ 
+		& GPUs		& 			& threads		& GPU threads (ms) \\ 
+\hline
+		& 1			& 1002		& 696413		& 1448	\\ 
+1		& 2			& 763		& 838938		& 918	\\ 
+		& 4			& 706		& 1113781	& 647	\\ 
+\hline
+		& 1			& 900		& 608172		& 1490	\\ 
+2		& 2			& 647		& 608172		& 1070	\\ 
+		& 4			& 568		& 608172		& 947	\\ 
+\hline
+		& 1			& 696		& 599847		& 1181	\\ 		
+3		& 2			& 554		& 599847		& 951	\\ 
+		& 4			& 500		& 599847		& 864	\\ 
+\hline
 
-	- not deterministic
-
-	- dependant on density of nodes: how many neighbors have to be tested per connection
-
-\section{Conclusion}
+\end{tabular}
+\caption{Timing results for the parallelization versions from chapter \ref{solver}}
+\end{table}
 

### 2016-03-30 paper/paper.tex
paper im prinzip fertig

@@
 
-- introduction
+The solver builds in the first phase the roadmap graph from both the start and the end node, until it is connected.
+It is always ensured that the graph consists of two connected components containing the start and the end.
+With this, it can be easily determined at each time, if the whole graph has become connected.
 
-- dijkstra at the end makes almost no contribution to the whole time
-
-	- can be parallelized, but this is again complicated
-
-- focus on building the graph
+At the end a shortest path from the start to the end node is searched by the Dijkstra algorithm.
+It has shown that this part a neglectible amount of time compared to the graph building phase.
+Therefore it is not parallelized here and the focus lies on building the roadmap.
 
@@
 \begin{equation*}
-	key(q) = \lfloor q(1)/H \rfloor
+	key(q) = \left\lfloor \frac{q_1}{H} \right\rfloor
 \end{equation*}
@@
 
+For getting the neighbors of a node $q$ it is then enough to look at the keys from $key(q-cD)$ to $key(q+cD)$. For $D=H$, $c=1$ is possible.
+The key mapping could be generalized by taking more components of $q$ into account, but that was not done here.
+
 \subsection{Parallelization Models}
@@
 The indicator function is called and returnes the edges.
-After this the graph must be synchronized on every processes, which leads to an alltoall communication.
-As in this graph building does not need the edge data, it is sufficient to exchange only the nodes in this phase.
+After this, the graph must be synchronized between all processes, which needs an all-to-all communication.
+As the graph building phase does not need the edge data, it is sufficient to exchange only the nodes and collect the edges at the end.
 
@@
 Then the indicator function computation is splitted over their configuration space instances and the returned data shared afterwards with each other.
-With this data, all graphs can be updated and it is guaranteed, that they are the same on each process.
+With this data, all graphs can be updated and it is guaranteed, that they grow exactly in the same way on each process.
 The purpose of this version compared to the first one is, that less data has to be sent. 
-Furthermore, in the first version it can occur, that the different generated nodes can result in very different counts of neighbours and therefore different loads on the GPUs.
-Here, the distribution of the computation amount over the different GPUs can be chosen almost freely.
+Furthermore, in the first version it can occur, that the different generated nodes can result in very different counts of neighbors and therefore uneven loads on the GPUs.
+Here, the distribution of the computation amount over the GPUs can be chosen almost freely.
 
@@
 
-\ \\ 
 \begin{table}[h]\label{resulttable}
@@
 \end{tabular}
-\caption{Timing results for the parallelization versions from chapter \ref{solver}}
+\caption{Timing results for the parallelization versions from chapter \ref{solver} on a machine with 4 GeForce GTX 480}
 \end{table}
+
+As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
+In most cases, the GPU code was around 10 times faster. However, the robot code is not optimized for the CPU version, such that this comparison is not quite fair.
+
+\newpage
 

### 2016-03-31 paper/paper.tex
...
rechtschreibung

@@
 Therefore $I(q)$ is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every thread computes one evaluation.
-In the robotics case, this means, the direct kinematic and collision calculation are made on the GPU parallely for different joint angle configurations.
-Although this will the only major application in the project, the PRM solver and the configuration space are implemented seperated from each other and connected by an interface to be free to simply add other applications.
+In the robotics case, this means, the direct kinematic and collision calculation are made on the GPU parallelly for different joint angle configurations.
+Although this will the only major application in the project, the PRM solver and the configuration space are implemented separated from each other and connected by an interface to be free to simply add other applications.
 As an example and for testing, additionally a simple 2D configuration space is provided,
@@
 While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
-a second level can be achieved by sampling and connecting several new nodes parallely. 
+a second level can be achieved by sampling and connecting several new nodes parallelly. 
 This is more difficult, because all new nodes have to be connected 
@@
 
-In the project, the probabilistic roadmap algorithm is applied  to create collisionfree trajectories of a robot arm.
+In the project, the probabilistic roadmap algorithm is applied  to create collision-free trajectories of a robot arm.
 For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
@@
 Now the movement of the robot can be described by a curve $q(t) \in \R^d$ in the configuration space.
-It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
+It is collision-free, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
 To apply the probabilistic roadmap algorithm, we express this by the indicator function
@@
 The Denavit-Hartenberg (DH) convention is a method to describe the geometry of an robots joints, 
-that is, how the transformations between the robots coordinate frames look like in dependancy of its joint values $q_1,...,q_d$.
+that is, how the transformations between the robots coordinate frames look like in dependency of its joint values $q_1,...,q_d$.
 For this it is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
@@
 To check if two polytopes collide an algorithm by Chung and Wang is used here.
-It uses the separating vector theorem and tries to find a seperating vector by iterative search.
-Furthermore, a subalgorithm determines in each iteration from the candidate history, if it is still theoretically possible to find a seperating vector.
+It uses the separating vector theorem and tries to find a separating vector by iterative search.
+Furthermore, a sub-algorithm determines in each iteration from the candidate history, if it is still theoretically possible to find a separating vector.
 If not, it is returned a collision.
-If on the other side a seperating vector is found, the algorithm can return no collision.
+If on the other side a separating vector is found, the algorithm can return no collision.
 For detailed informations see \cite{chung} and \cite{ericson}, pp. 410-412.
@@
 \begin{lstlisting}
-__host__ __device__ int seperating_vector_algorithm(const polytope4& P, const polytope4& Q, 
+__host__ __device__ int separating_vector_algorithm(const polytope4& P, const polytope4& Q, 
 													const trafo4& tp, const trafo4& tq);
@@
 	\For{all registered pairs pairs of polytopes $P, Q$}{
-		$res_{tmp}=seperating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
+		$res_{tmp}=separating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
 		\If{$res_{tmp} \neq 0$}{
@@
 In order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
-Here the function is split in everything until the kernel launch and a wait function, which waits until the kernel finishes and recieves the result data.
+Here the function is split in everything until the kernel launch and a wait function, which waits until the kernel finishes and receives the result data.
 
@@
 
-The implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corrsponding waiting calls. 
+The implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corresponding waiting calls. 
 With this it is possible to use the full computation time of the GPU.
@@
 At the end a shortest path from the start to the end node is searched by the Dijkstra algorithm.
-It has shown that this part a neglectible amount of time compared to the graph building phase.
+It has shown that this part a neglectable amount of time compared to the graph building phase.
 Therefore it is not parallelized here and the focus lies on building the roadmap.
@@
 
-In the first approach, that was evaluated, every process generates new nodes independantly from each other.
-Then the possible neighbours are determined and the connections, which have to be tested.
-The indicator function is called and returnes the edges.
+In the first approach, that was evaluated, every process generates new nodes independently from each other.
+Then the possible neighbors are determined and the connections, which have to be tested.
+The indicator function is called and returns the edges.
 After this, the graph must be synchronized between all processes, which needs an all-to-all communication.
@@
 By using this, every process can generate all new nodes and determine the connection data. 
-Then the indicator function computation is splitted over their configuration space instances and the returned data shared afterwards with each other.
+Then the indicator function computation is split over their configuration space instances and the returned data shared afterwards with each other.
 With this data, all graphs can be updated and it is guaranteed, that they grow exactly in the same way on each process.
@@
 	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
-										recieve\\ result, \\
+										receive\\ result, \\
 										store nodes \\ 
@@
 	\node[block, mywidth, right=of l6] (r6)	{\begin{tabular}{c}
-										recieve\\ result, \\
+										receive\\ result, \\
 										store nodes \\
@@
 	\includegraphics[width=0.5\textwidth]{robotsolution.pdf}
-	\caption{PRM generated collisionfree movement for a 4-axis robot}
+	\caption{PRM generated collision-free movement for a 4-axis robot}
 \end{wrapfigure}
@@
 Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
-Because of the dependancy on random numbers, the algorithm takes different amounts of time for different runs.
+Because of the dependency on random numbers, the algorithm takes different amounts of time for different runs.
 Therefore, 10 runs were made for each version and number of processes.
@@
 \end{tabular}
-\caption{Timing results for the parallelization versions from chapter \ref{solver} on a machine with 4 GeForce GTX 480}
+\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
 \end{table}

### 2016-03-31 paper/paper.tex
paper korrekturen

@@
 A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form.
-An efficient approximative approach is using probabilistic roadmaps (PRMs).
+An efficient approximative solution for this is using probabilistic roadmaps (PRMs).
 
@@
 
-The abstract problem is to find a path on a d-dimensional space, which avoids intersecting obstacles. The obstacles are  given by an indicator function
+The abstract problem treated here is to find a path in a d-dimensional space, which avoids intersecting obstacles. The area occupied by obstacles is given by an indicator function
 \begin{equation*}
@@
 \end{equation*}
-on the domain \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \) and the goal is to find a continuous way
+on a region \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \). Hence, the goal is to find a continuous curve
 \begin{equation*}
@@
 \end{equation*}
-The probabilistic roadmap (PRM) approach grows a graph \(G=(V,E)\) form \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in V\subset \Omega\).
-Two nodes \(v,w \in V\) are connected, if they are near enough each other and the linear connection in between is free from obstacles.
-The algorithm terminates, when there is a path from \(q_s\) to \(q_e\) on G.
+The PRM algorithm now grows a graph \(G=(V,E)\) form points \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in \Omega\) and inserting them into $V$.
+Two nodes \(v,w \in V\) are connected by an edge, if they are near enough each other and the line between them is free from obstacles.
+The algorithm terminates, when there exists a path from \(q_s\) to \(q_e\) on G.
 \begin{defi}
@@
 	\end{equation*}
-	We say, \(q_1\) and \(q_2\) are connected (with stepsize \(h>0\)) on the map \(\Omega\), if
+	We say, \(q_1\) and \(q_2\) are connected with stepsize \(h>0\), if
 	\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
@@
 \begin{algorithm} 
-	\caption{Probabilistic Roadmap in general}
+	\caption{Probabilistic Roadmap Algorithm}
 	\KwData{ \(q_s, q_e\in \Omega, h>0, d_0>0\)}
@@
     \includegraphics[width=0.8\textwidth]{configspacetrafo.png}
-    \caption{Example for an indicator function in robotics from \cite{prmlec}. 
+    \caption{Example from \cite{prmlec} for an indicator function in robotics. 
     		The real obstacles on the left are transformed 
@@
 
-In the project, the PRM algorithm is used to find trajectories for a robot arm, which avoid self collisions and collisions with obstacles in the environment.
-Such trajectories can be expressed as functions describing the joint variables changing over time, that means curves in the configuration space with dimension equal to the number of degrees of freedom of the robot.
-With this association, the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form stated in the previous chapter.
-The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given set of joints, if the configuration is feasible or a collision occurs.
+In this project, the PRM algorithm is used to find trajectories for a robot arm, which avoid self collisions and collisions with obstacles in the environment.
+Such trajectories can be expressed as functions describing the joint variables changing over time, that means as a curve in the configuration space with dimension equal to the number of degrees of freedom of the robot.
+With this association, the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form defined in the previous section.
+The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given vector of joint angles, if it is feasible or leads to a collision.
 
@@
 which however can be done independently for every new \(q\).
-Therefore $I(q)$ is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every thread computes one evaluation.
+Therefore it is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every GPU thread computes one evaluation.
 In the robotics case, this means, the direct kinematic and collision calculation are made on the GPU parallelly for different joint angle configurations.
@@
 
+%%%%%%%%%%%%%%%%%
+
 \section{Implementation of an Indicator Function for Robot Arms}\label{indicator}
 
-In the project, the probabilistic roadmap algorithm is applied  to create collision-free trajectories of a robot arm.
+The probabilistic roadmap algorithm is applied  to create collision-free trajectories of a robot arm.
 For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
@@
 The coordinate frames of the single parts are described by transformation matrices 
-$T_i \in R^{4x4}$, which transform points from the body frame frame into the world frame.
+$T_i \in R^{4x4}$, which transform points from the body frame into the world frame.
 They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
@@
 Applying the corresponding transformation $T_i$, its vertices can be transformed into the world frame.
-Environment obstacles are also defined by convex polytopes, in the world frame.
+Environment obstacles are also defined by convex polytopes, given in the world frame.
 
@@
 
-It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable.
+The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
+Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.
+
+It has to be noted that the trajectories \(q(t)\), the PRM algorithm will generate, are piecewise linear and therefore not differentiable.
 Hence, they have to be smoothed before the use on a robot.
@@
 
-The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
-Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.
-
 \subsection{Direct Kinematics}\label{dk}
 
-The Denavit-Hartenberg (DH) convention is a method to describe the geometry of an robots joints, 
+The Denavit-Hartenberg (DH) convention is a method to describe the geometry of a robots joints, 
 that is, how the transformations between the robots coordinate frames look like in dependency of its joint values $q_1,...,q_d$.
-For this it is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
+A robot consists of $d+1$ bodies each of which has a body fixed coordinate frame $B_i$, which is described by a transformation matrix $T_i$ into the world frame.
+It is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
 or prismatic, a translation along an axis.
-The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by a transformation matrix $T_i$ into the world frame.
-For a robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
-it is now defined, that the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies are
+These joints may now be defined through DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
+by setting the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies to
 \begin{equation}
@@
 where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
-From this follows, that from given DH parameters all transformations are determined.
-The other way round, one can show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by a set of DH parameters.
+With this and $T_0=I$, all transformations $T_i$ are determined.
+
+The other way round, one can now show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by such a set of DH parameters, i.e.
+it exists a set of parameters, such that the above definition leads to the correct transformations for this robot.
 For the exact calculations, see \cite{robodyn}.
@@
 
-For the CUDA implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes float2 for 2D vectors and float4 for 3D vectors.
+For the implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes float2 for 2D vectors and float4 for 3D vectors.
 To guarantee aligned access on the GPU, float4 is used instead of float3.
@@
 
+The second constructor implements the relative DH transformations.
+
 \subsection{Polytopes}
 
-To calculate the intersection between two convex polytopes, the Chung-Wang algorithm only needs their vertices 
-and the data which vertices are connected by an edge. 
-For this the adjacency matrix of the edge graph is stored compactly in the crs format.
+To calculate the intersection between two convex polytopes, the here used Chung-Wang algorithm only needs their vertices 
+and the data, which of them are connected by an edge. 
+For this the adjacency matrix of the edge graph is stored compactly in compact row storage format.
 \begin{lstlisting}
@@
 It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
-The polytopes used in this projects are given by the user as convex hull of a list of vertices.
+The polytopes used in this project are given by the user as convex hull of a list of vertices.
 However, to determine the edge graph only from the vertices is a nontrivial task, which is done by Matlab here using provided high level functions.
@@
 In our application, the polytopes are mostly parts of the robot, which are given in a body frame and have to be transformed into world frame according to section \ref{dk} first.
-However, from the structure of the Chung-Wang algorithm drops out that mostly only a small part of the polytopes vertices are needed for the computations.
-Therefore it would be rather time wasting to transform all vertices before passing to the algorithm.
-To avoid this, all polytopes are passed in their body frames together with their transformations and the algorithm transforms vertices at the time they occur in calculations.
+However, from the structure of the Chung-Wang algorithm drops out, that mostly only a small part of the polytopes vertices are needed for the computations.
+Therefore, it would be rather time wasting to transform all vertices before passing them to the algorithm.
+To avoid this, all polytopes are passed in their body frames together with their transformations and the algorithm transforms a vertex, whenever it occurs in calculations.
+This is also the version proposed in \cite{chung}.
 
@@
 private:
-  const Robot<ndof>* robot; //host
-  Robot<ndof>* robotdev;    //GPU
-  collision4::polytope4data* polydata;            //data on host
-  collision4::polytope4data* polydatadev;         //pointer to GPU obj
+  const Robot<ndof>* robot; //host object
+  Robot<ndof>* robotdev;    //GPU object
+  collision4::polytope4data* polydata;    //host object
+  collision4::polytope4data* polydatadev; //GPU object
   .
@@
 \end{lstlisting}
-where the constant geometry and robot data is passed once at the beginning and copied to the GPU by an init function.
-The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if on the line in between the indicator function is always 0.
+where the constant geometry and robot data is loaded once at the beginning and copied to the GPU by an init function.
+The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
 \begin{equation}
@@
 
-As the PRMSolver only needs to know, which node points can be connected by a line, and not, which points on the line exactly are free, this version delivers exactly the minimal necessary information.
-Furthermore it saves CPU and communication time, because the intermediate points on the lines can be directly computed by the GPU threads and do not need to be exchanged.
+As the PRMSolver only needs to know, which node points can be connected by a line, and not, which intermediate points are free in detail, this version delivers exactly the minimal necessary information.
+Furthermore, it saves CPU and communication time, because the intermediate points on the lines can be directly computed by the GPU threads and do not need to be exchanged.
 
-Algorithms \ref{host} and \ref{kernel} two show in pseudo code, how the work is split between host and device.
+Algorithms \ref{host} and \ref{kernel} show in pseudo code, how the work is split between host and device.
 
@@
 In order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
-Here the function is split in everything until the kernel launch and a wait function, which waits until the kernel finishes and receives the result data.
+Here the function is divided into everything until the kernel launch and a function, which waits, until the kernel has finished and receives the result data.
 
@@
 The implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corresponding waiting calls. 
-With this it is possible to use the full computation time of the GPU.
+With this, it is possible to use the full computation time of the GPU.
 
 \newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{The PRM Solver}\label{solver}
 
-The solver builds in the first phase the roadmap graph from both the start and the end node, until it is connected.
-It is always ensured that the graph consists of two connected components containing the start and the end.
+The solver builds a first phase the roadmap graph from both the start and the end node, until it is connected.
+It is always ensured that the graph consists of two connected components containing the start resp. the end node.
 With this, it can be easily determined at each time, if the whole graph has become connected.
 
-At the end a shortest path from the start to the end node is searched by the Dijkstra algorithm.
-It has shown that this part a neglectable amount of time compared to the graph building phase.
-Therefore it is not parallelized here and the focus lies on building the roadmap.
+At the end, in a second phase, a shortest path from the start to the end node is searched by the Dijkstra algorithm.
+It has shown, that this part needs a neglectable amount of time compared to building the graph.
+Therefore, it is not parallelized here and the focus lies on the first phase.
 
@@
 That is the set of all nodes $\tilde q$ of the graph, with $\norm{\tilde q - q} \le D$.
-If no special order of the nodes is given, one has to calculate the norms for all nodes of the graph, which is rather time consuming.
+If no special order of the nodes is given, one has to calculate this norms for all nodes of the graph, which is rather time consuming.
 
-Therefore a special indexing is used here. 
+Therefore, a special indexing is used here. 
 The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map.
@@
 This has to be chosen so large that enough nodes are stored in one block in order to have efficient memory access.
-On the other side, the smaller it is, the less computations have to be done for the neighborhood requests.
-For illustration the following pseudocode shows the insertion of a node.
+On the other side, the smaller it is, the less unnecessary computations have to be done for the neighborhood requests.
+For illustration of the structure, the following pseudocode shows the insertion of a node.
 
@@
 	}
-	Insert node in block
+	Insert node into block
 	\caption{Inserting a node}
@@
 The work of the single MPI processes could be parallelized further with OpenMP, but this has not been done here.
-Instead only different approaches to distribute the work on several GPUs have been investigated.
+Instead, different approaches to distribute the work over several GPUs have been investigated.
 
@@
 
-In the first approach, that was evaluated, every process generates new nodes independently from each other.
+In the first approach, that has been evaluated, every process generates new nodes independently from each other.
 Then the possible neighbors are determined and the connections, which have to be tested.
-The indicator function is called and returns the edges.
+The indicator function is called and returns the edge data.
 After this, the graph must be synchronized between all processes, which needs an all-to-all communication.
@@
 Here, every process runs two workers overlapped. 
-Here, in the time, when the first worker has generated his indicator function request and launched the kernel and is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round.
+In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round.
 With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
@@
 However, the different versions lead to a different average amount of GPU threads needed.
-So, to be able to better compare how they scale, the time per number of GPU threads used as a quantity.
-Table \ref{resulttable} shows the results.
-One can see, that while the versions with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
-The versions with less communication however have better absolute times.
+Hence, to be able to better compare how they scale, the time per number of GPU threads is used as a quantity.
+Table 1 shows the results.
+One can see, that the version with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
+The versions with less communication, however, have better absolute times.
 Furthermore, one sees the improvement from version 2 to version 3 from the asynchronous indicator function.

### 2016-03-31 prm/run.sh
...

@@ -1,3 +1,4 @@
+#do not call without argument
 #make -j4
 mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 1
 mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 1

### 2016-03-31 prm/matlab/graphplot.m
array example

@@ -1,6 +1,6 @@
 clear all
 
-if 0
+if 1
     A=imread('labyrinth3.jpg');
     A=A(:,:,1);
 
@@ -11,7 +11,7 @@ Ql=binread('../prmoutput/graphl/qstoragec.bin','float');
 Qr=binread('../prmoutput/graphr/qstoragec.bin','float');
 Ql=reshape(Ql,2,[]);
 Qr=reshape(Qr,2,[]);
-if 0
+if 1
 Ql=[h,0;
     0,h]*Ql;
 Qr=[h,0;
@@ -49,7 +49,7 @@ Qpathr=Qr(:,pathr);
 close all;
 pause(0.1);
 
-%image(A);
+image(A);
 hold on;
 
 Gl=graph(Elfrom,Elto);

### 2016-03-31 prm/src/array.cpp

@@ -9,11 +9,11 @@
 #include "lib/tictoc.h"
 
 #include "lib/robotspace/robotconfigspace.h"
+#include "lib/arrayspace/arrayconfigspace.hpp"
 #include "lib/prm/prmsolver.h"
 
 
 using namespace std;
-
 
 
 int main(int argc, char** argv)
@@ -25,7 +25,7 @@ int main(int argc, char** argv)
   MPI_Comm_size(MPI_COMM_WORLD, &size);
 
 
-  const int ndof=4;
+  const int ndof=2;
 
 
   tick(tinit);
@@ -48,31 +48,30 @@ int main(int argc, char** argv)
   //srand(rank+time(NULL));
   srand(seed+rank*10);
   int firstrand=rand();
+  //printvar(firstrand);
 
 
   //! Initialization
 
-  RobotConfigspace<ndof> space("config1",dq, confignbuf);
-  space.init(rank,size);
+
+  int b=640,h=480,n=b*h;
+
+  int *array=new int[n];
+  read_file("array.bin",array,n);
+  ArrayConfigspace space(array,b,h,0.0,b/(float)h,0.0,1.0);
 
   PRMSolver<ndof> prm(&space, D, D, maxstorage, blocksize);
 
 
-  float qstart[ndof];
-  float qend[ndof];
-  for(int dof=0;dof<ndof;++dof){
-    qstart[dof]=0.0;
-    qend[dof]=3;
-  }
+  float qstart[ndof]={0.1f*space.max(0),0.5f*space.max(1)};
+  float qend[ndof]={0.9f*space.max(0),0.5f*space.max(1)};
 
   if(rank==0){
-    printvar(space.indicator(&qstart[0]));
-    printvar(space.indicator(&qend[0]));
+    assert(space.indicator(&qstart[0])==0);
+    assert(space.indicator(&qend[0])==0);
   }
 
   prm.init(&qstart[0],&qend[0]);
-
-  //prm.print();
 
   tock(tinit);
 
@@ -81,22 +80,29 @@ int main(int argc, char** argv)
 
   tick(trun);
 
-
-  //prm.process_mpi(num,nbuf,maxsteps);
-  if(prmversion==2){
+  int version=-1;
+  if(prmversion==1){
+    prm.process_mpi(num,nbuf,maxsteps);
+    version=1;
+  }else if(prmversion==2){
     prm.process_mpi2(num,nbuf,maxsteps, seed);
+    version=-1;
   }else if(prmversion==3){
[…]

### 2016-03-31 prm/src/lib/prm/prmsolver.cpp

@@ -2654,6 +2654,7 @@
 
 
   template class PRMSolver<4>;
+  template class PRMSolver<2>;
 
 
 

### 2016-03-31 prm/run.sh
finale version

@@ -1,12 +1,0 @@
-#do not call without argument
-#make -j4
-mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 1
-mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 1
-mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 1
-mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 4
-mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 4
-mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 4
-mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 5
-mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 5
-mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 5
-

### 2016-03-31 prm/bench.sh
final version

@@ -1,0 +1,12 @@
+#do not call without argument
+#make -j4
+mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 1
+mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 1
+mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 1
+mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 4
+mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 4
+mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 4
+mpirun -n 1 ./maincuda 32 2048 0.01 1.0 $1 256 5
+mpirun -n 2 ./maincuda 32 2048 0.01 1.0 $1 256 5
+mpirun -n 4 ./maincuda 32 2048 0.01 1.0 $1 256 5
+

### 2016-05-31 paper/presentation.tex
presentation

--- 2016-03-30 paper/paper.tex
@@
-\documentclass{scrartcl} 
-%\documentclass[a4paper]{article}
-%\usepackage{a4wide}
+\documentclass[11pt]{beamer}
 
+%% Sprache und Zeichensatz:
+\usepackage[applemac]{inputenc}
+%\usepackage[ngerman]{babel}
+\usepackage[T1]{fontenc}
 \usepackage[ngerman,english]{babel}
- \usepackage[T1]{fontenc}
 \usepackage{lmodern}
-\usepackage[applemac]{inputenc}
 
@@
 \usepackage{amstext}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 
-\usepackage{fancyhdr}
-\usepackage{paralist}
+\usepackage{setspace}
 
-\usepackage{mathtools}
+\usetheme{boxes}
+\usepackage{dsfont}
 
-\usepackage{hyperref}
+\usecolortheme{rose}
+\useinnertheme[shadow]{rounded}
 
-\usepackage{graphicx}
+%---------------------------------------------------------------------------------------------
+%
+%%Pakete Mathematik
+%\usepackage{amsmath, amssymb, amstext, amsfonts, mathrsfs}
+%\usepackage[squaren]{SIunits}
 
-\usepackage{wrapfig}
+%---------------------------------------------------------------------------------------------
+%% Für eine Titelseite:
 
-% PSEUDOCODE
+%%%%%%
 
-\usepackage[ruled]{algorithm2e}
+\usepackage[matrix,arrow]{xy}
+\usepackage{stmaryrd}
 
-% C++ CODE
+\DeclareMathOperator*{\esssup}{ess.sup}
+\DeclareMathOperator*{\defgl}{:=}
+\DeclareMathOperator*{\defgr}{=:}
+\DeclareMathOperator*{\argmax}{argmax}
 
-\usepackage{listings}
-\usepackage{xcolor}
-\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
-
-\lstdefinestyle{customc}{
-  belowcaptionskip=1\baselineskip,
-  breaklines=true,
-  %frame=LR,
-  %xleftmargin=\parindent,
-  language=C,
-  showstringspaces=false,
-  basicstyle=\footnotesize\ttfamily,
-  keywordstyle=\bfseries\color{green!40!black},
-  commentstyle=\itshape\color{purple!40!black},
-  identifierstyle=\color{blue},
-  stringstyle=\color{orange},
-}
-
-\lstdefinestyle{customcpp}{
-	backgroundcolor=\color{lbcolor},
-    	tabsize=4,
-    	language=[GNU]C++,
-        basicstyle=\tiny,%{\scriptsize},
-        upquote=true,
-        %aboveskip={1.5\baselineskip},
-        columns=fixed,
-        showstringspaces=false,
-        extendedchars=false,
-        breaklines=true,
-        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
-        frame=single,
-        numbers=left,
-        showtabs=false,
-        showspaces=false,
-        showstringspaces=false,
-        identifierstyle=\ttfamily,
-        keywordstyle=\color[rgb]{0,0,1},
-        commentstyle=\color[rgb]{0.026,0.112,0.095},
-        stringstyle=\color[rgb]{0.627,0.126,0.941},
-        numberstyle=\color[rgb]{0.205, 0.142, 0.73},
-}
-
-\lstset{escapechar=@,style=customcpp}
+\newcommand{\3}{ ^{3\times3} }
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\M}{\mathbb{M}}
+\newcommand{\Sym}{\mathbb{S}^3}
+\newcommand{\Mpos}{\M_+^3}
+\newcommand{\Orth}{\mathbb{O}^3}
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
+\newcommand{\abs}[1]{\left |#1\right |}
+\newcommand{\ub}{\bar{u}}
+\newcommand{\yb}{\bar{y}}
+\newcommand{\scprod}[2]{\langle#1 , #2 \rangle}
+\newcommand{\E}{\mathcal{E}}
+\newcommand{\F}{\mathcal{F}}
+\newcommand{\G}{\mathcal{G}}
 
@@
 
-\DeclareMathOperator*{\esssup}{ess.sup}
-\DeclareMathOperator*{\defgl}{\vcentcolon=}
+%%%%%%
 
-%\pagestyle{fancy}
-%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
-%\fancyhead[C]{} %zentrierte Kopfzeile
-%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
-\title{\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
+\title{High Performance Computing Project\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration} 
+\author{Manuel Demmeler}
+\institute{Technische Universität München}
 \date{}
-\author{Manuel Demmeler}
 
-%\thispagestyle{fancy}
-
-\newcommand{\3}{ ^{3\times3} }
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\M}{\mathbb{M}}
-\newcommand{\Sym}{\mathbb{S}^3}
-\newcommand{\Mpos}{\M_+^3}
-\newcommand{\Orth}{\mathbb{O}^3}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\newcommand{\abs}[1]{\left |#1\right |}
-\newcommand{\ub}{\bar{u}}
-\newcommand{\yb}{\bar{y}}
-\newcommand{\scprod}[2]{\langle#1 , #2 \rangle}
-
-%\vspace{-3.4cm}
+\AtBeginSection[]{ 
+	\frame {
+		%\frametitle{Gliederung}
+		\tableofcontents[current, currentsubsection]
+	} 
+}
 
 \begin{document}
-\maketitle
 
+\frame{\titlepage}
+
+\begin{frame}
+	\frametitle{Gliederung}
+	\tableofcontents
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%
 \section{Motivation}
-In trajectory optimization with obstacles, it is often necessary to provide feasible initial trajectories. 
-For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive.
-A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form.
-An efficient approximative approach is using probabilistic roadmaps (PRMs).
+\begin{frame}
+\frametitle{Motivation}
+	\begin{center}
+		\includegraphics[height=4cm]{configspacetrafo.png}
+	\end{center}
+	\begin{itemize}
+		\item Problem: find a feasible trajectory of a robot through a setting of obstacles
+		\item often necessary to pass feasible initial tajectories to optimization algorithms
+	\end{itemize}
+\end{frame}
 
-In this project, a parallelized version of the PRM method has been developed in general.
-The robotics application then was implemented as an special case of this with GPU support.
+\newtheorem*{defi}{Definition}
 
-\section{Introduction}
-\subsection{Problem Statement and the Sequential PRM Method}\label{algintro}
-\newtheorem*{defi}{Definition}
-\newtheorem*{alg1}{Algorithm}
+\begin{frame}
+\frametitle{Problem Statement}
+	\begin{itemize}
+		\item Obstacles can not be assumed to have a closed form
+		\item Therefore they are defined by an indicator function
+		\begin{equation*}
+			I: \Omega \subset \R^d \rightarrow \{ 0,1 \}
+		\end{equation*}
+		\item collision for $q \in \Omega$, $I(q) = 1$ 
+	\end{itemize}
 
-The abstract problem is to find a path on a d-dimensional space, which avoids intersecting obstacles. The obstacles are  given by an indicator function
+\end{frame}
+
+\begin{frame}
+\frametitle{PRM Algorithm}
+	PRM: Grow graphs from given start and endpoints $q_s, q_e \in \Omega$ until a connection is found, by repeating
+	\begin{itemize}
+		\item sample new node $v$ randomly in environment of old nodes
+		\item check for all possible neighbors $w$, if the linear connection is feasible, in this case add edge $\{v,w\}$ to the graph
+		\item break, if the two graphs have been connected.
+	\end{itemize}
+	\begin{defi}
+	For two points \(v\) and \(w\) we define the connection with stepsize \(h>0\) as
 		\begin{equation*}
-	I: \Omega \rightarrow \{0, 1\}
+			[v,w]_h \defgl \{ q = \lambda v+(1-\lambda) w\ |\ \lambda \in [0,1], \norm{q-v} \in \N_0 h \}.
 		\end{equation*}
-on the domain \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \) and the goal is to find a continuous way
-\begin{equation*}
-	\Gamma: [0,1] \rightarrow \Omega
-\end{equation*}
-between a given start and endpoint, \(\Gamma(0)=q_b,\ \Gamma(1)=q_e \), 
-such that
-\begin{equation*}
-	I(\Gamma(s))=0 \text{\ f.a.\ } s \in [0,1].
-\end{equation*}
-The probabilistic roadmap (PRM) approach grows a graph \(G=(V,E)\) form \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in V\subset \Omega\).
-Two nodes \(v,w \in V\) are connected, if they are near enough each other and the linear connection in between is free from obstacles.
-The algorithm terminates, when there is a path from \(q_s\) to \(q_e\) on G.
-\begin{defi}
-For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
-	\begin{equation*}
-		[q_1,q_2]_h \defgl \{ q = \lambda q_1+(1-\lambda) q_2\ |\ \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
-	\end{equation*}
-	We say, \(q_1\) and \(q_2\) are connected (with stepsize \(h>0\)) on the map \(\Omega\), if
+		We say, \(q_1\) and \(q_2\) are connected with stepsize \(h>0\), if
 		\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
 	\end{defi}
+\end{frame}
 
-\begin{algorithm} 
-	\caption{Probabilistic Roadmap in general}
-	\KwData{ \(q_s, q_e\in \Omega, h>0, d_0>0\)}
-	\KwResult{ \(q_s=q_1, ... ,q_n=q_e\in \Omega\), 
-		such that all \(q_i\) and \(q_{i+1}\) are connected with stepsize \(h\)}
-	Initialize graph \(G=(V,E)\) with \(V=\{q_s, q_e\}\) \\
-	If \(q_s\) and \(q_e\) are connected, \(E \leftarrow\{ q_s, q_e\}\) \\
-	\While{ there exists no path in G between \(q_s\) and \(q_e\)}{
-		Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\) \\
-		\For{ all nodes \(\tilde{q} \in V\), with \( \norm{q- \tilde q} \le d_0 \) }{
-			\If { $q$ and $\tilde q$ connected with stepsize h} {
-				\(E \leftarrow \{q, \tilde{q}\}\)
-			}
-		}
-		\(V \leftarrow{q}\), if at least one edge was inserted \\
-	}
-	Determine the shortest path \(q_1, ... , q_n\in V\) on \(G\) from \(q_s\) to \(q_e\).
-\end{algorithm}
+%%%%%%%%%%%%%%%%%%%%%%%
+\section{Robotics Application}
+\begin{frame}
+\frametitle{Robotics Application}
+	\begin{center}
+		\includegraphics[height=3cm]{configspacetrafo.png}
+	\end{center}
+	\begin{itemize}
+		\item For each linear connection, every intermediate configuration has to be checked for collisions
+		\item $\Rightarrow$ Many independent and similar kinematics and collision calculations
+		\item $\Rightarrow$ Well suited for usage of GPU
+	\end{itemize}
+\end{frame}
 
-The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\).
-Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.
-
-\subsection{Application to Robotics}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.8\textwidth]{configspacetrafo.png}
-    \caption{Example for an indicator function in robotics from \cite{prmlec}. 
-    		The real obstacles on the left are transformed 
-		into occupied areas of the configuration space on the right.}
-    \label{fig:awesome_image}
-\end{figure}
-
-In the project, the PRM algorithm is used to find trajectories for a robot arm, which avoid self collisions and collisions with obstacles in the environment.
-Such trajectories can be expressed as functions describing the joint variables changing over time, that means curves in the configuration space with dimension equal to the number of degrees of freedom of the robot.
-With this association, the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form stated in the previous chapter.
-The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given set of joints, if the configuration is feasible or a collision occurs.
-
-\newpage
-\subsection{Project Overview}
-
-Like in the stated robotics example, in many cases the most costly operation of the PRM algorithm are the many evaluations of the indicator function \(I(q)\), 
-which however can be done independently for every new \(q\).
-Therefore $I(q)$ is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every thread computes one evaluation.
-In the robotics case, this means, the direct kinematic and collision calculation are made on the GPU parallely for different joint angle configurations.
-Although this will the only major application in the project, the PRM solver and the configuration space are implemented seperated from each other and connected by an interface to be free to simply add other applications.
-As an example and for testing, additionally a simple 2D configuration space is provided,
-whose indicator function is read from a black-and-white image.
-
+\begin{frame}
+\frametitle{Project Overview}
 \begin{center} 
+\resizebox{\textwidth}{!}{%
 \begin{tikzpicture}[auto]%node distance=0.5cm and 1.8cm]
@@
 \end{tikzpicture}
+}
 \end{center}
+	\begin{itemize}
+	\item PRM algorithm and Configuration space independent from each other
+	\item Only connected through indicator function represented by Configspace interface
+	\end{itemize}
+\end{frame}
 
-While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
-a second level can be achieved by sampling and connecting several new nodes parallely. 
-This is more difficult, because all new nodes have to be connected 
-and the graph data has to be updated on all processes. 
-For this, different approaches have been implemented, which are described in chapter \ref{solver}.
-
-Before that, chapter \ref{indicator} gives more details about the robotics indicator function.
-
-\section{Implementation of an Indicator Function for Robot Arms}\label{indicator}
-
-In the project, the probabilistic roadmap algorithm is applied  to create collisionfree trajectories of a robot arm.
-For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
-For this, coordinate frames according to the Denavit-Hartenberg convention are used.
-The coordinate frames of the single parts are described by transformation matrices 
-$T_i \in R^{4x4}$, which transform points from the body frame frame into the world frame.
-They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
-
-The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame.
-Applying the corresponding transformation $T_i$, its vertices can be transformed into the world frame.
-Environment obstacles are also defined by convex polytopes, in the world frame.
-
-Now the movement of the robot can be described by a curve $q(t) \in \R^d$ in the configuration space.
-It is collisionfree, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
-To apply the probabilistic roadmap algorithm, we express this by the indicator function
-
-\begin{equation*}
-	I(q)=
-	\begin{cases}
-		0,	& \text{ for no collision for all given pairs of polytopes } \\
-		1,	& \text{ if at least one collision occurs }
-	\end{cases}
-\end{equation*}
-
-It has to be noted that the trajectories \(q(t)\) the PRM algorithm generates are piecewise linear and therefore not differentiable.
-Hence, they have to be smoothed before the use on a robot.
-This is not part of the project.
-The task is only to generate feasible trajectories to be passed to an optimizer.
-
-The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
-Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.
-
-\subsection{Direct Kinematics}\label{dk}
-
-The Denavit-Hartenberg (DH) convention is a method to describe the geometry of an robots joints, 
-that is, how the transformations between the robots coordinate frames look like in dependancy of its joint values $q_1,...,q_d$.
-For this it is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
-or prismatic, a translation along an axis.
-The robot consists of $d+1$ bodies each of which has an body fixed coordinate frame $B_i$, which is described by a transformation matrix $T_i$ into the world frame.
-For a robot with DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
-it is now defined, that the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies are
-\begin{equation}
-	T_{i-1,i} = R_z(\theta_i) T_z(d_i) R_x(\alpha_{i-1}) R_x(a_{i-1})
-\end{equation}
-where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
-From this follows, that from given DH parameters all transformations are determined.
-The other way round, one can show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by a set of DH parameters.
-For the exact calculations, see \cite{robodyn}.
-
-\subsection{2D and 3D Geometry Library}
-
-For the CUDA implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes float2 for 2D vectors and float4 for 3D vectors.
-To guarantee aligned access on the GPU, float4 is used instead of float3.
-Furthermore, as the geometric calculations made here are numerically not very critical, the usage of double precision has turned out to be not necessary.
-The library contains hybrid host and device inline functions for common vector operations such as
-\begin{lstlisting}
-__host__ __device__ inline float4& operator += (float4& u, const float4& v);
-__host__ __device__ inline float4& operator *=  (float4& u, const float f);
-__host__ __device__ inline float4& add(const float4& u, const float4& v, float4& w);
-\end{lstlisting}
-Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
-
-Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
-\begin{lstlisting}
-class trafo4{
-  public:
-    float4 col[4];
-    __host__ __device__ trafo4(){}
-    __host__ __device__ trafo4(float a, float alpha, float q, float d);
-    __host__ __device__ inline float4& apply(float4& u) const;
-    __host__ __device__ inline float4& apply(const float4& u, float4& Tu) const;
-    .
-    .
-    .
-}
-\end{lstlisting}
-
-\subsection{Polytopes}
-
-To calculate the intersection between two convex polytopes, the Chung-Wang algorithm only needs their vertices 
-and the data which vertices are connected by an edge. 
-For this the adjacency matrix of the edge graph is stored compactly in the crs format.
-\begin{lstlisting}
-struct polytope4{
-    float4* vertices;	//length n
-    int n;
-    //edges
-    int* dsp;	//length n
-    int* cnt;	//length n
-    int* dest;	//length m
-    int m;
-};
-\end{lstlisting}
-It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
-The polytopes used in this projects are given by the user as convex hull of a list of vertices.
-However, to determine the edge graph only from the vertices is a nontrivial task, which is done by Matlab here using provided high level functions.
-
-\subsection{Chung-Wang Collision Algorithm}
-
-To check if two polytopes collide an algorithm by Chung and Wang is used here.
-It uses the separating vector theorem and tries to find a seperating vector by iterative search.
-Furthermore, a subalgorithm determines in each iteration from the candidate history, if it is still theoretically possible to find a seperating vector.
-If not, it is returned a collision.
-If on the other side a seperating vector is found, the algorithm can return no collision.
-For detailed informations see \cite{chung} and \cite{ericson}, pp. 410-412.
-The method is implemented as device code for the use on GPU kernels.
-\begin{lstlisting}
-__host__ __device__ int seperating_vector_algorithm(const polytope4& P, const polytope4& Q, 
-													const trafo4& tp, const trafo4& tq);
-\end{lstlisting}
-In our application, the polytopes are mostly parts of the robot, which are given in a body frame and have to be transformed into world frame according to section \ref{dk} first.
-However, from the structure of the Chung-Wang algorithm drops out that mostly only a small part of the polytopes vertices are needed for the computations.
-Therefore it would be rather time wasting to transform all vertices before passing to the algorithm.
-To avoid this, all polytopes are passed in their body frames together with their transformations and the algorithm transforms vertices at the time they occur in calculations.
-
-\subsection{Robot Kernel}
-
-The robot configuration space is implemented in the form
-\begin{lstlisting}
-template<int ndof>
-class RobotConfigspace : public Configspace<ndof>
-{
-public:
-  RobotConfigspace(const Robot<ndof>* robot_,
-                   const polytope *polys_,
-                   const int* sys_,
-                   const int N_,
-                   const int *from_, const int *to_,
-                   const int M_,
-                   const float* mins_, const float* maxs_, const float dq_,
-                   const int nbuf_);
-  int init(const int ressource_rank=0, const int ressource_size=1);
-  int indicator2(const float* qs, const float* qe, int *res, const int N, const int offset);
-  .
-  .
-  .
-private:
-  const Robot<ndof>* robot; //host
-  Robot<ndof>* robotdev;    //GPU
-  collision4::polytope4data* polydata;            //data on host
-  collision4::polytope4data* polydatadev;         //pointer to GPU obj
-  .
-  .
-  .
-}
-\end{lstlisting}
-\begin{lstlisting}
-__global__ void kernel_indicator2(const Robot<ndof>* robot,
-                                  const collision4::polytope4data* polydata,
-                                  const float* qs, int offsets,
-                                  const float* qe, int offsete,
-                                  int* res,
-                                  const int* testpos, const int* testnum,
-                                  int N, int numthreads);
-\end{lstlisting}
-where the constant geometry and robot data is passed once at the beginning and copied to the GPU by an init function.
-The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if on the line in between the indicator function is always 0.
-\begin{equation}
-	res[j]=\begin{cases}
-		0	&\text{if $I(q)=0$ f.a. $q \in \left[qs[j], qe[j]\right]_{\Delta q}$} \\
-		1	&else
-	\end{cases}
-\end{equation}
-The border points are stored in structures of arrays.
-\begin{equation}
-\begin{aligned}
-	q_{start}^j &= \left(qs[j],qs[j+\text{offset}],...,qs[j+(d-1)*\text{offset}]\right)^T\\
-	q_{end}^j &= \left(qe[j],qe[j+\text{offset}],...,qe[j+(d-1)*\text{offset}]\right)^T
-\end{aligned}
-\end{equation}
-
-As the PRMSolver only needs to know, which node points can be connected by a line, and not, which points on the line exactly are free, this version delivers exactly the minimal necessary information.
-Furthermore it saves CPU and communication time, because the intermediate points on the lines can be directly computed by the GPU threads and do not need to be exchanged.
-
-Algorithms \ref{host} and \ref{kernel} two show in pseudo code, how the work is split between host and device.
-
-\begin{algorithm}\label{host}
-	\KwData{qs, qe, N, offset}
-	\KwResult{res}
-	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
-	testnum[j] = $d_j/{\Delta q}$: number of threads for each line \;
-	testpos[j] = thread displacements \;
-	reset res \;
-	call kernel \;
-	\caption{Host indicator function}
-\end{algorithm}
-
-\begin{algorithm}\label{kernel}
-	\KwData{arrays qs, qe, testpos, testnum and threadindex}
-	\KwResult{res}
-	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
-	$c = (threadindex - testpos[j]) / (testnum[k]-1) $ \;
-	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
-	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
-	$res_{tmp} = 0; $ \;
-	\For{all registered pairs pairs of polytopes $P, Q$}{
-		$res_{tmp}=seperating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
-		\If{$res_{tmp} \neq 0$}{
-			$res[j]=res_{tmp}$ \;
-			break \;
-		}
-	}
-	\caption{Robot kernel: code for one thread}
-\end{algorithm}
-
-In order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
-Here the function is split in everything until the kernel launch and a wait function, which waits until the kernel finishes and recieves the result data.
-
-\begin{lstlisting}
-int indicator2_async(const float* qs, const float* qe, int *res, 
-					 const int N, const int offset, int &request);
-int indicator2_async_wait(int request);
-\end{lstlisting}
-
-The implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corrsponding waiting calls. 
-With this it is possible to use the full computation time of the GPU.
-
-\newpage
-\section{The PRM Solver}\label{solver}
-
-The solver builds in the first phase the roadmap graph from both the start and the end node, until it is connected.
-It is always ensured that the graph consists of two connected components containing the start and the end.
-With this, it can be easily determined at each time, if the whole graph has become connected.
-
-At the end a shortest path from the start to the end node is searched by the Dijkstra algorithm.
-It has shown that this part a neglectible amount of time compared to the graph building phase.
-Therefore it is not parallelized here and the focus lies on building the roadmap.
-
-\subsection{Storing the Graph}
-
-Apart from the indicator function evaluation, the main problem in building the probabilistic roadmap is, that for each new node $q$ it have to be determined all possible neighbors.
-That is the set of all nodes $\tilde q$ of the graph, with $\norm{\tilde q - q} \le D$.
-If no special order of the nodes is given, one has to calculate the norms for all nodes of the graph, which is rather time consuming.
-
-Therefore a special indexing is used here. 
-The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map.
-
-\begin{lstlisting}
-struct block{
-	int pos;	 //! position of first vertex
-	int num;	 //! current number of vertices stored
-	block* next; //! if num==blocksize -> pointer to next block
-};
-\end{lstlisting}
-
-\begin{lstlisting}
-struct graph{
-    std::map<int,block*> map; //map for accessing blocks
-    std::vector<block> blocks;
-    std::vector<float> qstorage; //length ndof*N
-    int newblockpos;  //position of next new block
-    int blocknum;     //number of used blocks
-    .
-    .
-};
-\end{lstlisting}
-
-The key of an block is computed by the mapping
-\begin{equation*}
-	key(q) = \left\lfloor \frac{q_1}{H} \right\rfloor
-\end{equation*}
-such that $H>0$ becomes a parameter to adjust the grid density.
-This has to be chosen so large that enough nodes are stored in one block in order to have efficient memory access.
-On the other side, the smaller it is, the less computations have to be done for the neighborhood requests.
-For illustration the following pseudocode shows the insertion of a node.
-
-\begin{algorithm}[H]
-	\KwData{$q\in\R^d$}
-	Compute key=key(q)\\ 
-	\eIf{key exists in map}{
-		block=map[key]\\ 
-		\While{ block full }{
-			block=block->next
-		}
-	}{
-		insert new block at map[key]
-	}
-	Insert node in block
-	\caption{Inserting a node}
-\end{algorithm}
-
-For getting the neighbors of a node $q$ it is then enough to look at the keys from $key(q-cD)$ to $key(q+cD)$. For $D=H$, $c=1$ is possible.
-The key mapping could be generalized by taking more components of $q$ into account, but that was not done here.
-
-\subsection{Parallelization Models}
-
-The parallelization is done with MPI. All processes store their own versions of the graph, which are hold up to date between each other.
-Furthermore, it is assumed that every process has its own GPU.
-The work of the single MPI processes could be parallelized further with OpenMP, but this has not been done here.
-Instead only different approaches to distribute the work on several GPUs have been investigated.
-
-\subsubsection*{Version 1}
-
-In the first approach, that was evaluated, every process generates new nodes independantly from each other.
-Then the possible neighbours are determined and the connections, which have to be tested.
-The indicator function is called and returnes the edges.
-After this, the graph must be synchronized between all processes, which needs an all-to-all communication.
-As the graph building phase does not need the edge data, it is sufficient to exchange only the nodes and collect the edges at the end.
+\begin{frame}
+\frametitle{Second Level of Parallelization}
+	Version 1:
+	\begin{itemize}
+	\item Use multiple GPUs parallely
+	\item Sample and connect new nodes on each processor
+	\item Exchange nodes and edges
+	\end{itemize}
 
 \begin{center}
+\resizebox{\textwidth}{!}{%
 \begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
@@
 \end{tikzpicture}
+}
 \end{center}
 
-\subsubsection*{Version 2}
-The second approach uses the fact, that by synchronizing the random seed, 
-different processes can produce the same sequence of random numbers independently from each other.
-By using this, every process can generate all new nodes and determine the connection data. 
-Then the indicator function computation is splitted over their configuration space instances and the returned data shared afterwards with each other.
-With this data, all graphs can be updated and it is guaranteed, that they grow exactly in the same way on each process.
-The purpose of this version compared to the first one is, that less data has to be sent. 
-Furthermore, in the first version it can occur, that the different generated nodes can result in very different counts of neighbors and therefore uneven loads on the GPUs.
-Here, the distribution of the computation amount over the GPUs can be chosen almost freely.
+\end{frame}
 
+\begin{frame}
+\frametitle{Second Level of Parallelization}
+	Version 2:
+	\begin{itemize}
+		\item Grow whole graphs on every processor
+		\item Exchange only edges
+	\end{itemize}
 \begin{center}
+\resizebox{\textwidth}{!}{%
 \begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
@@
 \end{tikzpicture}
+}
 \end{center}
+\end{frame}
 
-\subsubsection*{Version 3}
+\begin{frame}
+\frametitle{Implementation and Testing}
+	\begin{itemize}
+	\item GPU imlementation with CUDA
+	\item CPU Parallelization with MPI
+	\item Tested with a 4 axis robot scenario created with CAD/Matlab
+	\end{itemize}
+	\begin{center}
+	\includegraphics[height=4cm]{robotsolution.pdf}
+	\end{center}
+\end{frame}
 
-The third version does a further optimization by using the asynchronous version of the indicator function.
-Here, every process runs two workers overlapped.
-Here, in the time, when the first worker has generated his indicator function request and launched the kernel and is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round.
-With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
-
-\begin{center}
-\begin{tikzpicture}
-	\tikzset{mywidth/.style={minimum width=2.5cm},};
-	\node[block, mywidth] (l1) {\textbf{Worker 1}};
-	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
-	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
-										generate\\ nodes, \\
-										start kernel
-									\end{tabular}};
-	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
-	\node[frameless, mywidth, below=of l6] (l7) {};
-	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
-										recieve\\ result, \\
-										store nodes \\ 
-										and edges
-									\end{tabular}};
-	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};
-
-	\node[block, mywidth, right=of l1] (r1) {\textbf{Worker 2}};
-	\node[frameless, mywidth, right=of l2, minimum height=1.5cm] (r2) {$\vdots$};
-	\node[frameless, mywidth, right=of l3] (r3) {};
-	\node[block, mywidth, right=of l6] (r6)	{\begin{tabular}{c}
-										recieve\\ result, \\
-										store nodes \\
-										and edges
-									\end{tabular}};
-	\node[block, mywidth, right=of l7] (r7) 	{\begin{tabular}{c}
-										generate\\ nodes, \\
-										start kernel
-									\end{tabular}};
-	\node[frameless, mywidth, right=of l12] (r12) {$\vdots$};
-
-	\node[frameless, left=of l1] (g1) {};
-	\node[block, below=of g1, minimum height=14cm] (graph) {\rotatebox{90}{\textbf{Graph}}};
-
-	%gpu column
-	\tikzset{gpuwidth/.style={minimum width=0.8cm},};
-
-	\node[block, right=2cm of r1, minimum width=2.5cm] {\textbf{GPU}};
-	\node[frameless, right=2cm of r2, gpuwidth] (topa) {$\vdots$};
-	\node[frameless, right=of topa, gpuwidth] (topb) {$\vdots$};
-	\node[block2, below=0.7cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
-	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=4.2cm] (stream1b) {\rotatebox{90}{kernel}};
-	\node[frameless, below=0.3cm of stream1b, gpuwidth] (bota) {$\vdots$};
-
-	\node[block2, below=2.2cm of topb, gpuwidth, minimum height=7.6cm] (stream2) {\rotatebox{90}{kernel}};
-	\node[frameless, right=of bota, gpuwidth] (botb) {$\vdots$};
-
-	%EDGES
-
-	\path[-latex',thick]	(graph.east |- l3.west) edge (l3.west)
-					(l11.west) edge (graph.east |- l11.west);
-	\path[-latex',thick]	(graph.east |- r7.west) edge (r7.west)
-					(r6.west) edge (graph.east |- r6.west);
-	\path[dotted,thick]	(l2) edge (l3)
-					(l3) edge (l11)
-					(l11) edge (l12)
-					(r2) edge (r6)
-					(r6) edge (r7)
-					(r7) edge (r12);
-
-	\path[-latex',thick]	(r6.east -| stream1b.west) edge (r6.east);
-	\path[-latex',thick]	(r7.east) edge (r7.east -| stream1b.west);
-
-	\path[-,thick]		(l3.east) edge (l3.east -| stream1a.west);
-	\path[-latex',thick]	(stream1a.east |- l3.east) edge (l3.east -| stream2.west);
-	\path[latex'-,thick]		(l11.east) edge (l11.east -| stream1b.west);
-	\path[-,thick]		(stream1b.east |- l11.east) edge (l11.east -| stream2.west);
-
-\end{tikzpicture}
-\end{center}
-
-\newpage
-\section{Results}
-
-\begin{wrapfigure}{r}{0.5\textwidth}\label{robotsolution}
-	\centering
-	\includegraphics[width=0.5\textwidth]{robotsolution.pdf}
-	\caption{PRM generated collisionfree movement for a 4-axis robot}
-\end{wrapfigure}
-
-Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
-Because of the dependancy on random numbers, the algorithm takes different amounts of time for different runs.
-Therefore, 10 runs were made for each version and number of processes.
-However, the different versions lead to a different average amount of GPU threads needed.
-So, to be able to better compare how they scale, the time per number of GPU threads used as a quantity.
-Table \ref{resulttable} shows the results.
-One can see, that while the versions with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
-The versions with less communication however have better absolute times.
-Furthermore, one sees the improvement from version 2 to version 3 from the asynchronous indicator function.
-
-\begin{table}[h]\label{resulttable}
-\centering
-\begin{tabular}{| c | c | r | r | r |}
-
-\hline
-version	&  processes/	& time (ms)	& GPU 		& time per 1e6  \\ 
-		& GPUs		& 			& threads		& GPU threads (ms) \\ 
-\hline
-		& 1			& 1002		& 696413		& 1448	\\ 
-1		& 2			& 763		& 838938		& 918	\\ 
-		& 4			& 706		& 1113781	& 647	\\ 
-\hline
-		& 1			& 900		& 608172		& 1490	\\ 
-2		& 2			& 647		& 608172		& 1070	\\ 
-		& 4			& 568		& 608172		& 947	\\ 
-\hline
-		& 1			& 696		& 599847		& 1181	\\ 		
-3		& 2			& 554		& 599847		& 951	\\ 
-		& 4			& 500		& 599847		& 864	\\ 
-\hline
-
-\end{tabular}
-\caption{Timing results for the parallelization versions from chapter \ref{solver} on a machine with 4 GeForce GTX 480}
-\end{table}
-
-As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
-In most cases, the GPU code was around 10 times faster. However, the robot code is not optimized for the CPU version, such that this comparison is not quite fair.
-
-\newpage
-
+\begin{frame}
+\frametitle{Literatur}
 \bibliographystyle{alpha}
@@
 \end{thebibliography}
+\end{frame}
 

### 2016-06-01 prm/src/main.cpp
commandline arguments

@@ -2,6 +2,13 @@
 #include <time.h>
 #include <mpi.h>
 
+//for parsing parameters
+#include <unistd.h>
+#include <stdio.h>
+#include <stdlib.h>
+#include <getopt.h>
+
+
 #define MPI_CODE
 
 #include "lib/config.h"
@@ -30,7 +37,7 @@ int main(int argc, char** argv)
   tick(tinit);
 
   //! Parameters
-
+#if 0
   int num=(argc>=10 ? atoi(argv[9]) : 32 );
   int nbuf=(argc>=3 ? atoi(argv[2]) : 2048);
   int confignbuf=(argc>=3 ? 2*atoi(argv[2]) : 4096);
@@ -44,6 +51,61 @@ int main(int argc, char** argv)
   int maxsteps=10000;//(argc>=2 ? atoi(argv[1]) : 100000 );
 
   bool new_kernel=(argc>=2 ? atoi(argv[1]) : 1 );
+#else
+  int num=32;
+  int nbuf=2048;
+  float dq=0.01;
+  float D=1.0;
+  int seed=0;
+  int blocksize=256;
+  int prmversion=5;
+  int store=1;
+  int maxstorage=1024*1024;
+  int maxsteps=10000;//(argc>=2 ? atoi(argv[1]) : 100000 );
+
+  bool new_kernel=(argc>=2 ? atoi(argv[1]) : 1 );
+#endif
+
+  //! Parse inputs
+  static struct option long_options[] = {
+      {"num",      required_argument,        0,  'n' },
+      {"nbuf", required_argument,            0,  'b' },
+      {"dq",    required_argument,           0,  'q' },
+      {"seed",   required_argument,          0,  's' },
+      {"prmversion", required_argument,      0,  'v' },
+      {"store",      required_argument,      0,  'w' },
+      {"maxsteps",   required_argument,      0,  'N' },
+      {0,           0,                 0,        0   }
+  };
+
+  int opt= 0;
+  int long_index =0;
+  while ((opt = getopt_long(argc, argv,"apl:b:",
+                            long_options, &long_index )) != -1) {
+      switch (opt) {
+      case 'n' : num = atoi(optarg);
+          break;
+      case 'b' : nbuf = atoi(optarg);
+          break;
+      case 'q' : dq = atof(optarg);
+          break;
+      case 's' : seed = atoi(optarg);
+          break;
+      case 'v' : prmversion = atoi(optarg);
+          break;
[…]

### 2016-06-01 prm/src/main.cpp
commandline arguments

@@ -75,6 +75,7 @@ int main(int argc, char** argv)
       {"prmversion", required_argument,      0,  'v' },
       {"store",      required_argument,      0,  'w' },
       {"maxsteps",   required_argument,      0,  'N' },
+      {"new_kernel", required_argument,      0,  'k' },
       {0,           0,                 0,        0   }
   };
 
@@ -96,6 +97,8 @@ int main(int argc, char** argv)
       case 'w' : store = atoi(optarg);
           break;
       case 'N' : maxsteps = atoi(optarg);
+          break;
+      case 'k': new_kernel = atoi(optarg);
           break;
       default:
           exit(EXIT_FAILURE);

### 2016-06-01 prm/src/main.cpp
..

@@ -101,6 +101,7 @@ int main(int argc, char** argv)
       case 'k': new_kernel = atoi(optarg);
           break;
       default:
+          cout<<"bad argument: "<< opt <<endl;
           exit(EXIT_FAILURE);
       }
   }

### 2016-06-01 prm/src/main.cpp
..

@@ -79,7 +79,7 @@ int main(int argc, char** argv)
       {0,           0,                 0,        0   }
   };
 
-  int opt= 0;
+  char opt= 0;
   int long_index =0;
   while ((opt = getopt_long(argc, argv,"apl:b:",
                             long_options, &long_index )) != -1) {

### 2016-06-01 prm/src/main.cpp
..

@@ -81,7 +81,7 @@ int main(int argc, char** argv)
 
   char opt= 0;
   int long_index =0;
-  while ((opt = getopt_long(argc, argv,"apl:b:",
+  while ((opt = getopt_long(argc, argv,"nbqsvwNk",
                             long_options, &long_index )) != -1) {
       switch (opt) {
       case 'n' : num = atoi(optarg);

### 2016-06-01 prm/src/main.cpp
..

@@ -81,7 +81,7 @@ int main(int argc, char** argv)
 
   char opt= 0;
   int long_index =0;
-  while ((opt = getopt_long(argc, argv,"nbqsvwNk",
+  while ((opt = getopt_long(argc, argv,"",
                             long_options, &long_index )) != -1) {
       switch (opt) {
       case 'n' : num = atoi(optarg);

### 2016-06-01 paper/presentation.tex
presentation

@@
 
-\begin{frame}
-	\frametitle{Gliederung}
-	\tableofcontents
-\end{frame}
+%\begin{frame}
+%	\frametitle{Gliederung}
+%	\tableofcontents
+%\end{frame}
 
 %%%%%%%%%%%%%%%%%%%%%%%
-\section{Motivation}
+%\section{Motivation}
 \begin{frame}
@@
 %%%%%%%%%%%%%%%%%%%%%%%
-\section{Robotics Application}
+%\section{Robotics Application}
+
 \begin{frame}
@@
 		\item $\Rightarrow$ Well suited for usage of GPU
+		\item Every GPU thread can make one intermediate test
 	\end{itemize}
 \end{frame}
+
+%\section{Parallelization}
 
@@
 \begin{frame}
+\frametitle{Second Level of Parallelization}
+Version 3: Asynchronous, overlapping kernel calls
+
+\begin{center}
+\resizebox{!}{0.8\textheight}{%
+\begin{tikzpicture}
+	\tikzset{mywidth/.style={minimum width=2.5cm},};
+	\node[block, mywidth] (l1) {\textbf{Worker 1}};
+	\node[frameless, mywidth, below=of l1, minimum height=1.5cm] (l2) {$\vdots$};
+	\node[block, mywidth, below=of l2] (l3) 	{\begin{tabular}{c}
+										generate\\ nodes, \\
+										start kernel
+									\end{tabular}};
+	\node[frameless, mywidth, below=of l3, minimum height=2cm] (l6) {};
+	\node[frameless, mywidth, below=of l6] (l7) {};
+	\node[block, mywidth, below=of l7] (l11)	{\begin{tabular}{c}
+										receive\\ result, \\
+										store nodes \\ 
+										and edges
+									\end{tabular}};
+	\node[frameless, mywidth, below=of l11] (l12) {$\vdots$};
+
+	\node[block, mywidth, right=of l1] (r1) {\textbf{Worker 2}};
+	\node[frameless, mywidth, right=of l2, minimum height=1.5cm] (r2) {$\vdots$};
+	\node[frameless, mywidth, right=of l3] (r3) {};
+	\node[block, mywidth, right=of l6] (r6)	{\begin{tabular}{c}
+										receive\\ result, \\
+										store nodes \\
+										and edges
+									\end{tabular}};
+	\node[block, mywidth, right=of l7] (r7) 	{\begin{tabular}{c}
+										generate\\ nodes, \\
+										start kernel
+									\end{tabular}};
+	\node[frameless, mywidth, right=of l12] (r12) {$\vdots$};
+
+	\node[frameless, left=of l1] (g1) {};
+	\node[block, below=of g1, minimum height=14cm] (graph) {\rotatebox{90}{\textbf{Graph}}};
+
+	%gpu column
+	\tikzset{gpuwidth/.style={minimum width=0.8cm},};
+
+	\node[block, right=2cm of r1, minimum width=2.5cm] {\textbf{GPU}};
+	\node[frameless, right=2cm of r2, gpuwidth] (topa) {$\vdots$};
+	\node[frameless, right=of topa, gpuwidth] (topb) {$\vdots$};
+	\node[block2, below=0.7cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
+	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=4.2cm] (stream1b) {\rotatebox{90}{kernel}};
+	\node[frameless, below=0.3cm of stream1b, gpuwidth] (bota) {$\vdots$};
+
+	\node[block2, below=2.2cm of topb, gpuwidth, minimum height=7.6cm] (stream2) {\rotatebox{90}{kernel}};
+	\node[frameless, right=of bota, gpuwidth] (botb) {$\vdots$};
+
+	%EDGES
+
+	\path[-latex',thick]	(graph.east |- l3.west) edge (l3.west)
+					(l11.west) edge (graph.east |- l11.west);
+	\path[-latex',thick]	(graph.east |- r7.west) edge (r7.west)
+					(r6.west) edge (graph.east |- r6.west);
+	\path[dotted,thick]	(l2) edge (l3)
+					(l3) edge (l11)
+					(l11) edge (l12)
+					(r2) edge (r6)
+					(r6) edge (r7)
+					(r7) edge (r12);
+
+	\path[-latex',thick]	(r6.east -| stream1b.west) edge (r6.east);
+	\path[-latex',thick]	(r7.east) edge (r7.east -| stream1b.west);
+
+	\path[-,thick]		(l3.east) edge (l3.east -| stream1a.west);
+	\path[-latex',thick]	(stream1a.east |- l3.east) edge (l3.east -| stream2.west);
+	\path[latex'-,thick]		(l11.east) edge (l11.east -| stream1b.west);
+	\path[-,thick]		(stream1b.east |- l11.east) edge (l11.east -| stream2.west);
+
+\end{tikzpicture}
+}
+\end{center}
+
+\end{frame}
+
+\begin{frame}
 \frametitle{Implementation and Testing}

### 2016-06-01 paper/presentation.tex
presentation

@@
 		\item Problem: find a feasible trajectory of a robot through a setting of obstacles
-		\item often necessary to pass feasible initial tajectories to optimization algorithms
+		\item Often necessary to pass feasible initial tajectories to optimization algorithms
 	\end{itemize}
@@
 		\end{equation*}
-		\item collision for $q \in \Omega$, $I(q) = 1$ 
+		\item Collision for $q \in \Omega$, $I(q) = 1$ 
 	\end{itemize}
@@
 		\item Every GPU thread can make one intermediate test
+	\end{itemize}
+\end{frame}
+
+%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{frame}
+\frametitle{Indicator Function Implementation}
+\begin{itemize}
+	\item Input: vectors of start and end nodes for each potential edge
+	\item Output: vector of feasible edges
+	\item CPU determines number of threads needed for each edge
+		 and passes nodes to GPU 
+	\item Each CUDA thread determines his intermediate configuration and checks for collsion
+	\begin{itemize}
+		\item Kinematics with Denavit-Hartenberg transformations
+		\item Geometry parts are convex polytopes
+		\item Collision detection: Chung-Wang Collision Algorithm
+	\end{itemize}
 \end{itemize}

### 2016-06-29 paper/paper.tex
paper journal format angefangen

@@
-\documentclass{scrartcl} 
+\documentclass[twocolumn]{svjour3} 
 %\documentclass[a4paper]{article}
@@
 \usepackage{amstext}
-\usepackage{amsthm}
+%\usepackage{amsthm}
 
-\usepackage{fancyhdr}
+%\usepackage{fancyhdr}
 \usepackage{paralist}
@@
 \usepackage{listings}
-\usepackage{xcolor}
-\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
+\lstset{breaklines=true}
 
-\lstdefinestyle{customc}{
-  belowcaptionskip=1\baselineskip,
-  breaklines=true,
-  %frame=LR,
-  %xleftmargin=\parindent,
-  language=C,
-  showstringspaces=false,
-  basicstyle=\footnotesize\ttfamily,
-  keywordstyle=\bfseries\color{green!40!black},
-  commentstyle=\itshape\color{purple!40!black},
-  identifierstyle=\color{blue},
-  stringstyle=\color{orange},
-}
-
-\lstdefinestyle{customcpp}{
-	backgroundcolor=\color{lbcolor},
-    	tabsize=4,
-    	language=[GNU]C++,
-        basicstyle=\tiny,%{\scriptsize},
-        upquote=true,
-        %aboveskip={1.5\baselineskip},
-        columns=fixed,
-        showstringspaces=false,
-        extendedchars=false,
-        breaklines=true,
-        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
-        frame=single,
-        numbers=left,
-        showtabs=false,
-        showspaces=false,
-        showstringspaces=false,
-        identifierstyle=\ttfamily,
-        keywordstyle=\color[rgb]{0,0,1},
-        commentstyle=\color[rgb]{0.026,0.112,0.095},
-        stringstyle=\color[rgb]{0.627,0.126,0.941},
-        numberstyle=\color[rgb]{0.205, 0.142, 0.73},
-}
-
-\lstset{escapechar=@,style=customcpp}
+%\usepackage{xcolor}
+%\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
+%
+%\lstdefinestyle{customc}{
+%  belowcaptionskip=1\baselineskip,
+%  breaklines=true,
+%  %frame=LR,
+%  %xleftmargin=\parindent,
+%  language=C,
+%  showstringspaces=false,
+%  basicstyle=\footnotesize\ttfamily,
+%  keywordstyle=\bfseries\color{green!40!black},
+%  commentstyle=\itshape\color{purple!40!black},
+%  identifierstyle=\color{blue},
+%  stringstyle=\color{orange},
+%}
+%
+%\lstdefinestyle{customcpp}{
+%	backgroundcolor=\color{lbcolor},
+%    	tabsize=4,
+%    	language=[GNU]C++,
+%        basicstyle=\tiny,%{\scriptsize},
+%        upquote=true,
+%        %aboveskip={1.5\baselineskip},
+%        columns=fixed,
+%        showstringspaces=false,
+%        extendedchars=false,
+%        breaklines=true,
+%        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
+%        frame=single,
+%        numbers=left,
+%        showtabs=false,
+%        showspaces=false,
+%        showstringspaces=false,
+%        identifierstyle=\ttfamily,
+%        keywordstyle=\color[rgb]{0,0,1},
+%        commentstyle=\color[rgb]{0.026,0.112,0.095},
+%        stringstyle=\color[rgb]{0.627,0.126,0.941},
+%        numberstyle=\color[rgb]{0.205, 0.142, 0.73},
+%}
+%
+%\lstset{escapechar=@,style=customcpp}
+%
 
@@
 %\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
-\title{\vspace{-1.3cm} \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
+\title{ \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
 \date{}
@@
 \subsection{Problem Statement and the Sequential PRM Method}\label{algintro}
-\newtheorem*{defi}{Definition}
-\newtheorem*{alg1}{Algorithm}
+%\newtheorem*{defi}{Definition}
+%\newtheorem*{alg1}{Algorithm}
 
@@
 The algorithm terminates, when there exists a path from \(q_s\) to \(q_e\) on G.
-\begin{defi}
+\begin{definition}
 For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
@@
 	\(I(q)=0\) for all \(q \in [q_1, q_2]_h\).
-\end{defi}
+\end{definition}
 
@@
 
-\begin{figure}
+\begin{figure*}
     \centering
@@
     \label{fig:awesome_image}
-\end{figure}
+\end{figure*}
 
@@
 
+\begin{figure*}
 \begin{center} 
@@
 \end{center}
+\end{figure*}
 
@@
 The library contains hybrid host and device inline functions for common vector operations such as
-\begin{lstlisting}
+
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*]
 __host__ __device__ inline float4& operator += (float4& u, const float4& v);
@@
 \end{lstlisting}
+\normalsize
+
 Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
@@
 Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
-\begin{lstlisting}
+
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*]
 class trafo4{
@@
 For this the adjacency matrix of the edge graph is stored compactly in compact row storage format.
-\begin{lstlisting}
+
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*]
 struct polytope4{
@@
 \end{lstlisting}
+
 It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
@@
 The method is implemented as device code for the use on GPU kernels.
-\begin{lstlisting}
-__host__ __device__ int separating_vector_algorithm(const polytope4& P, const polytope4& Q, 
-													const trafo4& tp, const trafo4& tq);
+
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*]
+__host__ __device__ int separating_vector_algorithm(const polytope4& P, const polytope4& Q, const trafo4& tp, const trafo4& tq);
 \end{lstlisting}
+
 In our application, the polytopes are mostly parts of the robot, which are given in a body frame and have to be transformed into world frame according to section \ref{dk} first.
@@
 The robot configuration space is implemented in the form
-\begin{lstlisting}
+
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*]
 template<int ndof>
@@
 \end{lstlisting}
-\begin{lstlisting}
+
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*]
 __global__ void kernel_indicator2(const Robot<ndof>* robot,
@@
 \end{lstlisting}
+
 where the constant geometry and robot data is loaded once at the beginning and copied to the GPU by an init function.
 The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
+
 \begin{equation}
@@
 \end{equation}
+
 The border points are stored in structures of arrays.
+
 \begin{equation}
@@
 
-\begin{algorithm}[H]
+\begin{algorithm}
 	\KwData{$q\in\R^d$}
@@
 
+\begin{figure*}
 \begin{center}
@@
 \end{center}
+\caption{test}
+\end{figure*}
 
@@
 
-\begin{wrapfigure}{r}{0.5\textwidth}\label{robotsolution}
+\begin{figure}%{r}{0.5\textwidth}
+\label{robotsolution}
 	\centering
@@
 	\caption{PRM generated collision-free movement for a 4-axis robot}
-\end{wrapfigure}
+\end{figure}
 

### 2016-07-12 paper/paper.tex
paper anpassungen

@@
 
-\usepackage[ngerman,english]{babel}
+\usepackage[english]{babel}
+%\usepackage[english,ngerman]{babel}
+
 \usepackage[T1]{fontenc}
-\usepackage{lmodern}
+%\usepackage{lmodern}
 \usepackage[applemac]{inputenc}
+
+\tolerance=250
 
@@
 
-\usepackage{hyperref}
+%\usepackage{hyperref}
 
-\usepackage{graphicx}
-
-\usepackage{wrapfig}
+%\usepackage{graphicx}
+%\usepackage{wrapfig}
 
@@
 \lstset{breaklines=true}
+\renewcommand{\lstlistingname}{Code}
 
@@
 %\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
-\title{ \textmd{\normalsize{High Performance Computing Project}}\\ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
-\date{}
-\author{Manuel Demmeler}
+\title{ %\textmd{\normalsize{High Performance Computing Project}}\\ 
+	Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
+\subtitle{High Performance Computing Project}
+%\date{}
+\author{Manuel Demmeler \and Manfred Liebmann}
+
+\institute{Manuel Demmeler
+\at todo ... \\ \email{manuel.demmeler@yahoo.de}
+\and
+Manfred Liebmann
+\at todo ... \\ \email{todo}
+}
 
@@
 
-\newcommand{\3}{ ^{3\times3} }
 \newcommand{\R}{\mathbb{R}}
 \newcommand{\N}{\mathbb{N}}
-\newcommand{\M}{\mathbb{M}}
 \newcommand{\Sym}{\mathbb{S}^3}
@@
 In this project, a parallelized version of the PRM method has been developed in general.
-The robotics application then was implemented as an special case of this with GPU support.
+A robotics application then was implemented as an special case of this with GPU support.
 
@@
 The abstract problem treated here is to find a path in a d-dimensional space, which avoids intersecting obstacles. The area occupied by obstacles is given by an indicator function
-\begin{equation*}
+\begin{equation}
 	I: \Omega \rightarrow \{0, 1\}
-\end{equation*}
+\end{equation}
 on a region \( \Omega = [ q_{min,1}, q_{max,1} ] \times ... \times  [ q_{min,d}, q_{max,d} ] \subset \R^d  \). Hence, the goal is to find a continuous curve
-\begin{equation*}
+\begin{equation}
 	\Gamma: [0,1] \rightarrow \Omega
-\end{equation*}
+\end{equation}
 between a given start and endpoint, \(\Gamma(0)=q_b,\ \Gamma(1)=q_e \), 
 such that
-\begin{equation*}
+\begin{equation}
 	I(\Gamma(s))=0 \text{\ f.a.\ } s \in [0,1].
-\end{equation*}
+\end{equation}
 The PRM algorithm now grows a graph \(G=(V,E)\) form points \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in \Omega\) and inserting them into $V$.
-Two nodes \(v,w \in V\) are connected by an edge, if they are near enough each other and the line between them is free from obstacles.
+Two nodes \(v,w \in V\) are connected by an edge, if they are near enough each other and the line between them is free from obstacles, based on the following definition \ref{connectiondef}.
 The algorithm terminates, when there exists a path from \(q_s\) to \(q_e\) on G.
-\begin{definition}
+
+\begin{definition}\label{connectiondef}
 For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
-	\begin{equation*}
-		[q_1,q_2]_h \defgl \{ q = \lambda q_1+(1-\lambda) q_2\ |\ \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
-	\end{equation*}
+	\begin{equation}
+	\begin{aligned}
+		[q_1,q_2]_h \defgl \{ &q = \lambda q_1+(1-\lambda) q_2\ | \\ 
+				& \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
+	\end{aligned}
+	\end{equation}
 	We say, \(q_1\) and \(q_2\) are connected with stepsize \(h>0\), if
@@
 
-\begin{algorithm} 
+Algorithm \ref{prmalg} shows the sequential version.
+Hereby, the probability density \(p(\ \cdot\ ,G)\) (depending on the current graph) can be for example first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\).
+Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.
+
+\begin{algorithm} \label{prmalg}
 	\caption{Probabilistic Roadmap Algorithm}
@@
 
-The probability density \(p(\ \cdot\ ,G)\) (depending on the actual graph) can be for example first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\).
-Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.
-
 \subsection{Application to Robotics}
@@
 
-\newpage
+%\newpage
 \subsection{Project Overview}
@@
 Therefore it is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every GPU thread computes one evaluation.
-In the robotics case, this means, the direct kinematic and collision calculation are made on the GPU parallelly for different joint angle configurations.
-Although this will the only major application in the project, the PRM solver and the configuration space are implemented separated from each other and connected by an interface to be free to simply add other applications.
+In the robotics case, this means, the direct kinematics and collision calculations are made on the GPU parallelly for different joint angle configurations.
+Although this will the only major application in the pro\-ject, the PRM solver and the configuration space are implemented separated from each other and connected by an interface to be free to simply add other applications, as shown in figure \ref{interfacefig}.
 As an example and for testing, additionally a simple 2D configuration space is provided,
 whose indicator function is read from a black-and-white image.
+
+While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
+a second level can be achieved by sampling and connecting several new nodes parallelly. 
+This is more difficult, because all new nodes have to be connected 
+and the graph data has to be updated on all processes. 
+For this, different approaches have been implemented, which are described in chapter \ref{solver}.
+
+Before that, chapter \ref{indicator} gives more details about the robotics indicator function.
 
@@
 		\begin{tabular}{c}
-		\textbf{Configpace}\\
+		\textbf{Configspace}\\
 		indicator fct. \\
@@
 \end{center}
+\label{interfacefig}
+\caption{Project structure. Probabilistic roadmap solver and configuration space implementation are separated.}
 \end{figure*}
-
-While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
-a second level can be achieved by sampling and connecting several new nodes parallelly. 
-This is more difficult, because all new nodes have to be connected 
-and the graph data has to be updated on all processes. 
-For this, different approaches have been implemented, which are described in chapter \ref{solver}.
-
-Before that, chapter \ref{indicator} gives more details about the robotics indicator function.
 
@@
 The coordinate frames of the single parts are described by transformation matrices 
-$T_i \in R^{4x4}$, which transform points from the body frame into the world frame.
+$T_i \in R^{4 \times 4}$, which transform points from the body frame into the world frame.
 They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
@@
 To apply the probabilistic roadmap algorithm, we express this by the indicator function
-
-\begin{equation*}
+\begin{equation}
 	I(q)=
 	\begin{cases}
-		0,	& \text{ for no collision for all given pairs of polytopes } \\
+		0,	& \text{ for no collision at any polytope pair} \\
 		1,	& \text{ if at least one collision occurs }
 	\end{cases}
-\end{equation*}
+\end{equation}
 
-The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
+The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given a set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
 Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.
@@
 It has to be noted that the trajectories \(q(t)\), the PRM algorithm will generate, are piecewise linear and therefore not differentiable. 
-Hence, they have to be smoothed before the use on a robot.
+Hence, before the use on a robot they have to be smoothed. 
 This is not part of the project. 
@@
 or prismatic, a translation along an axis.
-These joints may now be defined through DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
+These joints can now be defined through DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
 by setting the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies to
@@
 \end{equation}
-where $R_x, R_z$ are rotations and $T_x, T_z$ translations around the $x$ and $z$ axes.
+where $R_x, R_z$ are rotations and $T_x, T_z$ translations aro\-und the $x$ and $z$ axes.
 With this and $T_0=I$, all transformations $T_i$ are determined.
@@
 Furthermore, as the geometric calculations made here are numerically not very critical, the usage of double precision has turned out to be not necessary.
-The library contains hybrid host and device inline functions for common vector operations such as
+The library contains hybrid host and device inline functions for common vector operations as shown in code \ref{code1}.
+Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
+Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
+The second constructor in code \ref{code2} implements the relative DH transformations.
 
 \footnotesize
-\begin{lstlisting}[language=C++, frame=single, float=*]
+\begin{lstlisting}[language=C++, frame=single, float=*, 
+	caption=Operator examples of the written float4 library.,
+	label=code1
+]
 __host__ __device__ inline float4& operator += (float4& u, const float4& v);
@@
 
-Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
-
-Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
-
 \footnotesize
-\begin{lstlisting}[language=C++, frame=single, float=*]
+\begin{lstlisting}[language=C++, frame=single, float=*,
+	caption=Class structure for a geomeric transformation matrix.,
+	label=code2
+]
 class trafo4{
@@
 \end{lstlisting}
-
-The second constructor implements the relative DH transformations.
+\normalsize
 
@@
 \footnotesize
-\begin{lstlisting}[language=C++, frame=single, float=*]
+\begin{lstlisting}[language=C++, frame=single,
+	caption=Storing polytopes in CRS format.,
+	label=code3
+]
 struct polytope4{
+    //vertices
     float4* vertices;	//length n
     int n; 	
-    //edges
+    //edge adjacency matrix
     int* dsp;	//length n
@@
 \end{lstlisting}
+\normalsize
 
-It means, that for $i=0,...,n-1$ the vertices $dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
+It means, that for $i=0,...,n-1$ the vertices with indices $dest[dsp[i]],..., dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
 The polytopes used in this project are given by the user as convex hull of a list of vertices.
@@
 
-\footnotesize
-\begin{lstlisting}[language=C++, frame=single, float=*]
-__host__ __device__ int separating_vector_algorithm(const polytope4& P, const polytope4& Q, const trafo4& tp, const trafo4& tq);
-\end{lstlisting}
-
 In our application, the polytopes are mostly parts of the robot, which are given in a body frame and have to be transformed into world frame according to section \ref{dk} first.
@@
 To avoid this, all polytopes are passed in their body frames together with their transformations and the algorithm transforms a vertex, whenever it occurs in calculations.
-This is also the version proposed in \cite{chung}.
+This also is the version proposed in \cite{chung}.
+
+%\footnotesize
+%\begin{lstlisting}[language=C++, frame=single, float=*]
+%__host__ __device__ int separating_vector_algorithm(const polytope4& P, const polytope4& Q, const trafo4& tp, const trafo4& tq);
+%\end{lstlisting}
+%\normalsize
 
@@
 
-The robot configuration space is implemented in the form
-
 \footnotesize
-\begin{lstlisting}[language=C++, frame=single, float=*]
+\begin{lstlisting}[language=C++, frame=single, float=*,
+	caption=Robot configuration space class structure.,
+	label=code4
+]
 template<int ndof>
@@
 \end{lstlisting}
+\normalsize
+
+%\footnotesize
+%\begin{lstlisting}[language=C++, frame=single, float=*,
+%	caption=GPU collision kernel for testing robot configurations.,
+%	label=code5
+%]
+%__global__ void kernel_indicator2(const Robot<ndof>* robot,
+%                                  const collision4::polytope4data* polydata,
+%                                  const float* qs, int offsets,
+%                                  const float* qe, int offsete,
+%                                  int* res,
+%                                  const int* testpos, const int* testnum,
+%                                  int N, int numthreads);
+%\end{lstlisting}
+%\normalsize
 
 \footnotesize
-\begin{lstlisting}[language=C++, frame=single, float=*]
-__global__ void kernel_indicator2(const Robot<ndof>* robot,
-                                  const collision4::polytope4data* polydata,
-                                  const float* qs, int offsets,
-                                  const float* qe, int offsete,
+\begin{lstlisting}[language=C++, frame=single, float=*,
+	caption=GPU collision kernel for testing robot configurations.,
+	label=code5
+]
+__global__ void kernel_indicator2(const Robot<ndof>* __restrict__ robot,
+                                  const collision4::polytope4data_restrict polydata,
+                                  const float* __restrict__ qs, int offsets,
+                                  const float* __restrict__ qe, int offsete,
                                   int* res,
-                                  const int* testpos, const int* testnum,
-                                  int N, int numthreads);
+                                  const int* __restrict__ testpos, const int* __restrict__ testnum,
+                                  int N, int numthreads)
 \end{lstlisting}
+\normalsize
 
-where the constant geometry and robot data is loaded once at the beginning and copied to the GPU by an init function.
+%The robot configuration space is implemented in the form
+Code \ref{code4} exemplarily shows the implementation of the robot configuration space.
+The geometry and robot data is load\-ed once at the beginning and copied to the GPU by an init function.
+It is constant during runtime.
 The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
-
 \begin{equation}
@@
 		0	&\text{if $I(q)=0$ f.a. $q \in \left[qs[j], qe[j]\right]_{\Delta q}$} \\
+		%0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
 		1	&else
@@
 \end{equation}
-
+This computation is done by calling the kernel in Code \ref{code5}.
 The border points are stored in structures of arrays.
-
 \begin{equation}
 \begin{aligned}
-	q_{start}^j &= \left(qs[j],qs[j+\text{offset}],...,qs[j+(d-1)*\text{offset}]\right)^T\\
-	q_{end}^j &= \left(qe[j],qe[j+\text{offset}],...,qe[j+(d-1)*\text{offset}]\right)^T
+	q_{start}^j = (&qs[j],qs[j+\text{offset}], \\
+			&..., qs[j+(d-1)*\text{offset}])^T\\
+	q_{end}^j = (&qe[j],qe[j+\text{offset}], \\
+			&..., qe[j+(d-1)*\text{offset}])^T
 \end{aligned}
 \end{equation}
-
-As the PRMSolver only needs to know, which node points can be connected by a line, and not, which intermediate points are free in detail, this version delivers exactly the minimal necessary information.
-Furthermore, it saves CPU and communication time, because the intermediate points on the lines can be directly computed by the GPU threads and do not need to be exchanged.
-
+As the PRMSolver only needs to know, which node points can be connected by a line, and not, which intermediate points are free in detail, the kernel delivers exactly this minimal necessary information.
+Furthermore, CPU and communication time is saved,
+by computing intermediate points of the lines directly on the GPU and not exchanging them.
 Algorithms \ref{host} and \ref{kernel} show in pseudo code, how the work is split between host and device.
@@
 
-In order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
+Furthermore, in order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
 Here the function is divided into everything until the kernel launch and a function, which waits, until the kernel has finished and receives the result data.
+This implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corresponding waiting calls. 
+With that, it is possible to use the full computation time of the GPU.
 
-\begin{lstlisting}
-int indicator2_async(const float* qs, const float* qe, int *res, 
-					 const int N, const int offset, int &request);
-int indicator2_async_wait(int request);
-\end{lstlisting}
+%\footnotesize
+%\begin{lstlisting}[language=C++, frame=single, float=*,
+%	caption=Asynchronous versions of indicator functions.,
+%	label=code6
+%]
+%int indicator2_async(const float* qs, const float* qe, int *res, 
+%                     const int N, const int offset, int &request);
+%int indicator2_async_wait(int request);
+%\end{lstlisting}
+%\normalsize
 
-The implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corresponding waiting calls. 
-With this, it is possible to use the full computation time of the GPU.
-
-\newpage
+%\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\begin{lstlisting}
+\footnotesize
+\begin{lstlisting}[language=C++, frame=single, float=*,
+	caption=Roadmap graph structure.,
+	label=code7
+]
 struct block{
@@
 };
-\end{lstlisting}
 
-\begin{lstlisting}
 struct graph{
@@
 \end{lstlisting}
+\normalsize
 
 The key of an block is computed by the mapping
-\begin{equation*}
+\begin{equation}
 	key(q) = \left\lfloor \frac{q_1}{H} \right\rfloor
-\end{equation*}
+\end{equation}
 such that $H>0$ becomes a parameter to adjust the grid density.
@@
 
-For getting the neighbors of a node $q$ it is then enough to look at the keys from $key(q-cD)$ to $key(q+cD)$. For $D=H$, $c=1$ is possible.
+For getting the neighbors of a node $q$ it is then enough to look at the keys from $key(q_1-cD)$ to $key(q_1+cD)$. For $D=H$, $c=1$ is possible.
 The key mapping could be generalized by taking more components of $q$ into account, but that was not done here.
 
-\subsection{Parallelization Models}
+\subsection{Parallelization Approaches}
 
@@
 
+%\begin{figure}	
 \begin{center}
-\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
+\resizebox{\columnwidth}{!}{
+\begin{tikzpicture}[auto, node distance=0.5cm and 1.0cm]
 	\node[block, minimum width=3cm] (prm) {
@@
 		\begin{tabular}{c}
-		\textbf{Configpace 1}
+		\textbf{Configspace 1}
 		\end{tabular}
@@
 		\begin{tabular}{c}
-		\textbf{Configpace n}
+		\textbf{Configspace n}
 		\end{tabular}
@@
 \end{tikzpicture}
+}
 \end{center}
+%\caption{First version of parallelization: distributed node generation.}
+%\end{figure}
 
@@
 
+%\begin{figure*}	
 \begin{center}
-\begin{tikzpicture}[auto, node distance=0.5cm and 1.8cm]
+\resizebox{\columnwidth}{!}{
+\begin{tikzpicture}[auto, node distance=0.5cm and 1.0cm]
 	\node[block, minimum width=3cm] (prm) {
@@
 \end{tikzpicture}
+}
 \end{center}
+%\caption{Second version of parallelization: redundant node generation. }
+%\end{figure*}
 
@@
 Here, every process runs two workers overlapped. 
-In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round.
+In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round, as shown in figure \ref{figversion3}.
 With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
@@
 	\node[frameless, left=of l1] (g1) {};
-	\node[block, below=of g1, minimum height=14cm] (graph) {\rotatebox{90}{\textbf{Graph}}};
+	\node[block, below=of g1, minimum height=13cm] (graph) {\rotatebox{90}{\textbf{Graph}}};
 
@@
 	\node[frameless, right=of topa, gpuwidth] (topb) {$\vdots$};
-	\node[block2, below=0.7cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
-	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=4.2cm] (stream1b) {\rotatebox{90}{kernel}};
-	\node[frameless, below=0.3cm of stream1b, gpuwidth] (bota) {$\vdots$};
+	\node[block2, below=0.4cm of topa, gpuwidth, minimum height=4.3cm] (stream1a) {\rotatebox{90}{kernel}};
+	\node[block2, below=2.3cm of stream1a, gpuwidth, minimum height=3.8cm] (stream1b) {\rotatebox{90}{kernel}};
+	\node[frameless, below=0.1cm of stream1b, gpuwidth] (bota) {$\vdots$};
 
-	\node[block2, below=2.2cm of topb, gpuwidth, minimum height=7.6cm] (stream2) {\rotatebox{90}{kernel}};
+	\node[block2, below=2.0cm of topb, gpuwidth, minimum height=7.1cm] (stream2) {\rotatebox{90}{kernel}};
 	\node[frameless, right=of bota, gpuwidth] (botb) {$\vdots$};
@@
 \end{center}
-\caption{test}
+\label{figversion3}
+\caption{Structure of asynchronous kernel launches.}
 \end{figure*}
 
-\newpage
+%\newpage
 \section{Results}
@@
 	\centering
-	\includegraphics[width=0.5\textwidth]{robotsolution.pdf}
+	\includegraphics[width=0.5\textwidth]{robotsolution2_5.pdf}
 	\caption{PRM generated collision-free movement for a 4-axis robot}
@@
 \begin{table}[h]\label{resulttable}
+\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
 \centering
-\begin{tabular}{| c | c | r | r | r |}
+%\begin{tabular}{| c | c | r | r | r |}
+
+\begin{tabular}{ c  c  r  r  r }
 
 \hline
-version	&  processes/	& time (ms)	& GPU 		& time per 1e6  \\ 
-		& GPUs		& 			& threads		& GPU threads (ms) \\ 
+version	&  processes/	& time		& GPU 		& time per 1e6  \\ 
+		& GPUs		& (ms)		& threads		& threads (ms) \\ 
 \hline
@@
 \end{tabular}
-\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
 \end{table}
@@
 As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
-In most cases, the GPU code was around 10 times faster. However, the robot code is not optimized for the CPU version, such that this comparison is not quite fair.
+In most cases, the GPU code was around 10 times faster. 
+However, the robot code is not optimized for the CPU version, such that this comparison is not quite fair.
 
-\newpage
+%\newpage
 
-\bibliographystyle{alpha}
-\begin{thebibliography}{999}
+%\bibliographystyle{alpha}
+\begin{thebibliography}{}%{999}
 	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, source: http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html

### 2016-07-12 prm/matlab/lib/pathplot.m

@@ -18,7 +18,7 @@ function [] = pathplot( dhparams, ndof, polys, N, Q, dt, infinite)
         lim=[xlim;ylim;zlim];
         [AZ,EL]=view;
         
-        clf;
+        %clf;
         stateplot(dhparams,ndof,polys,N,Q(:,t),colors);
         
         
@@ -29,7 +29,7 @@ function [] = pathplot( dhparams, ndof, polys, N, Q, dt, infinite)
         zlim(limneu(3,:));
         view(AZ,EL);
         
-        pause(dt);
+        %pause(dt);
     end
         if ~infinite
             break;

### 2016-07-12 prm/matlab/robotplot.m

@@ -19,7 +19,12 @@ Qpathcont=interp1q((1:l)',prmoutput.Qpath',(1:0.4:l)')';
 %Qpathcont=interp1q((1:l)',prmoutput.Qpath',[1;18;20;21;25;l])';
 
 
-pathplot(dhparams,ndof,polys,N,Qpathcont, 0.06, true);
+%pathplot(dhparams,ndof,polys,N,Qpathcont, 0.06, true);
+
+pathplot(dhparams,ndof,polys,N, ...
+    [prmoutput.Qpath(:,1),prmoutput.Qpath(:,round(l/4)),...
+    prmoutput.Qpath(:,round(l*0.7)),prmoutput.Qpath(:,l)], ...
+    0.06, false);
 
 
 

### 2016-07-12 paper/paper.tex
paper -> fuer korrektur

@@
   collision4::polytope4data* polydatadev; //GPU object
+  collision4::polytope4data_restrict* polydatadev_restrict; //restricted pointers collection
   .
@@
     int blocknum;     //number of used blocks
+    .
     .
@@
 
-\begin{figure*}
+\begin{figure}
 \begin{center}
+\resizebox{\columnwidth}{!}{
 \begin{tikzpicture}
@@
 \end{tikzpicture}
+}
 \end{center}
@@
 \caption{Structure of asynchronous kernel launches.}
-\end{figure*}
+\end{figure}
 
@@
 \begin{figure}%{r}{0.5\textwidth}
-	\label{robotsolution}
 	\centering
 	\includegraphics[width=0.5\textwidth]{robotsolution2_5.pdf}
-	\caption{PRM generated collision-free movement for a 4-axis robot}
+	\caption{PRM generated collision-free movement for a 4-axis robot}\label{robotpic}
 \end{figure}
 
-Tests have been made for a scenario with a robot with 4 joints. For creating and plotting the environment and the resulting trajectories, Matlab functions are provided.
+Tests have been made for a scenario with a robot with 4 joints, shown in figure \ref{robotpic}.
+This scenario is difficult because of the small path to be passed in the middle.
+For creating and plotting the environment and the resulting trajectories, Matlab functions were written.
 Because of the dependency on random numbers, the algorithm takes different amounts of time for different runs.
@@
 Table 1 shows the results.
-One can see, that the version with node communication has a better time to GPU threads ratio, but leads to a greater amount of threads on average.
-The versions with less communication, however, have better absolute times.
-Furthermore, one sees the improvement from version 2 to version 3 from the asynchronous indicator function.
+
+%\begin{table}[h]\label{resulttable}
+%\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
+%\centering
+%%\begin{tabular}{| c | c | r | r | r |}
+%
+%\begin{tabular}{ c  c  r  r  r }
+%
+%\hline
+%version	&  processes/	& time		& GPU 		& time per 1e6  \\ 
+%		& GPUs		& (ms)		& threads		& threads (ms) \\ 
+%\hline
+%		& 1			& 1002		& 696413		& 1448	\\ 
+%1		& 2			& 763		& 838938		& 918	\\ 
+%		& 4			& 706		& 1113781	& 647	\\ 
+%\hline
+%		& 1			& 900		& 608172		& 1490	\\ 
+%2		& 2			& 647		& 608172		& 1070	\\ 
+%		& 4			& 568		& 608172		& 947	\\ 
+%\hline
+%		& 1			& 696		& 599847		& 1181	\\ 
+%3		& 2			& 554		& 599847		& 951	\\ 
+%		& 4			& 500		& 599847		& 864	\\ 
+%\hline
+%
+%\end{tabular}
+%\end{table}
 
 \begin{table}[h]\label{resulttable}
-\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
+\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Eight-Core CPU E5-2650 0 @ 2.00GHz and 4x Tesla K20m.}
 \centering
@@
 \hline
-version	&  processes/	& time		& GPU 		& time per 1e6  \\ 
+version	&  processes/	& time		& GPU 		& time per $10^6$  \\ 
 		& GPUs		& (ms)		& threads		& threads (ms) \\ 
 \hline
-		& 1			& 1002		& 696413		& 1448	\\ 
-1		& 2			& 763		& 838938		& 918	\\ 
-		& 4			& 706		& 1113781	& 647	\\ 
+		& 1			& 6864		& 6171525	& 1070	\\ 
+1		& 2			& 1580		& 4489213	& 342	\\ 
+		& 4			& 869		& 5853763	& 149	\\ 
 \hline
-		& 1			& 900		& 608172		& 1490	\\ 
-2		& 2			& 647		& 608172		& 1070	\\ 
-		& 4			& 568		& 608172		& 947	\\ 
+		& 1			& 2136		& 6933768	& 337	\\ 
+2		& 2			& 1778		& 6933768	& 290	\\ 
+		& 4			& 1677		& 6933768	& 279	\\ 
 \hline
-		& 1			& 696		& 599847		& 1181	\\ 		
-3		& 2			& 554		& 599847		& 951	\\ 
-		& 4			& 500		& 599847		& 864	\\ 
+		& 1			& 2492		& 19096275	& 130	\\ 
+3		& 2			& 1345		& 19096275	& 72		\\ 
+		& 4			& 942		& 19096275	& 52		\\ 
+\hline
+3		& 4			& 32730		& 10038198	& 3281	\\ 
+ (CPU	& 8			& 17676		& 10038198	& 1785	\\ 
+only)		& 16			& 9936		& 10038198	& 1012	\\ 
 \hline
@@
 As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
-In most cases, the GPU code was around 10 times faster. 
-However, the robot code is not optimized for the CPU version, such that this comparison is not quite fair.
+The timing results for the last one are shown in the last section of table 1.
+One can see an improvement of up to a factor around 10.
+However, the robot code is not optimized for the CPU version, as the GPU kernel is just rewritten with an for loop.
+Therefore this comparison is not quite fair.
 

### 2016-07-15 paper/paper.tex
paper final version (bis auf autor daten..)

@@
 
-%\usepackage{fancyhdr}
 \usepackage{paralist}
@@
 \usepackage{mathtools}
-
-%\usepackage{hyperref}
-
-%\usepackage{graphicx}
-%\usepackage{wrapfig}
 
@@
 \renewcommand{\lstlistingname}{Code}
-
-%\usepackage{xcolor}
-%\definecolor{lbcolor}{rgb}{0.9,0.9,0.9}
-%
-%\lstdefinestyle{customc}{
-%  belowcaptionskip=1\baselineskip,
-%  breaklines=true,
-%  %frame=LR,
-%  %xleftmargin=\parindent,
-%  language=C,
-%  showstringspaces=false,
-%  basicstyle=\footnotesize\ttfamily,
-%  keywordstyle=\bfseries\color{green!40!black},
-%  commentstyle=\itshape\color{purple!40!black},
-%  identifierstyle=\color{blue},
-%  stringstyle=\color{orange},
-%}
-%
-%\lstdefinestyle{customcpp}{
-%	backgroundcolor=\color{lbcolor},
-%    	tabsize=4,
-%    	language=[GNU]C++,
-%        basicstyle=\tiny,%{\scriptsize},
-%        upquote=true,
-%        %aboveskip={1.5\baselineskip},
-%        columns=fixed,
-%        showstringspaces=false,
-%        extendedchars=false,
-%        breaklines=true,
-%        prebreak = \raisebox{0ex}[0ex][0ex]{\ensuremath{\hookleftarrow}},
-%        frame=single,
-%        numbers=left,
-%        showtabs=false,
-%        showspaces=false,
-%        showstringspaces=false,
-%        identifierstyle=\ttfamily,
-%        keywordstyle=\color[rgb]{0,0,1},
-%        commentstyle=\color[rgb]{0.026,0.112,0.095},
-%        stringstyle=\color[rgb]{0.627,0.126,0.941},
-%        numberstyle=\color[rgb]{0.205, 0.142, 0.73},
-%}
-%
-%\lstset{escapechar=@,style=customcpp}
-%
 
@@
 
-\DeclareMathOperator*{\esssup}{ess.sup}
 \DeclareMathOperator*{\defgl}{\vcentcolon=}
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\N}{\mathbb{N}}
+\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
+\newcommand{\abs}[1]{\left |#1\right |}
 
-%\pagestyle{fancy}
-%\fancyhead[L]{Seminar \glqq Optimale Steuerung\grqq\\ Prof. Brokate, Prof. Kuttler} %Kopfzeile links
-%\fancyhead[C]{} %zentrierte Kopfzeile
-%\fancyhead[R]{10. Dezember 2013\\ Manuel Demmeler} %Kopfzeile rechts
-\title{ %\textmd{\normalsize{High Performance Computing Project}}\\ 
-	Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
+\usepackage{hyperref}
+
+\title{ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
 \subtitle{High Performance Computing Project}
@@
 \institute{Manuel Demmeler
-\at todo ... \\ \email{manuel.demmeler@yahoo.de}
+\at todo ... \\ 
+\email{manuel.demmeler@yahoo.de}
 \and
 Manfred Liebmann
-\at todo ... \\ \email{todo}
+\at todo ... \\  %TODO: which adresses to use?
+\email{todo}
 }
-
-%\thispagestyle{fancy}
-
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\N}{\mathbb{N}}
-\newcommand{\Sym}{\mathbb{S}^3}
-\newcommand{\Mpos}{\M_+^3}
-\newcommand{\Orth}{\mathbb{O}^3}
-\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
-\newcommand{\abs}[1]{\left |#1\right |}
-\newcommand{\ub}{\bar{u}}
-\newcommand{\yb}{\bar{y}}
-\newcommand{\scprod}[2]{\langle#1 , #2 \rangle}
-
-%\vspace{-3.4cm}
 
@@
 \subsection{Problem Statement and the Sequential PRM Method}\label{algintro}
-%\newtheorem*{defi}{Definition}
-%\newtheorem*{alg1}{Algorithm}
 
@@
     \caption{Example from \cite{prmlec} for an indicator function in robotics. 
-    		The real obstacles on the left are transformed 
+    		The given obstacles on the left are transformed 
 		into occupied areas of the configuration space on the right.}
@@
 
-%\newpage
 \subsection{Project Overview}
@@
 For this, different approaches have been implemented, which are described in chapter \ref{solver}.
-
 Before that, chapter \ref{indicator} gives more details about the robotics indicator function.
@@
 \end{center}
-\label{interfacefig}
-\caption{Project structure. Probabilistic roadmap solver and configuration space implementation are separated.}
+\caption{Project structure. Probabilistic roadmap solver and configuration space implementation are separated.}\label{interfacefig}
 \end{figure*}
@@
     .
-    .
 }
@@
 
-It means, that for $i=0,...,n-1$ the vertices with indices $dest[dsp[i]],..., dest[dsp[i]+cnt[i]-1]$ are the ones connected with vertex $i$ by an edge.
+It means, that for $i=0,...,n-1$ the vertices with indices $\mathrm{ dest[dsp[i]],..., dest[dsp[i]+cnt[i]-1] }$ are the ones connected with vertex $i$ by an edge.
 The polytopes used in this project are given by the user as convex hull of a list of vertices.
@@
   .
-  .
 private:
@@
   collision4::polytope4data_restrict* polydatadev_restrict; //restricted pointers collection
-  .
   .
@@
 \begin{equation}
-	res[j]=\begin{cases}
-		0	&\text{if $I(q)=0$ f.a. $q \in \left[qs[j], qe[j]\right]_{\Delta q}$} \\
+	\mathrm{res[j]}=\begin{cases}
+		0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
 		%0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
-		1	&else
+		1	&\text{else}
 	\end{cases}
@@
 \begin{aligned}
-	q_{start}^j = (&qs[j],qs[j+\text{offset}], \\
-			&..., qs[j+(d-1)*\text{offset}])^T\\
-	q_{end}^j = (&qe[j],qe[j+\text{offset}], \\
-			&..., qe[j+(d-1)*\text{offset}])^T
+	q_{start}^j = (&\mathrm{qs[j],qs[j+\text{offset}], }\\
+			&\mathrm{..., qs[j+(d-1)*\text{offset}]})^T\\
+	q_{end}^j = (&\mathrm{qe[j],qe[j+\text{offset}], }\\
+			&{..., qe[j+(d-1)*\text{offset}]})^T
 \end{aligned}
@@
 	compute distances $d_j = \abs{q_{end}^j - q_{start}^j}$ \;
-	testnum[j] = $d_j/{\Delta q}$: number of threads for each line \;
-	testpos[j] = thread displacements \;
+	$\mathrm{testnum[j]} = d_j/{\Delta q}$: number of threads for each line \;
+	$\mathrm{testpos[j]} = \mathrm{thread\ displacements}$ \;
 	reset res \;
-	call kernel \;
+	call kernel with $\sum_{j=0}^N \mathrm{testnum[j]}$ threads\;
 	\caption{Host indicator function}
@@
 	compute associated edge number $j$, defined by $testpos[j] \leq threadindex \le testpos[j]+testnum[j]$ \;
-	$c = (threadindex - testpos[j]) / (testnum[k]-1) $ \;
+	$c = (\mathrm{threadindex - testpos[j]}) / (\mathrm{testnum[k]}-1) $ \;
 	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
 	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
-	$res_{tmp} = 0; $ \;
+	$\mathrm{res\_tmp} = 0; $ \;
 	\For{all registered pairs pairs of polytopes $P, Q$}{
-		$res_{tmp}=separating\_vector\_algorithm(P, T_{i_P}, Q, T_{i_Q})$ \;
-		\If{$res_{tmp} \neq 0$}{
-			$res[j]=res_{tmp}$ \;
+		$\mathrm{res\_tmp}=\mathrm{separating\_vector\_algorithm}(P, T_{i_P}, Q, T_{i_Q})$ \;
+		\If{$\mathrm{res\_tmp} \neq 0$}{
+			$\mathrm{res}[j]=\mathrm{res\_tmp}$ \;
 			break \;
@@
 
-%\newpage
 %%%%%%%%%%%%%%%%%%%%%%%%
@@
 If no special order of the nodes is given, one has to calculate this norms for all nodes of the graph, which is rather time consuming.
-
 Therefore, a special indexing is used here. 
-The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map.
+The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map. See Code \ref{code7} for the data structures.
 
@@
     int blocknum;     //number of used blocks
-    .
     .
@@
 
-\subsubsection*{Version 3}
-
-The third version does a further optimization by using the asynchronous version of the indicator function.
-Here, every process runs two workers overlapped. 
-In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round, as shown in figure \ref{figversion3}.
-With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
-
 \begin{figure}
@@
 \end{center}
-\label{figversion3}
-\caption{Structure of asynchronous kernel launches.}
+\caption{Structure of asynchronous kernel launches.}\label{fig:v3}
 \end{figure}
 
-%\newpage
+\subsubsection*{Version 3}
+
+The third version does a further optimization by using the asynchronous version of the indicator function.
+Here, every process runs two workers overlapped. 
+In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round, as shown in figure \ref{fig:v3}.
+With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
+
 \section{Results}
@@
 
-%\begin{table}[h]\label{resulttable}
-%\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Six-Core CPU X5650 @2.67GHz and 4x GeForce GTX 480}
-%\centering
-%%\begin{tabular}{| c | c | r | r | r |}
-%
-%\begin{tabular}{ c  c  r  r  r }
-%
-%\hline
-%version	&  processes/	& time		& GPU 		& time per 1e6  \\ 
-%		& GPUs		& (ms)		& threads		& threads (ms) \\ 
-%\hline
-%		& 1			& 1002		& 696413		& 1448	\\ 
-%1		& 2			& 763		& 838938		& 918	\\ 
-%		& 4			& 706		& 1113781	& 647	\\ 
-%\hline
-%		& 1			& 900		& 608172		& 1490	\\ 
-%2		& 2			& 647		& 608172		& 1070	\\ 
-%		& 4			& 568		& 608172		& 947	\\ 
-%\hline
-%		& 1			& 696		& 599847		& 1181	\\ 
-%3		& 2			& 554		& 599847		& 951	\\ 
-%		& 4			& 500		& 599847		& 864	\\ 
-%\hline
-%
-%\end{tabular}
-%\end{table}
-
-\begin{table}[h]\label{resulttable}
-\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Eight-Core CPU E5-2650 0 @ 2.00GHz and 4x Tesla K20m.}
+\begin{table}
+\caption{Timing results for the parallelization versions from chapter \ref{solver} on a cluster node with 2x Intel Xeon Eight-Core CPU E5-2650 0 @ 2.00GHz and 4x Tesla K20m.}\label{resulttable}
 \centering
-%\begin{tabular}{| c | c | r | r | r |}
 
@@
 \hline
+\hline
+version	&  CPU		& time		& indicator 	& time per $10^6$  \\ 
+		& threads		& (ms)		& evaluations	& evals. (ms) \\
+
+\hline
 3		& 4			& 32730		& 10038198	& 3281	\\ 
@@
 As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
-The timing results for the last one are shown in the last section of table 1.
+The timing results for the last one are shown in the last section of table \ref{resulttable}.
 One can see an improvement of up to a factor around 10.
-However, the robot code is not optimized for the CPU version, as the GPU kernel is just rewritten with an for loop.
+However, the robot code is not optimized for the CPU version, as the GPU kernel has just been rewritten with an for loop
+reusing all the device functions for collision detection.
 Therefore this comparison is not quite fair.
 
-%\newpage
+%%%%%%%%%%%%%%%%%%%%%%%%
 
-%\bibliographystyle{alpha}
 \begin{thebibliography}{}%{999}
-	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, source: http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html
-	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press, 2005
-	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press, 2006
-	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München, 2014
+	\bibitem{prmlec} D. Burschka, Lecture Robot Motion Planning, \url{http://robvis01.informatik.tu-muenchen.de/courses/wegtraj/index.html}
+	\bibitem{prm1} H. Choset, K. Lynch, S. Hutchinson, G. Kantor, W. Burgard, L. Kavraki and S. Thrun, Principles of Robot Motion: Theory, Algorithms, and Implementation, MIT Press 2005
+	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press 2006
+	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München 2014
 	\bibitem{chung} K.
-Chung, Wenping Wang, Quick Collision Detection of Polytopes in Virtual Environments, Proceedings of the ACM Symposium on Virtual Reality Software and Technology (VRST 96), Mark Green (ed.), 1996
-	\bibitem{ericson} C. Ericson, Real-Time Collision Detection, Elsevier, 2005
+Chung, Wenping Wang, Quick Collision Detection of Polytopes in Virtual Environments, Proceedings of the ACM Symposium on Virtual Reality Software and Technology (VRST 96), Mark Green (ed.) 1996
+	\bibitem{ericson} C. Ericson, Real-Time Collision Detection, MK 2005
+	\bibitem{cuda} Nvidia Corporation, Cuda Programming Guide, 2015, \url{http://docs.nvidia.com/cuda/pdf/CUDA_C_Programming_Guide.pdf}
+	\bibitem{cuda2} D. B. Kirk, W. W. Hwu, Programming Massively Parallel Processors, MK 2010
+	\bibitem{c++} Pitt-Francis, Guide to Scientific Computing in C++, Springer 2012
+	\bibitem{mpi} W. Gropp, Using MPI: Portable Parallel Programming with the Message-Passing Interface, MIT Press 2014
 \end{thebibliography}

### 2016-07-26 paper/paper.tex
Anmerkungen

@@
 
-\section{Motivation}
+\section{Abstract}
 In trajectory optimization with obstacles, it is often necessary to provide feasible initial trajectories. 
@@
 
-In this project, a parallelized version of the PRM method has been developed in general.
+In this paper, a parallelized version of the PRM method has been developed in general.
 A robotics application then was implemented as an special case of this with GPU support.
 
-\section{Introduction}
+\section{Probabilistic Roadmap Method}
 \subsection{Problem Statement and the Sequential PRM Method}\label{algintro}
 
-The abstract problem treated here is to find a path in a d-dimensional space, which avoids intersecting obstacles. The area occupied by obstacles is given by an indicator function
+The abstract problem treated here is to find a path in a d-dimensional space $\Omega \subset \R^d$, which avoids intersecting obstacles.
+The area occupied by obstacles is given by an indicator function
 \begin{equation}
@@
 \begin{equation}
-	I(\Gamma(s))=0 \text{\ f.a.\ } s \in [0,1].
+	\forall s\in [0,1]: I(\Gamma(s))=0 .
 \end{equation}
-The PRM algorithm now grows a graph \(G=(V,E)\) form points \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in \Omega\) and inserting them into $V$.
-Two nodes \(v,w \in V\) are connected by an edge, if they are near enough each other and the line between them is free from obstacles, based on the following definition \ref{connectiondef}.
-The algorithm terminates, when there exists a path from \(q_s\) to \(q_e\) on G.
+The PRM algorithm now grows a graph \(G=(V,E)\), where $V$ is the set of vertices and $E$ the set of edges,
+ form points \(q_s\) and \(q_e\) by randomly sampling nodes \(v\in \Omega\) and inserting them into the set $V$.  
+ Two nodes \(v,w \in V\) are connected by an edge, if they are near enough each other and the line between them is free from obstacles, based on definition \ref{connectiondef}, see below. 
+ The algorithm terminates, when there exists a path from \(q_s\) to \(q_e\) on the graph G.
 
 \begin{definition}\label{connectiondef}
-For two points \(q_1\) and \(q_2\) we define the connection with stepsize \(h>0\) as
+For two points \(q_1\) and \(q_2\) we define the line segment partitioned with stepsize \(h>0\) as
 	\begin{equation}
@@
 		[q_1,q_2]_h \defgl \{ &q = \lambda q_1+(1-\lambda) q_2\ | \\ 
-				& \lambda \in [0,1], \norm{q-q_1} \in \N_0 h \}.
+				& \lambda \in [0,1] \land \norm{q-q_1} = k h, k \in \N_0 \}.
 	\end{aligned}
@@
 Algorithm \ref{prmalg} shows the sequential version.
-Hereby, the probability density \(p(\ \cdot\ ,G)\) (depending on the current graph) can be for example first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\).
+Hereby, the probability density \(p_G\) (depending on the current graph) determines the sampling of new nodes.
+$p_G$ is defined for example as first selecting randomly a node \(q_0 \in V\) and then sampling a new point \(q\) in the neighbourhood of \(q_0\).
 Theory about success probabilities and convergence results can be found in \cite{prmlec}, \cite{prm1} or \cite{prm2}.
@@
 	\While{ there exists no path in G between \(q_s\) and \(q_e\)}{
-		Sample \(q\in \Omega\) from a probability density  \(p(\ \cdot\ , G)\), resample until \(I(q)=0\) \\
+		Sample \(q\in \Omega\) from a probability density  \(p_G\), resample until \(I(q)=0\) \\
 		\For{ all nodes \(\tilde{q} \in V\), with \( \norm{q- \tilde q} \le d_0 \) }{
@@
 
-In this project, the PRM algorithm is used to find trajectories for a robot arm, which avoid self collisions and collisions with obstacles in the environment.
+In the paper, the PRM algorithm is used to find trajectories for a robot arm, which avoid self collisions and collisions with obstacles in the environment.
 Such trajectories can be expressed as functions describing the joint variables changing over time, that means as a curve in the configuration space with dimension equal to the number of degrees of freedom of the robot.
 With this association, the trajectory finding problem becomes a path finding problem in the configuration space exactly of the form defined in the previous section.
-The real obstacles transform into obstacles in the configuration space, given by the indicator function, that returns for a given vector of joint angles, if it is feasible or leads to a collision.
+The real obstacles transform into occupied areas in the configuration space.
+Those areas are given by the indicator function, which returns for a given vector of joint angles, if it is feasible or leads to a collision.
 
@@
 
-Like in the stated robotics example, in many cases the most costly operation of the PRM algorithm are the many evaluations of the indicator function \(I(q)\), 
+Like in the stated robotics example, in many cases the most costly operations of the PRM algorithm are the many evaluations of the indicator function \(I(q)\), 
 which however can be done independently for every new \(q\).
-Therefore it is implemented in a vectorized way as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\) by a CUDA kernel, where every GPU thread computes one evaluation.
-In the robotics case, this means, the direct kinematics and collision calculations are made on the GPU parallelly for different joint angle configurations.
-Although this will the only major application in the pro\-ject, the PRM solver and the configuration space are implemented separated from each other and connected by an interface to be free to simply add other applications, as shown in figure \ref{interfacefig}.
-As an example and for testing, additionally a simple 2D configuration space is provided,
-whose indicator function is read from a black-and-white image.
+Therefore the evaluation is implemented in a vectorized form as a function \((q_1, ... , q_P) \mapsto (I(q_1),...,I(q_P))\). 
+The single indicator function computations are done in parallel on one or more graphics processing units (GPU).
+The implementation is done with the programming interface CUDA \cite{cuda}, \cite{cuda2} by Nvidia.
+In the robotics case, this means, the direct kinematics and collision calculations are made on the GPU in parallel for different joint angle configurations.
+Although this will the only major application in the paper, the PRM solver and the configuration space are implemented separately to provide a flexible interface, as shown in figure \ref{interfacefig}.
+%As an example and for testing, additionally a simple 2D configuration space is provided, whose indicator function is read from a black-and-white image.
 
 While the distribution of the indicator function to different CUDA threads is a first level of parallelization, 
-a second level can be achieved by sampling and connecting several new nodes parallelly. 
+a second level can be achieved by sampling and connecting several new nodes in parallel. 
 This is more difficult, because all new nodes have to be connected 
 and the graph data has to be updated on all processes. 
-For this, different approaches have been implemented, which are described in chapter \ref{solver}.
+Different approaches have been implemented, with a description provided in chapter \ref{solver}.
 Before that, chapter \ref{indicator} gives more details about the robotics indicator function.
@@
 The probabilistic roadmap algorithm is applied  to create collision-free trajectories of a robot arm.
-For a robot with given joint angles \(q_i\) one can calculate straightforwardly all positions and orientations  of the single parts of the robot.
-For this, coordinate frames according to the Denavit-Hartenberg convention are used.
-The coordinate frames of the single parts are described by transformation matrices 
-$T_i \in R^{4 \times 4}$, which transform points from the body frame into the world frame.
-They depend on the state $q=(q_1,...,q_d)$ of the robot and on the geometry of the joint angles given by the Denavit-Hartenberg parameters of the robot.
+For a robot with given joint angles \(q_i\) one can calculate straightforward all positions and orientations  of the single parts of the robot.
+For this, coordinate frames according to the Denavit-Hartenberg (DH) \cite{robodyn} convention are used.
+The coordinate frames of reference of the single parts are described by transformation matrices 
+$T_i \in \R^{4 \times 4}, i=1,...,d$, which transform points from the body frame of reference into the world frame of reference.
+They depend on the state $q=(q_1,...,q_d)^T$ of the robot and on the geometry of the joint angles given by the DH parameters of the robot.
 
-The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robots coordinate frame.
-Applying the corresponding transformation $T_i$, its vertices can be transformed into the world frame.
+The geometry of the robot parts is defined by a set of convex polytopes, each of them belonging to one of the robot coordinate frames.
+Applying the corresponding transformation $T_i$, the vertices are transformed into the world frame.
 Environment obstacles are also defined by convex polytopes, given in the world frame.
 
-Now the movement of the robot can be described by a curve $q(t) \in \R^d$ in the configuration space.
-It is collision-free, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
-To apply the probabilistic roadmap algorithm, we express this by the indicator function
+Now the movement of the robot can be described by a curve $q(t) \in \R^d, t\in[0,1]$, in the configuration space. 
+The movement is collision-free, if at no time, any two polytopes of the robot intersect with each other or with the environment after being transformed into the world frame.
+To apply the probabilistic roadmap algorithm, we express this setting by the indicator function
 \begin{equation}
@@
 	\begin{cases}
-		0,	& \text{ for no collision at any polytope pair} \\
-		1,	& \text{ if at least one collision occurs }
+		0,	& \text{ for no collision of any polytope pair} \\
+		1,	& \text{ if at least one collision occurs },
 	\end{cases}
@@
 
-The indicator function is implemented in a vectorized way as an CUDA kernel, which means that, given a set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
-Therefore, the kinematics and collision algorithm are implemented as device code for the use in the kernel.
+The indicator function is implemented in a vectorized form in a CUDA kernel, which means that, given a set of states $q^1,..,q^M$, the indicator function is evaluated for each state by one CUDA thread.
+Therefore, the kinematics and collision algorithm are implemented as device code for the use in the CUDA kernel.
 
-It has to be noted that the trajectories \(q(t)\), the PRM algorithm will generate, are piecewise linear and therefore not differentiable. 
+It has to be noted that the trajectories \(q(t)\), generated by the PRM algorithm, are piecewise linear and are therefore not differentiable at all points. 
 Hence, before the use on a robot they have to be smoothed. 
-This is not part of the project. 
-The task is only to generate feasible trajectories to be passed to an optimizer.
+This is not included in the project. 
+The task is only to generate feasible trajectories to be possibly passed to an optimizer.
 
@@
 
-The Denavit-Hartenberg (DH) convention is a method to describe the geometry of a robots joints, 
-that is, how the transformations between the robots coordinate frames look like in dependency of its joint values $q_1,...,q_d$.
-A robot consists of $d+1$ bodies each of which has a body fixed coordinate frame $B_i$, which is described by a transformation matrix $T_i$ into the world frame.
-It is assumed that each joint is either rotational, that means an rotation around a fixed axis, 
-or prismatic, a translation along an axis.
+The Denavit-Hartenberg convention is a method to describe the geometry of the joints of a robot, 
+that is, how the transformations between the robot coordinate frames look like in dependency of its joint values $q_1,...,q_d$.
+A robot consists of $d+1$ bodies each of which has a coordinate frame $B_i$ attached to the body, which is described by the transformation matrix $T_i$ to the world frame.
+It is assumed that each joint is either rotational, that means a rotation around a fixed axis is performed, 
+or prismatic, where a translation along an axis is performed.
 These joints can now be defined through DH parameters $a_{i-1}, \alpha_{i-1}, d_i, \theta_i \in \R$, $i=1,...,d$, 
-by setting the relative transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies to
+by setting up the transformations $T_{i-1,i}=T_{i-1}^{-1} T_{i}$ between two bodies as
 \begin{equation}
@@
 \end{equation}
-where $R_x, R_z$ are rotations and $T_x, T_z$ translations aro\-und the $x$ and $z$ axes.
+where $R_x, R_z$ are rotations aro\-und and $T_x, T_z$ are translations along the $x$ and $z$ axes.
 With this and $T_0=I$, all transformations $T_i$ are determined.
 
-The other way round, one can now show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by such a set of DH parameters, i.e.
-it exists a set of parameters, such that the above definition leads to the correct transformations for this robot.
+One can show, that by an proper choice of the body frames, every robot with rotational or prismatic joints can be described by such a set of DH parameters, i.e.
+there exists a set of parameters, such that the above definition leads to the correct transformations for this robot.
 For the exact calculations, see \cite{robodyn}.
@@
 To guarantee aligned access on the GPU, float4 is used instead of float3. 
-Furthermore, as the geometric calculations made here are numerically not very critical, the usage of double precision has turned out to be not necessary.
-The library contains hybrid host and device inline functions for common vector operations as shown in code \ref{code1}.
+Furthermore, as the precision of the geometric calculations is numerically not very critical, the usage of single precision instead of double precision did not cause problems.
+The library contains host and device inline functions for common vector operations as shown in code \ref{code1}.
 Operators like $+$ or $*$ are not overloaded to prevent unnecessary temporary variables.
 Transformation matrices are implemented as float4 arrays with inline functions for common matrix operations. Again operators with temporaries were avoided, wherever possible.
-The second constructor in code \ref{code2} implements the relative DH transformations.
 
@@
 
-To calculate the intersection between two convex polytopes, the here used Chung-Wang algorithm only needs their vertices 
+To calculate the intersection between two convex polytopes, the Chung-Wang algorithm \cite{chung} is used. It only needs the polytope vertices 
 and the data, which of them are connected by an edge. 
-For this the adjacency matrix of the edge graph is stored compactly in compact row storage format.
+For this purpose, the adjacency matrix of the edge graph is stored in compressed row storage format.
 
@@
 
-It means, that for $i=0,...,n-1$ the vertices with indices $\mathrm{ dest[dsp[i]],..., dest[dsp[i]+cnt[i]-1] }$ are the ones connected with vertex $i$ by an edge.
-The polytopes used in this project are given by the user as convex hull of a list of vertices.
-However, to determine the edge graph only from the vertices is a nontrivial task, which is done by Matlab here using provided high level functions.
+This means, that for $i=0,...,n-1$ the vertices with indices $\mathrm{ dest[dsp[i]],..., dest[dsp[i]+cnt[i]-1] }$ are the ones connected to vertex $i$ by an edge.
+The polytopes are given by the user as convex hull of a list of vertices.
+However, to determine the edge graph only from the vertices is a nontrivial task, which is done by Matlab using builtin high level functions.
 
@@
 
-To check if two polytopes collide an algorithm by Chung and Wang is used here.
-It uses the separating vector theorem and tries to find a separating vector by iterative search.
+To check if two polytopes collide an algorithm by Chung and Wang \cite{chung} is used here.
+It uses the separating vector theorem and tries to find a separating vector by an iterative search.
 Furthermore, a sub-algorithm determines in each iteration from the candidate history, if it is still theoretically possible to find a separating vector.
-If not, it is returned a collision.
-If on the other side a separating vector is found, the algorithm can return no collision.
-For detailed informations see \cite{chung} and \cite{ericson}, pp. 410-412.
-The method is implemented as device code for the use on GPU kernels.
+If not, a collision is returned.
+If on the other hand a separating vector is found, the algorithm returns no collision.
+For a detailed description of the algorithm see \cite{chung} and \cite{ericson}, pp. 410-412.
+The method is implemented as device code for the use in CUDA kernels.
 
-In our application, the polytopes are mostly parts of the robot, which are given in a body frame and have to be transformed into world frame according to section \ref{dk} first.
-However, from the structure of the Chung-Wang algorithm drops out, that mostly only a small part of the polytopes vertices are needed for the computations.
-Therefore, it would be rather time wasting to transform all vertices before passing them to the algorithm.
-To avoid this, all polytopes are passed in their body frames together with their transformations and the algorithm transforms a vertex, whenever it occurs in calculations.
-This also is the version proposed in \cite{chung}.
+In our application, the polytopes are mostly parts of the robot, which are given in the body frame of reference and have to be first transformed into world frame according to section \ref{dk}.
+However, from the structure of the Chung-Wang algorithm follows, that only a small number of the polytope vertices are needed for the computations. 
+Therefore, it would be rather wasteful to transform all vertices before passing them to the algorithm. 
+To avoid this problem, all polytopes are passed in their body frames together with their transformations 
+and the algorithm transforms a vertex, whenever it is needed in the calculations.
+This is also the version proposed in \cite{chung}.
 
@@
 
-%The robot configuration space is implemented in the form
-Code \ref{code4} exemplarily shows the implementation of the robot configuration space.
-The geometry and robot data is load\-ed once at the beginning and copied to the GPU by an init function.
-It is constant during runtime.
-The indicator function gets as input a list of border points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
+Code \ref{code4} exemplarily shows the class definition of the robot configuration space.
+The geometry and robot data does not change during runtime
+and is load\-ed once at the beginning and copied to the GPU by the 
+\footnotesize
+init
+\normalsize 
+function.
+The indicator function gets as input a list of start and end points $q_{start}^j, q_{end}^j, j=0,...,N$ and checks now for each pair, if the indicator function is always 0 on the line in between.
 \begin{equation}
 	\mathrm{res[j]}=\begin{cases}
-		0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
 		%0	&\text{if $I(q)=0$ f.a. $q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q}$} \\
+		0	&\text{if\ \ } \forall q \in \left[q_{start}^j, q_{end}^j\right]_{\Delta q} : I(q)=0 \\
 		1	&\text{else}
@@
 This computation is done by calling the kernel in Code \ref{code5}.
-The border points are stored in structures of arrays.
+The start and end points are stored in structures of arrays defined by
 \begin{equation}
 \begin{aligned}
-	q_{start}^j = (&\mathrm{qs[j],qs[j+\text{offset}], }\\
-			&\mathrm{..., qs[j+(d-1)*\text{offset}]})^T\\
-	q_{end}^j = (&\mathrm{qe[j],qe[j+\text{offset}], }\\
-			&{..., qe[j+(d-1)*\text{offset}]})^T
+	q_{start}^j &= (q_s^j , q_s^{j+N_{buf}} , ... , q_s^{j+(d-1)*N_{buf}})^T\\
+	q_{end}^j &= (q_e^j , q_e^{j+N_{buf}} , ... , q_e^{j+(d-1)*N_{buf}})^T
 \end{aligned}
 \end{equation}
-As the PRMSolver only needs to know, which node points can be connected by a line, and not, which intermediate points are free in detail, the kernel delivers exactly this minimal necessary information.
-Furthermore, CPU and communication time is saved,
-by computing intermediate points of the lines directly on the GPU and not exchanging them.
+As the PRM solver only needs to know, which start and end points can be connected by a line, the kernel delivers exactly this minimal necessary information. 
+CPU and communication time is also saved by computing interpolating points directly on the GPU and not to exchange them.
 Algorithms \ref{host} and \ref{kernel} show in pseudo code, how the work is split between host and device.
@@
 
-Furthermore, in order to overlap computation time of GPU and CPU, an asynchronous version of the indicator function was written.
-Here the function is divided into everything until the kernel launch and a function, which waits, until the kernel has finished and receives the result data.
-This implementation also allows overlapping of multiple requests, for example two calls of the launch function and after this the two corresponding waiting calls. 
-With that, it is possible to use the full computation time of the GPU.
+Furthermore, in order to overlap computations on the GPU and CPU, an asynchronous version of the indicator function was implemented. 
+Here the host function is split into the part till the kernel launch and a function, which waits, until the kernel has finished and receives the result data.
+This implementation also allows overlapping of multiple requests, for example two calls of the launch function and then the two corresponding waiting calls. 
+With that, it is possible to use the full computation power of the GPU.
 
@@
 
-The solver builds a first phase the roadmap graph from both the start and the end node, until it is connected.
-It is always ensured that the graph consists of two connected components containing the start resp. the end node.
-With this, it can be easily determined at each time, if the whole graph has become connected.
+The solver builds in a first phase the roadmap graph from both the start and the end node, until the graph is connected.
+It is always ensured that the graph consists of two connected components containing the start respectively the end node.
+With this fact, it can be easily determined at each step, if the whole graph is now connected.
 
-At the end, in a second phase, a shortest path from the start to the end node is searched by the Dijkstra algorithm.
-It has shown, that this part needs a neglectable amount of time compared to building the graph.
+At the end, in a second phase of the algorithm, a shortest path from the start to the end node is searched by the Dijkstra algorithm.
+It was observed, that this part of the algorithm needs a neglectable amount of time compared to building the graph.
 Therefore, it is not parallelized here and the focus lies on the first phase.
@@
 
-Apart from the indicator function evaluation, the main problem in building the probabilistic roadmap is, that for each new node $q$ it have to be determined all possible neighbors.
-That is the set of all nodes $\tilde q$ of the graph, with $\norm{\tilde q - q} \le D$.
-If no special order of the nodes is given, one has to calculate this norms for all nodes of the graph, which is rather time consuming.
-Therefore, a special indexing is used here. 
+Apart from the indicator function evaluation, the main problem in building the probabilistic roadmap is, 
+that for each new node $q$ all possible neighbors have to be determined.
+That is the set of all nodes $\tilde q$ of the graph, with $\norm{\tilde q - q} \le D$, with $D>0$ an ajustable algorithm parameter.
+If no special order of the nodes is given, one has to calculate these norms for all nodes of the graph, which is rather time consuming.
+Therefore, a special indexing is used. 
 The nodes are stored in a vector, which is organized in blocks with fixed sizes. The blocks are referred to by ids which are stored in a map. See Code \ref{code7} for the data structures.
@@
 
-The key of an block is computed by the mapping
+The key of a block is computed by the mapping
 \begin{equation}
@@
 such that $H>0$ becomes a parameter to adjust the grid density.
-This has to be chosen so large that enough nodes are stored in one block in order to have efficient memory access.
-On the other side, the smaller it is, the less unnecessary computations have to be done for the neighborhood requests.
-For illustration of the structure, the following pseudocode shows the insertion of a node.
+The parameter $H$ has to be chosen so large that enough nodes are stored in one block in order to have efficient memory accesses.
+On the other hand, the smaller $H$ is, the less unnecessary computations have to be done for the neighborhood requests.
+For illustration of the structure, the following pseudocode, algorithm \ref{keyalg}, shows the insertion of a node.
 
-\begin{algorithm}
+\begin{algorithm}\label{keyalg}
 	\KwData{$q\in\R^d$}
@@
 
-For getting the neighbors of a node $q$ it is then enough to look at the keys from $key(q_1-cD)$ to $key(q_1+cD)$. For $D=H$, $c=1$ is possible.
-The key mapping could be generalized by taking more components of $q$ into account, but that was not done here.
+For getting the neighbors of a node $q$ it is then enough to look at the keys from $key(q_1-cD)$ to $key(q_1+cD)$. 
+If $D=H$, then $c=1$ is possible.
+The key mapping could be generalized by taking more components of $q$ into account or using kd-trees, see \cite{prm1}, p. 209, but that was not implemented.
 
@@
 
-The parallelization is done with MPI. All processes store their own versions of the graph, which are hold up to date between each other.
-Furthermore, it is assumed that every process has its own GPU.
-The work of the single MPI processes could be parallelized further with OpenMP, but this has not been done here.
+The main parallelization is done with MPI. All processes store their own versions of the graph, which are held up to date between the processes.
+It is assumed that every process has its own GPU.
+The work of a single MPI process could be parallelized further with OpenMP, but this has not been pursued.
 Instead, different approaches to distribute the work over several GPUs have been investigated.
@@
 In the first approach, that has been evaluated, every process generates new nodes independently from each other.
-Then the possible neighbors are determined and the connections, which have to be tested.
+Then the possible neighbors are determined and the connections which have to be tested.
 The indicator function is called and returns the edge data.
-After this, the graph must be synchronized between all processes, which needs an all-to-all communication.
+After this, the graph must be synchronized between all processes, which needs an MPI all-to-all communication call.
 As the graph building phase does not need the edge data, it is sufficient to exchange only the nodes and collect the edges at the end.
@@
 \subsubsection*{Version 2}
-The second approach uses the fact, that by synchronizing the random seed, 
+The second approach uses the fact, that by synchronizing a random seed, 
 different processes can produce the same sequence of random numbers independently from each other.
-By using this, every process can generate all new nodes and determine the connection data. 
+By using this property, every process can generate all new nodes and determine the connection data. 
 Then the indicator function computation is split over their configuration space instances and the returned data shared afterwards with each other.
 With this data, all graphs can be updated and it is guaranteed, that they grow exactly in the same way on each process.
-The purpose of this version compared to the first one is, that less data has to be sent. 
-Furthermore, in the first version it can occur, that the different generated nodes can result in very different counts of neighbors and therefore uneven loads on the GPUs.
-Here, the distribution of the computation amount over the GPUs can be chosen almost freely.
+The advantage of this version compared to the first one is, that less data has to be exchanged. 
+Furthermore, in the first version it can occur, that the generated nodes can result in very different counts of neighbors and therefore uneven loads on the GPUs.
+Here, the distribution of the computation amount over the GPUs can be chosen very flexible, without performance loss.
 
@@
 
-The third version does a further optimization by using the asynchronous version of the indicator function.
-Here, every process runs two workers overlapped. 
-In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round, as shown in figure \ref{fig:v3}.
-With this it is possible, that the computation time of the GPU is fully used, and overlapped with the communication.
+The third version includes a further optimization by using the asynchronous version of the indicator function.
+Here, every process runs two workers overlapped,
+between which the node generation and GPU requests are split up.
+Figure \ref{fig:v3} shows the detailed work flow.
+%In the time, when the first worker has generated his indicator function request, launched the kernel and now is waiting until it has finished, the second worker updates the graph and sends his new request, and the other way round.
+With this, it is possible, that the computational ressources of the GPU are fully used, and overlapped with the communication.
 
@@
 
-Tests have been made for a scenario with a robot with 4 joints, shown in figure \ref{robotpic}.
-This scenario is difficult because of the small path to be passed in the middle.
+Benchmark runs have been performed for a realistic scenario with a robot with four joints, shown in figure \ref{robotpic}.
+This scenario is difficult because of the small path to be passed in the middle of the domain in figure \ref{robotpic}.
 For creating and plotting the environment and the resulting trajectories, Matlab functions were written.
-Because of the dependency on random numbers, the algorithm takes different amounts of time for different runs.
+Because of the dependency on random numbers, the execution times differ for different runs.
 Therefore, 10 runs were made for each version and number of processes.
 However, the different versions lead to a different average amount of GPU threads needed.
-Hence, to be able to better compare how they scale, the time per number of GPU threads is used as a quantity.
+Hence, to be able to better compare how the benchmarks scale, the time per number of GPU threads is used as a quantity.
 Table 1 shows the results.
@@
 The timing results for the last one are shown in the last section of table \ref{resulttable}.
-One can see an improvement of up to a factor around 10.
-However, the robot code is not optimized for the CPU version, as the GPU kernel has just been rewritten with an for loop
-reusing all the device functions for collision detection.
-Therefore this comparison is not quite fair.
+One can see an improvement of up to a factor around 10 between the CPU and GPU benchmarks (based on version 3).
+However, the robot code is not especially optimized for the CPU version, as the GPU kernel has been rewritten with an for loop
+reusing all device functions for collision detection.
+The GPU performance significantly improves from version 1 to 3.
+The code is available on Github, \cite{githubcode}.
 
@@
 	\bibitem{mpi} W. Gropp, Using MPI: Portable Parallel Programming with the Message-Passing Interface, MIT Press 2014
+	\bibitem{githubcode} M. Demmeler, Github Repository GPUPRM, \url{https://github.com/demmeler/GPUPRM}
 \end{thebibliography}

### 2016-07-27 paper/paper.tex
subtitle weg
qs qe, float2,.. schrift

@@
 \title{ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
-\subtitle{High Performance Computing Project}
+%\subtitle{High Performance Computing Project}
 %\date{}
@@
 		0,	& \text{ for no collision of any polytope pair} \\
-		1,	& \text{ if at least one collision occurs },
+		1,	& \text{ if at least one collision occurs }
 	\end{cases}
+	\; .
 \end{equation}
@@
 
-For the implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes float2 for 2D vectors and float4 for 3D vectors.
+For the implementation of the kinematics and later the collision algorithm, a geometry library was written, based on the CUDA datatypes \texttt{float2} for 2D vectors and \texttt{float4} for 3D vectors.
 To guarantee aligned access on the GPU, float4 is used instead of float3. 
@@
 and is load\-ed once at the beginning and copied to the GPU by the 
-\footnotesize
-init
-\normalsize 
+\texttt{init}
 function.
@@
 \begin{algorithm}\label{host}
-	\KwData{qs, qe, N, offset}
+	\KwData{$q_s$, $q_e$, N, $N_{buf}$}
 	\KwResult{res}
@@
 \begin{algorithm}\label{kernel}
-	\KwData{arrays qs, qe, testpos, testnum and threadindex}
+	\KwData{arrays $q_s$, $q_e$, testpos, testnum and threadindex}
 	\KwResult{res}

### 2016-07-27 paper/paper.tex
stand besprechung

@@
 
-This means, that for $i=0,...,n-1$ the vertices with indices $\mathrm{ dest[dsp[i]],..., dest[dsp[i]+cnt[i]-1] }$ are the ones connected to vertex $i$ by an edge.
+This means, that for $i=0,...,n-1$ the vertices with indices \texttt{ dest[dsp[i]],...,dest[dsp[i]+cnt[i]-1] } are the ones connected to vertex $i$ by an edge.
 The polytopes are given by the user as convex hull of a list of vertices.
@@
 \begin{aligned}
-	q_{start}^j &= (q_s^j , q_s^{j+N_{buf}} , ... , q_s^{j+(d-1)*N_{buf}})^T\\
-	q_{end}^j &= (q_e^j , q_e^{j+N_{buf}} , ... , q_e^{j+(d-1)*N_{buf}})^T
+	q_{start}^j &\defgl (q_s^j , q_s^{j+N_{buf}} , ... , q_s^{j+(d-1)*N_{buf}})^T\\
+	q_{end}^j &\defgl (q_e^j , q_e^{j+N_{buf}} , ... , q_e^{j+(d-1)*N_{buf}})^T
 \end{aligned}
@@
 	reset res \;
-	call kernel with $\sum_{j=0}^N \mathrm{testnum[j]}$ threads\;
+	call kernel with $\sum_{j=0}^N \mathrm{testnum[j]}$ threads \;
 	\caption{Host indicator function}
@@
 	$q=c q_{start}^j + (1-c) q_{end}^j $ \;
-	calculate robot kinematics $T_0(q), ..., T_d(q) $\;
-	$\mathrm{res\_tmp} = 0; $ \;
+	calculate robot kinematics $T_0(q), ..., T_d(q) $ \;
+	$\mathrm{res\_tmp} = 0$ \;
 	\For{all registered pairs pairs of polytopes $P, Q$}{

### 2016-07-27 paper/paper.tex
stand besprechung ..

@@
 
-\title{ Parallelization of the Probabilistic Roadmap Method with GPU Acceleration}
+\journalname{Journal of Intelligent \& Robotic Systems}
+\title{ Parallel Algorithms for the Probabilistic Roadmap Method for Robotic Motion Planning with GPU Acceleration}
 %\subtitle{High Performance Computing Project}
@@
 Manfred Liebmann
-\at todo ... \\  %TODO: which adresses to use?
-\email{todo}
+\at Technische Universität München\\
+%Chair of Optimal Control,
+Center for Mathematical Sciences, M17 \\
+Boltzmannstraße 3,
+85748 Garching by Munich \\  %TODO: which adresses to use?
+\email{manfred.liebmann@tum.de}
 }
@@
 
-\section{Abstract}
-In trajectory optimization with obstacles, it is often necessary to provide feasible initial trajectories. 
+\begin{abstract}
+In trajectory optimization with obstacles for robot motion planning, it is often necessary to provide feasible initial trajectories. 
 For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive. 
 A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. 
-An efficient approximative solution for this is using probabilistic roadmaps (PRMs).
-
+An efficient approximative solution for this problem is the probabilistic roadmap method (PRM).
 In this paper, a parallelized version of the PRM method has been developed in general.
-A robotics application then was implemented as an special case of this with GPU support.
+A robotics application then was implemented as a special case of this with GPU support.
+\keywords{probabilistic roadmap method \and robot motion planning \and parallel algorithms \and GPU acceleration}
+\end{abstract}
 

### 2016-07-30 paper/paper.tex
stand
neues abstract

@@
 \journalname{Journal of Intelligent \& Robotic Systems}
-\title{ Parallel Algorithms for the Probabilistic Roadmap Method for Robotic Motion Planning with GPU Acceleration}
+\title{Parallel Algorithms for the Probabilistic Roadmap Method for Robotic Motion Planning with GPU Acceleration}
 %\subtitle{High Performance Computing Project}
@@
 
-\institute{Manuel Demmeler
-\at todo ... \\ 
-\email{manuel.demmeler@yahoo.de}
-\and
-Manfred Liebmann
+\institute{Manuel Demmeler, Manfred Liebmann\\ 
+%\at Technische Universität München\\
+%Center for Mathematical Sciences, M17 \\
+%Boltzmannstraße 3,
+%85748 Garching by Munich \\ 
+%\email{manuel.demmeler@yahoo.de}
+%\and
+%Manfred Liebmann
 \at Technische Universität München\\
@@
 Boltzmannstraße 3,
-85748 Garching by Munich \\  %TODO: which adresses to use?
-\email{manfred.liebmann@tum.de}
+85748 Garching by Munich \\ 
+\email{manfred.liebmann@tum.de, manuel.demmeler@yahoo.de}
 }
@@
 \begin{abstract}
+%
 In trajectory optimization with obstacles for robot motion planning, it is often necessary to provide feasible initial trajectories. 
-For example in robot motion planning, where the dimensions of the underlying spaces are rather high, this can be expensive. 
-A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. 
-An efficient approximative solution for this problem is the probabilistic roadmap method (PRM).
-In this paper, a parallelized version of the PRM method has been developed in general.
-A robotics application then was implemented as a special case of this with GPU support.
+A common algorithm is the probabilistic roadmap method (PRM), which gives efficient approximative solutions and scales well with high dimensions of the underlying spaces.
+%
+%As the dimensions of the underlying spaces are rather high, this can be expensive. 
+%A further problem is that because of the transformation from work into configuration space, the obstacles are often no longer given in a closed form. 
+%An efficient approximative solution for this problem is the probabilistic roadmap method (PRM).
+%In this paper, a parallelized version of the PRM method has been developed in general.
+%A robotics application then was implemented as a special case of this with GPU support.
+%
+In this paper, the PRM method has been parallelized.
+A first level of parallelization is achieved by GPU support.
+A CUDA kernel has been implemented that runs robot kinematics calculations and collision detection for different robot configurations in parallel.
+For a further level of parallelization, different approaches for the usage of multiple GPUs have been developed and benchmarked.
+%
 \keywords{probabilistic roadmap method \and robot motion planning \and parallel algorithms \and GPU acceleration}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{configspacetrafo.png}
+    \includegraphics[width=0.7\textwidth]{configspacetrafo.png}
     \caption{Example from \cite{prmlec} for an indicator function in robotics. 
@@
 As the indicator function has also been implemented on the CPU, all versions could be run without GPU.
-The timing results for the last one are shown in the last section of table \ref{resulttable}.
+The timing results for the CPU version are shown in the last section of table \ref{resulttable}.
 One can see an improvement of up to a factor around 10 between the CPU and GPU benchmarks (based on version 3).

### 2016-07-31 paper/paper.tex
kleine quellenergaenzung: roboterdynamik buch

@@
 For a robot with given joint angles \(q_i\) one can calculate straightforward all positions and orientations  of the single parts of the robot.
-For this, coordinate frames according to the Denavit-Hartenberg (DH) \cite{robodyn} convention are used.
+For this, coordinate frames according to the Denavit-Hartenberg (DH) \cite{robodyn}, \cite{robodynbuch} convention are used.
 The coordinate frames of reference of the single parts are described by transformation matrices 
@@
 there exists a set of parameters, such that the above definition leads to the correct transformations for this robot.
-For the exact calculations, see \cite{robodyn}.
+For the exact calculations, see \cite{robodyn}, \cite{robodynbuch}.
 
@@
 	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München 2014
+	\bibitem{robodynbuch} F. Pfeiffer, E. Reithmeier, Roboterdynamik, B.G. Teubner Verlag 1987
 	\bibitem{chung} K.

### 2016-08-04 paper/paper.tex
Abstract updates

@@
 \institute{Manuel Demmeler, Manfred Liebmann\\ 
-%\at Technische Universität München\\
+%\at Technische Universitt Mnchen\\
 %Center for Mathematical Sciences, M17 \\
-%Boltzmannstraße 3,
+%Boltzmannstra§e 3,
 %85748 Garching by Munich \\ 
@@
 %Manfred Liebmann
-\at Technische Universität München\\
+\at Technische Universit\"at M\"unchen\\
 %Chair of Optimal Control,
 Center for Mathematical Sciences, M17 \\
-Boltzmannstraße 3,
+Boltzmannstra{\ss}e 3,
 85748 Garching by Munich \\ 
-\email{manfred.liebmann@tum.de, manuel.demmeler@yahoo.de}
+\email{manuel.demmeler@tum.de, manfred.liebmann@tum.de}
 }
@@
 %
-In trajectory optimization with obstacles for robot motion planning, it is often necessary to provide feasible initial trajectories. 
-A common algorithm is the probabilistic roadmap method (PRM), which gives efficient approximative solutions and scales well with high dimensions of the underlying spaces.
+In trajectory optimization with obstacles for robot motion planning it is often necessary to provide feasible initial trajectories. 
+A common algorithm to solve this problem is the probabilistic roadmap method (PRM), which gives efficient approximative solutions and scales well with the high dimensional spaces underlying the problem.
 %
@@
 %
-In this paper, the PRM method has been parallelized.
-A first level of parallelization is achieved by GPU support.
-A CUDA kernel has been implemented that runs robot kinematics calculations and collision detection for different robot configurations in parallel.
-For a further level of parallelization, different approaches for the usage of multiple GPUs have been developed and benchmarked.
+In this paper, the parallelization of the PRM method specifically for graphics processing units (GPUs) is investigated.
+%A first level of parallelization is achieved by GPU support.
+CUDA kernels have been implemented to run the robot kinematics calculations and collision detection for different robot configurations in parallel.
+Three different parallelization approaches to improve the performance for multi GPU configurations are analyzed and the resulting benchmarks are presented.
+%For a further level of parallelization, different approaches for the usage of multiple GPUs have been developed and benchmarked.
 %
@@
 	\bibitem{prm2} S. M. LaValle, Planning Algorithms, Cambridge University Press 2006
-	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl für Angewandte Mechanik, TU München 2014
+	\bibitem{robodyn} T. Buschmann, Skript zur Vorlesung: Roboterdynamik SS14, Lehrstuhl f\"ur Angewandte Mechanik, TU M\"unchen 2014
 	\bibitem{robodynbuch} F. Pfeiffer, E. Reithmeier, Roboterdynamik, B.G. Teubner Verlag 1987
